### 2016-11-14 SimulatedData.Rmd
intial website

---
title: Hierarchical Nmixture models for species interactions - Simulations
author: "Ben Weinstein - Stony Brook University"
output:
  html_document:
    toc: true
    number_sectionsf: true
    theme: spacelab
    keep_md: true
  word_document: default
---

```{r,warning=FALSE,message=FALSE,echo=FALSE,cache=FALSE}
library(reshape2)
library(foreach)
library(chron)
library(ggplot2)
library(knitr)
library(R2jags)
library(dplyr)
library(stringr)
library(gridExtra)
library(boot)
opts_chunk$set(message=FALSE,warning=FALSE,fig.width=5,fig.height=4,echo=TRUE,cache=F,fig.align='center',fig.path="figure/")

set.seed(3)
#source functions
source("Bayesian/BayesFunctions.R")

```

```{r,echo=F,cache=FALSE}
paste("Run Completed at",Sys.time())
```

#Simulation

## Parameters

* 10 hummingbird species
* 10 plant species
* Range of hummingbird bill sizes (in mm) ~ Pois(10)/10
* Range of corolla sizes (in mm) ~ Pois(15)/10
* Mean frequeny ($\lambda$) for each hummingbird is drawn from U(0,10)  
* Trait matching (minimizing Bill-Corolla difference) is drawn from a hierarcichal distribution
$$log(\lambda)<-\alpha_i + \beta_i *traitmatch$$
$$\alpha=N(3,0.2)$$
$$\beta1 = N(-1,0.2)$$

* Imperfect detection 
* $$ p_i = U(0.1,0.9) $$ 
* 10 camera
* 3 days per camera

**View simulated strength and form of trait matching **

```{r}
#load("AbundanceSimulation.RData")
gc()
```

#Simulation Parameters

```{r,fig.height=5,fig.width=8,eval=T}
#Number of hummingbird species
h_species=10
plant_species=10
cameras<-5
days<-3

#Bill sizes
Bill<-rpois(h_species,10)

#Corolla sizes
Corolla<-rpois(plant_species,15)

#Subtract both and take absolute value, convert cm
traitmatch<-abs(sapply(Corolla,function(x) x - Bill)/10)

#regression slopes
#traits
beta1_mu<- -1

#abundance

#species variance in slopes
beta1_sigma<- 0.2

#Species alpha_mu 
alpha_mu<- 3
alpha_sigma<- 0.2

#Poisson overdipersion
tauE<-10

#species level
detection= inv.logit(rnorm(h_species,0,1.67))
beta1<-rnorm(h_species,beta1_mu,beta1_sigma)
alpha<-rnorm(h_species,alpha_mu,alpha_sigma)
```

#Compute resource abundance

```{r,eval=T}
resources<-array(NA,dim=c(h_species,plant_species,cameras))

#fill for each month
for (x in 1:cameras){
   resources[,,x]<-rpois(1,10)   
  }
resources<-array(data=scale(resources),dim=c(h_species,plant_species,cameras))
```

#Compute true interaction matrices

```{r,eval=T}
#for each species loop through and create a replicate dataframe
obs<-array(dim=c(h_species,plant_species,cameras,days))
lambda<-array(dim=c(h_species,plant_species,cameras))
N<-array(dim=c(h_species,plant_species,cameras))

#create intensities
for(x in 1:h_species){
  for (y in 1:plant_species){
    for (z in 1:cameras){
      lambda[x,y,z]<-exp(alpha[x] + beta1[x] * traitmatch[x,y] + rnorm(1,0,1/sqrt(tauE)))
  }
  }
}

#draw latent states
for(x in 1:h_species){
  for (y in 1:plant_species){
    for (z in 1:cameras){
      # true latent count
      N[x,y,z]<-rpois(1,lambda[x,y,z])
    }
  }
}

#Observed counts in each day
for(x in 1:h_species){
  for (y in 1:plant_species){
    for (z in 1:cameras){
        for (d in 1:days){
      #true detection rate of that observed count
      obs[x,y,z,d]<-rbinom(1,N[x,y,z],p=detection[x])
      }
    }
  }
}
```

##View correlation in simulated latent state

```{r}
mdat<-melt(N)
colnames(mdat)<-c("Bird","Plant","Camera","Interactions")

traitmelt<-melt(traitmatch)
colnames(traitmelt)<-c("Bird","Plant","traitmatch")

mdat<-merge(mdat,traitmelt,c("Bird","Plant"))
ggplot(mdat,aes(x=traitmatch,y=Interactions,col=as.factor(Bird))) + geom_point() + geom_smooth(aes(group=1),method="glm",method.args = list(family = "poisson")) + labs(col="Bird") + xlab("Absolute value of Bill Length - Corolla Length ")

#Merge resources
mr<-melt(resources)
colnames(mr)<-c("Bird","Plant","Camera","Abundance")
mdat<-merge(mdat,mr,by=c("Bird","Plant","Camera"))
```

##View Detection Rates

```{r}
obs.state<-melt(obs)
colnames(obs.state)<-c("Bird","Plant","Camera","Day","Yobs")
obs.state<-merge(mdat,obs.state,by=c("Bird","Plant","Camera"))
ggplot(obs.state,aes(x=Interactions,y=Yobs,col=Camera)) + geom_point() + theme_bw() + geom_abline() + coord_equal()
```

# Hierarcichal Nmixture Model

For hummingbird i visiting plant j recorded by camera k on day d:

$$ Y_{i,j,k,d} \sim Binom(N_{i,j,k},\omega_i)$$
$$N_{i,j,k} \sim Pois(\lambda_{i,j}) $$
$$log(\lambda_{i,j})<-\alpha_i + \beta_{1,i} * |Bill_i - Corolla_j| * \beta_{2,i} * Abundance_{j,k} + \epsilon$$
$$ \epsilon \sim Normal(0,\tau_\epsilon)$$

**Priors**

Please recall that jags parameterizes models using precision, not sd (precision = 1/sd^2)

$$\omega_i \sim (\mu_{\omega},\tau_{\omega})$$  $$\mu_{\omega} \sim Normal(0,0.5)   
$$\tau_{\omega} \sim Uniform(0,10)

$$\alpha_i \sim Normal(\mu_{\alpha},\tau_{\alpha})$$
$$\beta_{i,1} \sim Normal(\mu_{\beta_1},\tau_{\beta_1})$$

**Hyperpriors**
$$\mu_{\alpha} \sim Normal(0,0.0001)$$
$$\mu_{\beta_1} \sim Normal(0,0.0001)$$
$$\mu_{\beta_2} \sim Normal(0,0.0001)$$

$$\tau_{\alpha} \sim Half-T(0.0001,0.0001)$$
$$\sigma_{\alpha} = \sqrt[2]{\frac{1}{\tau_\alpha}}$$

$$\tau_{\beta_1} \sim \sqrt[2]{\frac{1}{\sigma_{\beta_1}}}$$
$$\sigma_{\beta_1} \sim Half-T(0,1)$$

$$\tau_{\beta_2} \sim \sqrt[2]{\frac{1}{\sigma_{\beta_2}}}$$
$$\sigma_{\beta_2} \sim Half-T(0,1)$$

#Simulated data without detection

```{r,eval=T}
runs<-30000
#Source model
source("Bayesian/NoDetectNmixturePoissonRagged.R")

#print model
print.noquote(readLines("Bayesian//NoDetectNmixturePoissonRagged.R"))

  #for parallel run
  Yobs=obs.state$Yobs
  Bird=obs.state$Bird
  Plant=obs.state$Plant
  Plants=max(obs.state$Plant)
  Cameras=max(obs.state$Camera)
  Camera=obs.state$Camera
  Traitmatch=traitmatch
  Birds=max(obs.state$Bird)
  Nobs=length(obs.state$Yobs)
  resources=resources

  #Inits
  Ninit<-array(dim=c(h_species,plant_species,Cameras),data=1)

  InitStage <- function() {list(epsilon=Ninit)}

  #Parameters to track
  ParsStage <- c("alpha","beta1","alpha_mu","beta1_sigma","alpha_sigma","ynew","beta1_mu","fit","fitnew","tauE")

  #MCMC options
  ni <- runs  # number of draws from the posterior
  nt <- 4   #thinning rate
  nb <- max(0,runs-2000) # number to discard for burn-in
  nc <- 2  # number of chains

  Dat<-list("Yobs","Bird","Plant","Plants","Traitmatch","Birds","Nobs","Cameras","Camera","resources","Ninit")

    sim_niave<-do.call(jags.parallel,list(data=Dat,InitStage,parameters.to.save=ParsStage,model.file="Bayesian/NoDetectNmixturePoissonRagged.jags",n.thin=nt, n.iter=ni,n.burnin=nb,n.chains=nc))

```

```{r,eval=F}
#recompile if needed
load.module("dic")
runs<-5000
recompile(sim_niave)
sim_niave<-update(sim_niave,n.iter=runs,n.burnin=runs*.95,n.thin=10)
```

```{r}
pars_niave<-extract_par(sim_niave,data=obs.state,ynew=T)
pars_niave$Model<-c("Poisson GLMM")
```

##Assess Convergence

```{r,cache=FALSE,fig.width=11,fig.height=5}
ggplot(pars_niave[pars_niave$par %in% c("alpha","beta1"),],aes(x=Draw,y=estimate,col=as.factor(Chain))) + geom_line() + facet_grid(par~species,scale="free") + theme_bw() + labs(col="Chain") + ggtitle("Detection Probability")
```

```{r,fig.height=5,fig.width=11}
ggplot(pars_niave[pars_niave$par %in% c("beta1_mu","beta1_sigma","tauE"),],aes(x=Draw,y=estimate,col=as.factor(Chain))) + geom_line() + theme_bw() + labs(col="Chain") + ggtitle("Trait-matching regression") + facet_wrap(~par,scales="free")
```

##Posteriors

```{r,cache=FALSE,fig.width=7,fig.height=5}
###Posterior Distributions
p<-ggplot(pars_niave[pars_niave$par %in% c("alpha","beta1"),],aes(x=estimate)) + geom_histogram() + ggtitle("Estimate of parameters") + facet_grid(species~par,scales="free") + theme_bw() + ggtitle("Species Posteriors")

#Add true values
tr<-melt(data.frame(species=1:length(detection),alpha=alpha,beta1=beta1),id.var='species')
colnames(tr)<-c("species","par","value")
psim<-p + geom_vline(data=tr,aes(xintercept=value),col='red',linetype='dashed',size=1)
psim
ggsave("Figures/SimulationPosteriorsNoDetect.jpg",dpi=300,height=8,width=8)
```

```{r,cache=FALSE,fig.height=3,fig.width=10}
p<-ggplot(pars_niave[pars_niave$par %in% c("beta1_mu","alpha_mu","alpha_sigma","beta1_sigma","tauE"),],aes(x=estimate)) + geom_histogram() + ggtitle("Hierarchical Posteriors") + facet_grid(~par,scale="free") + theme_bw()

#Add true values
tr<-melt(list(beta1_mu=beta1_mu,alpha_mu=alpha_mu,alpha_sigma=alpha_sigma,beta1_sigma=beta1_sigma,tauE=tauE))

colnames(tr)<-c("value","par")

psim2<-p + geom_vline(data=tr,aes(xintercept=value),linetype='dashed',size=1,col="red")
```

**True values are given in the dashed lines.**

##Predicted Relationship

```{r,fig.height=4,fig.width=4}
castdf<-group_by(pars_niave,Chain) %>% select(par,estimate) %>% filter(par %in% c("beta1_mu","alpha_mu"))

castdf<-dcast(pars_niave[pars_niave$par %in% c("beta1_mu","alpha_mu"),], Chain + Draw~par,value.var="estimate")

#calculated predicted y
predyniave_trait<-trajF(alpha=castdf$alpha_mu,beta1=castdf$beta1_mu,trait=as.numeric(traitmatch),resources=as.numeric(resources[,1,]),tauSigma = 1/sqrt(tauE))

predyniave_both<-trajF(alpha=castdf$alpha_mu,beta1=castdf$beta1_mu,trait=as.numeric(traitmatch),resources=as.numeric(resources[,1,]),tauSigma=1/sqrt(tauE))
```

# Simulated data with detection

```{r,eval=T}
runs<-30000

#Source model
source("Bayesian/NmixturePoissonRagged.R")

#print model
print.noquote(readLines("Bayesian//NmixturePoissonRagged.R"))

  #for parallel run
  Yobs=obs.state$Yobs
  Bird=obs.state$Bird
  Plant=obs.state$Plant
  Time=obs.state$Camera
  Times=max(obs.state$Camera)
  Traitmatch=traitmatch
  Birds=max(obs.state$Bird)
  Plants=max(obs.state$Plant)
  Nobs=length(obs.state$Yobs)
  resources=resources

  #A blank Y matrix - all present
  Ninit<-array(dim=c(h_species,plant_species,Times),data=max(obs.state$Yobs)*1)

  #Inits
  InitStage <- function() {list(N=Ninit,epsilon=Ninit/10)}

  #Parameters to track
  ParsStage <- c("detect","tauE","alpha","beta1","alpha_mu","alpha_sigma","beta1_sigma","ynew","beta1_mu","fit","fitnew")

  #MCMC options
  ni <- runs  # number of draws from the posterior
  nt <- 8  #thinning rate
  nb <- max(0,runs-2000) # number to discard for burn-in
  nc <- 2  # number of chains

  Dat<-list("Yobs","Bird","Plant","Plants","Traitmatch","resources","Birds","Nobs","Ninit","Time","Times")

    system.time(sim_detect<-do.call(jags.parallel,list(Dat,InitStage,ParsStage,model.file="Bayesian/NmixturePoissonRagged.jags",n.thin=nt, n.iter=ni,n.burnin=nb,n.chains=nc)))
```

```{r,eval=F}
#recompile if needed
load.module("dic")
runs<-10000
recompile(sim_detect)
system.time(sim_detect<-update(sim_detect,n.iter=runs,n.burnin=runs*.95,n.thin=10,parameters.to.save=ParsStage))
```

```{r}
pars<-extract_par(sim_detect,data=obs.state,ynew=T)
pars$Model<-"Nmixture"
```

##Assess Convergence

```{r,cache=FALSE,fig.width=11,fig.height=5}
ggplot(pars[pars$par %in% c("detect","alpha","beta1"),],aes(x=Draw,y=estimate,col=as.factor(Chain))) + geom_line() + facet_grid(par~species,scale="free") + theme_bw() + labs(col="Chain") + ggtitle("Detection Probability")
```

```{r,fig.height=5,fig.width=11}
ggplot(pars[pars$par %in% c("beta1_mu","alpha_sigma","beta1_sigma","tauE"),],aes(x=Draw,y=estimate,col=as.factor(Chain))) + geom_line() + theme_bw() + labs(col="Chain") + ggtitle("Trait-matching regression") + facet_wrap(~par,scales="free")
```

##Posteriors

```{r,cache=FALSE,fig.width=7,fig.height=5}
###Posterior Distributions
p<-ggplot(pars[pars$par %in% c("detect","alpha","beta1"),],aes(x=estimate)) + geom_histogram() + ggtitle("Estimate of parameters") + facet_grid(species~par,scales="free") + theme_bw() + ggtitle("Species Posteriors")

#Add true values
tr<-melt(data.frame(species=1:length(detection),detect=detection,alpha=alpha,beta1=beta1),id.var='species')
colnames(tr)<-c("species","par","value")
psim<-p + geom_vline(data=tr,aes(xintercept=value),col='red',linetype='dashed',size=1)
#ggsave("Figures/SimulationPosteriors.jpg",dpi=300,height=8,width=8)
```

```{r,cache=FALSE,fig.height=3,fig.width=10}
p<-ggplot(pars[pars$par %in% c("beta1_mu","alpha_mu","alpha_sigma","beta1_sigma","tauE"),],aes(x=estimate)) + geom_histogram() + ggtitle("Hierarchical Posteriors") + facet_wrap(~par,scale="free",nrow=2) + theme_bw()

#Add true values
tr<-melt(list(beta1_mu=beta1_mu,alpha_mu=alpha_mu,alpha_sigma=alpha_sigma,beta1_sigma=beta1_sigma,tauE=tauE))

colnames(tr)<-c("value","par")

psim2<-p + geom_vline(data=tr,aes(xintercept=value),linetype='dashed',size=1,col="black")
#ggsave("Figures/SimulationH.jpg",dpi=300,height=4,width=10)
```

**True values are given in the dashed lines.**

##Compare simulation posteriors with and without detection

```{r,cache=FALSE,fig.width=8,fig.height=9}
#Bind to other dataset
parsall<-rbind.data.frame(pars[!pars$par %in% "ynew",],pars_niave[!pars_niave$par %in% "ynew",])
parsall$Model<-as.factor(parsall$Model)

###Posterior Distributions
p<-ggplot(parsall[parsall$par %in% c("detect","alpha","beta1"),],aes(x=estimate,fill=Model)) + geom_histogram(position="identity") + ggtitle("Estimate of parameters") + facet_grid(species~par,scales="free") + theme_bw()

#Add true values
tr<-melt(data.frame(species=1:length(detection),detect=detection,alpha=alpha,beta1=beta1),id.var='species')
colnames(tr)<-c("species","par","value")
psim<-p + geom_vline(data=tr,aes(xintercept=value),col='black',linetype='dashed',size=1)
psim
#ggsave("Figures/SimulationPosteriorsBoth.jpg",dpi=300,height=8,width=8)
```

```{r,cache=FALSE,fig.height=3,fig.width=10}
p<-ggplot(parsall[parsall$par %in% c("beta1_mu","alpha_mu","alpha_sigma","beta1_sigma","tauE_mu","tauE_tau"),],aes(x=estimate,fill=Model)) + geom_histogram(position="identity") + ggtitle("Hierarchical Posteriors") + facet_wrap(~par,scale="free",nrow=2) + theme_bw()

#Add true values
tr<-melt(list(beta1_mu=beta1_mu,alpha_mu=alpha_mu,alpha_sigma=alpha_sigma,beta1_sigma=beta1_sigma,tauE=tauE))

colnames(tr)<-c("value","par")

psim2<-p + geom_vline(data=tr,aes(xintercept=value),linetype='dashed',size=1,col="black")
psim2
#ggsave("Figures/SimulationHBoth.jpg",dpi=300,height=4,width=10)
```

##Strip plots
Plot the posterior mean density
```{r,fig.width=15,fig.height=13}
spars<-parsall %>% filter(par %in% c("alpha","beta1","detect")) %>% group_by(Model,species,par) %>% summarize(mean=mean(estimate),lower=quantile(estimate,0.05),upper=quantile(estimate,0.95))

tr<-melt(data.frame(species=1:length(detection),detect=detection,alpha=alpha,beta1=beta1,tauE=tauE),id.var='species')
colnames(tr)<-c("species","par","value")

ggplot(spars,aes(x=Model,ymin=lower,ymax=upper,y=mean,col=Model)) + geom_linerange(size=1.3) + facet_grid(par~species,scales="free") + geom_hline(data=tr,aes(yintercept=value),linetype='dashed',size=1,col="black") + geom_point(aes(y=mean),col='grey50',size=2) + theme_bw() + ylab("Estimate")

#Hierarchical posteriors
hpars<-parsall %>% filter(par %in% c("beta1_mu","alpha_mu","alpha_sigma","beta1_sigma","tauE")) %>% group_by(Model,species,par) %>% summarize(mean=mean(estimate),lower=quantile(estimate,0.05),upper=quantile(estimate,0.95))

tr<-melt(data.frame(species=1:length(detection),beta1_mu=beta1_mu,alpha_mu=alpha_mu,beta1_sigma=beta1_sigma,alpha_sigma=alpha_sigma,tauE=tauE),id.var='species')
colnames(tr)<-c("species","par","value")

ggplot(hpars,aes(x=Model,ymin=lower,ymax=upper,y=mean,col=Model)) + geom_linerange(size=1.3) + facet_wrap(~par,scales="free",nrow=1) + geom_hline(data=tr,aes(yintercept=value),linetype='dashed',size=1,col="black") + geom_point(aes(y=mean),col='grey50',size=2) + theme_bw() + ylab("Estimate") 
#ggsave("Figures/StripPlots.svg",height=4,width=9)
```

##Correlation in posteriors for Nmixture Model

```{r}
castdf<- pars %>% filter(Model =="Nmixture") %>% group_by(Chain) %>% select(par,estimate,Draw) %>% filter(par %in% c("beta1_mu","alpha_mu")) %>% dcast(Chain+Draw~par,value.var="estimate")

pairs(castdf[,3:4],main="Correlation in Group-Level Posteriors")

castdf<- pars %>% filter(Model =="Nmixture") %>% group_by(Chain) %>% select(par,estimate,Draw,species) %>% filter(par %in% c("alpha","beta1","detect","tauE")) %>% dcast(species+Chain+Draw~par,value.var="estimate")

pairs(castdf[,4:7],main="Correlation in Species-Level Posteriors")
```

##Predicted Relationship

```{r,fig.height=4,fig.width=4}

castdf<-dcast(pars[pars$par %in% c("beta1_mu","alpha_mu","tauE"),], Chain + Draw~par,value.var="estimate")
```

## Calculated predicted visitation rates

### Traits

```{r}
predy_trait<-trajF(alpha=castdf$alpha_mu,beta1=castdf$beta1_mu,tauSigma =1/sqrt(castdf$tauE),trait=as.numeric(traitmatch),resources=as.numeric(resources))
orig<-trajF(alpha=rnorm(1000,alpha_mu,alpha_sigma),beta1=rnorm(1000,beta1_mu,beta1_sigma),trait=as.numeric(traitmatch),resources=as.numeric(resources),tauSigma =1/sqrt(tauE))

pm<-melt(list(Nmixture=predy_trait,'Poisson GLMM'=predyniave_trait),id.vars=colnames(predy_trait))

tplot<-ggplot(data=pm[,],aes(x=trait)) + geom_ribbon(aes(ymin=lower,ymax=upper,fill=L1),alpha=0.5)  + geom_line(aes(y=mean,col=L1),size=.4,linetype="dashed") + theme_bw() + ylab("Daily Interactions") + xlab("Difference between Bill and Corolla Length") + geom_point(data=mdat,aes(x=traitmatch,y=Interactions),size=.5,alpha=.5) + labs(fill="Model",col="Model") + ggtitle("Simulation: Traits") + geom_line(data=orig,aes(y=mean),col='black',size=.8) + scale_fill_manual(values=c("grey70","Black"))+ scale_color_manual(values=c("grey70","Black"))
tplot
ggsave("Figures/SimPredictBoth.jpg",height=5,width=7)
```

Black line is the true relationship. The red line is the posterior mean with confidible intervals in shaded grey for the proposed bayesian model. The blue line is the same model, but assuming perfect detection rates.

**Conclusion:** Accounting for detection and non-independence greatly increases the accuracy of the predicted state. The perfect detection model underestimates the strength of trait matching among hummingbirds and their foodplants.

##Posterior Predictive Check

Since I have simualted the data, it should fit as well as any random dataset drawn from the estimated parameters. An ideal fit would be posterior values sitting along the 1:1 line.

```{r,fig.height=4,fig.width=8}
fitstat<-droplevels(parsall[parsall$par %in% c("fit","fitnew"),])
fitstat<-dcast(fitstat,Draw+Chain+Model~par,value.var="estimate")

#add 1:1 line
ymin<-round(min(c(fitstat$fit,fitstat$fitnew)))
ymax<-round(max(c(fitstat$fit,fitstat$fitnew)))
ab<-data.frame(x=ymin:ymax,y=ymin:ymax)
p<-ggplot(fitstat,aes(x=fit,y=fitnew)) + geom_point(aes(col=Model)) + theme_bw() + coord_equal()
psim4<-p  + labs(x="Discrepancy of observed data",y="Discrepancy of replicated data",col="Model") + geom_line(data=ab,aes(x=x,y=y)) + ggtitle("Simulated Data")
psim4
ggsave("Figures/SimulationDisc.jpeg",height=5,width=5)
```

```{r,echo=FALSE}
#Write Figures as panels
jpeg("Figures/SimulationPosteriorsBoth.jpg",height=12,width=12,units="in",res=300)
grid.arrange(psim,psim2,ncol=1,heights=c(3/4,1/4))
dev.off()

```

#By species

#Compare using true known interactions

## No Detection Nmixture Model

```{r}
true_state<-obs.state %>% group_by(Bird,Plant) %>% summarize(n=sum(Yobs)) %>% acast(.$Bird~.$Plant)

N_niave<-pars_niave %>% filter(par== "ynew")

#Discrepancy function
#define discrep function
chisq<-function(o,e){(o-e)^2/(e+0.5)}

bydraw<-split(N_niave,list(N_niave$Chain,N_niave$Draw))

#Don't need giant matrix sitting around
rm(N_niave)
gc()

occ_nodetect_matrix<-lapply(bydraw,function(x){
  r<-acast(x,species ~ plant,value.var = "estimate",fun.aggregate = sum)
})

rm(bydraw)
gc()

#calculate discrep on those deviates
occ_nodetect<-lapply(occ_nodetect_matrix,function(r){
    #for each position what is the chisq
  rmerge<-matrix(nrow = nrow(true_state),ncol=ncol(true_state))
  for (x in 1:nrow(r)){
    for (y in 1:ncol(r)){
     rmerge[x,y]<-chisq(e=r[x,y],o=true_state[x,y])
      }
    }
  return(rmerge)
})

names(occ_nodetect)<-1:length(occ_nodetect)
names(occ_nodetect_matrix)<-1:length(occ_nodetect_matrix)
```

##With Detection

```{r}
Ndetect<-pars %>% filter(par== "ynew")
bydraw<-split(Ndetect,list(Ndetect$Chain,Ndetect$Draw))

#Don't need giant matrix around
rm(Ndetect)
gc()

occ_matrix<-lapply(bydraw,function(x){
  r<-acast(x,species ~ plant,value.var = "estimate",fun.aggregate = sum)
  })

rm(bydraw)
gc()

#calculate discrep for those aggregated matrices
occ<-lapply(occ_matrix,function(r){
    #for each position what is the chisq
  rmerge<-matrix(nrow = nrow(true_state),ncol=ncol(true_state))
  for (x in 1:nrow(r)){
    for (y in 1:ncol(r)){
     rmerge[x,y]<-chisq(e=r[x,y],o=true_state[x,y])
      }
    }
  return(rmerge)
})

names(occ)<-1:length(occ)
names(occ_matrix)<-1:length(occ_matrix)
```

##Compare to observed data

```{r,fig.width=11}
#true number of observed interactions

mmat<-melt(true_state)
colnames(mmat)<-c("Bird","Plant","True_State")

#append to predicted matrices

#Nmixture with detection
mocc<-melt(occ_matrix)
colnames(mocc)<-c("Bird","Plant","Nmixture","Iteration")
simdat<-merge(mocc,mmat,by=c("Bird","Plant"),all.x=T)

#Nmixture with nodetection
moccd<-melt(occ_nodetect_matrix)
colnames(moccd)<-c("Bird","Plant","Poisson GLMM","Iteration")

simdat<-merge(simdat,moccd,by=c("Bird","Plant","Iteration"))

simdat<-melt(simdat,measure.vars = c("Nmixture","Poisson GLMM"))

ggplot(simdat,aes(x=True_State,y=value,col=variable),alpha=1) + geom_point() + geom_abline() + labs(col="Model") + ylab("Predicted State") + xlab("True State") + theme_bw() + facet_wrap(~variable)

ggsave("Figures/PredictedState.jpeg",height=3,width=8)

ggplot(simdat[simdat$variable %in% c("Nmixture","Poisson GLMM"),],aes(x=True_State,y=value,col=variable)) + geom_point(alpha=.3) + geom_abline() + labs(col="Model") + ylab("Predicted State") + xlab("True State") + theme_bw()
ggsave("Figures/PredictedState_noM.jpeg",height=3,width=8)

#difference in the middle
simd<-dcast(simdat,...~variable)
simd$Diff<-simd$Nmixture-simd$`Poisson GLMM`
ggplot(simd,aes(x=True_State,y=abs(Diff))) + geom_point() + ylab("|Nmixture - Poisson GLMM|") + theme_bw() + labs(x="True State")
ggsave("Figures/Difference_Pred.jpeg",height=4,width=6)
```

## View predicted trait-matching relationship with the number of visits.
```{r}
simdat<-merge(simdat,traitmelt,c("Bird","Plant"))
mmat<-merge(mmat,traitmelt,by=c("Bird","Plant"))
```

#Predicted total number of visits based on morphology

```{r,fig.height=5,fig.width=5}
simT<-simdat %>% group_by(variable,traitmatch) %>% summarize(Lower=quantile(value,0.05),Upper=quantile(value,0.95),y=mean(value))

ggplot(simT,aes(x=traitmatch)) + geom_ribbon(aes(ymin=Lower,ymax=Upper,fill=variable),alpha=0.4) + geom_line(aes(y=y,col=variable),linetype='dashed') + theme_bw()  + geom_point(data=mmat,aes(y=True_State)) + labs(x="Difference in Bill and Corolla Length",y="Total Predicted Visits",fill="Model",col='Model')

```

View a couple example data points from across the type of interactions.

```{r}
h<-simdat[which.max(simdat$True_State),c("Bird","Plant")]
d<-simdat[simdat$Bird %in% h$Bird & simdat$Plant %in% h$Plant,]

ggplot(data=d,aes(x=value,fill=variable))+ geom_histogram(position="identity") + labs(fill="Model") + geom_vline(aes(xintercept=True_State)) + ggtitle("High Visitation Example")

h<-simdat[which.min(simdat$True_State),c("Bird","Plant")]
d<-simdat[simdat$Bird %in% h$Bird & simdat$Plant %in% h$Plant,]

ggplot(data=d,aes(x=value,fill=variable))+ geom_histogram(position="identity") + labs(fill="Model") + geom_vline(aes(xintercept=True_State)) + ggtitle("Low Visitation Example")
```

##Summary of discrepancy of predicted matrices

```{r}
#Nmixture without detection
occno_disc<-sapply(occ_nodetect,function(x) median(x))

#Nmixture with detection
occ_disc<-sapply(occ,function(x) median(x))

#compared to bayesian
ggplot() + xlab("Chi-squared Discrepancy") + geom_histogram(data=data.frame(occ_disc),aes(x=occ_disc),fill="red",alpha=.6) + theme_bw() +geom_vline(aes(xintercept=mean(occ_disc)),linetype="dashed",col="red") + geom_histogram(data=data.frame(occno_disc),aes(x=occno_disc),fill="orange",alpha=.6) + geom_vline(aes(xintercept=mean(occno_disc)),linetype="dashed",col="orange")
```

##Comparison of summary statistics

```{r}
d<-list(Nmixture=occ,Poisson_GLM=occ_nodetect)
d<-melt(d)
colnames(d)<-c("Bird","Plant","value","Iteration","Model")

d %>% group_by(Model,Iteration) %>% summarize(mean=mean(value),sd=sd(value),sum=sum(value)) %>% group_by(Model) %>% summarize(mean_mean=round(mean(mean),2),mean_sd=round(sd(mean),2),mean_sum=round(mean(sum),2))
```

```{r}
save.image("AbundanceSimulation.RData")
```

### 2016-11-14 index.Rmd
website skeleton

---
title: "Inferring ecological interactions"
---

#Aims

To compare methods for inferring species interactions

#Where & When

Sherbrooke, Quebec

Nov 30th - Dec 2nd, 2016

#Workshop Participants

[Dominique Gravel](), [Tim Poisot](), [Kevin Cazelles,]() [Ben Weinstein](benweinstein.weebly.com), [Ignacio Bartomeus](),Ignacio Morales Castilla, David Beauchesne, Mickaël Hedde, Mathilde Besson, Steve Vissault, Guillaume Blanchet

# Access

This website and associated analysis can be cloned

```{}
git clone https://github.com/bw4sz/Montreal_Workshop.git
```

### 2016-11-30 SimulatedData.Rmd
revising simulation function

@@
 ---
-title: Hierarchical Nmixture models for species interactions - Simulations
-author: "Ben Weinstein - Stony Brook University"
+title: "Trait-matching simulation" 
+author: "Ben Weinstein"
 output:
@@
     theme: spacelab
-    keep_md: true
-  word_document: default
 ---
@@
 library(reshape2)
-library(foreach)
-library(chron)
+library(pez)
 library(ggplot2)
 library(knitr)
-library(R2jags)
 library(dplyr)
 library(stringr)
-library(gridExtra)
-library(boot)
 opts_chunk$set(message=FALSE,warning=FALSE,fig.width=5,fig.height=4,echo=TRUE,cache=F,fig.align='center',fig.path="figure/")
-
-set.seed(3)
-#source functions
-source("Bayesian/BayesFunctions.R")
-
-```
-
-```{r,echo=F,cache=FALSE}
-paste("Run Completed at",Sys.time())
 ```
@@
 
-## Parameters
+What are the data we want to simulate?
 
-* 10 hummingbird species
-* 10 plant species
-* Range of hummingbird bill sizes (in mm) ~ Pois(10)/10
-* Range of corolla sizes (in mm) ~ Pois(15)/10
-* Mean frequeny ($\lambda$) for each hummingbird is drawn from U(0,10)  
-* Trait matching (minimizing Bill-Corolla difference) is drawn from a hierarcichal distribution
-$$log(\lambda)<-\alpha_i + \beta_i *traitmatch$$
-$$\alpha=N(3,0.2)$$
-$$\beta1 = N(-1,0.2)$$
-
-* Imperfect detection 
-* $$ p_i = U(0.1,0.9) $$ 
-* 10 camera
-* 3 days per camera
-
-**View simulated strength and form of trait matching **
+size_x = Number of level 1 (eg. Plants)
+size_y = Number of level 2 (eg. Pollinators)
+traits_x = Trait frame for level 1
+traits_y = Trait frame for level 2
+phylogeny_x = phylogeny for level 1
+phylogeny_y = phylogeny for level 2
+Obs = a x by y matrix of visits
 
 ```{r}
-#load("AbundanceSimulation.RData")
-gc()
+#sim_assemblages<-function(x=,y=,traits=
+## Process Model
 ```
 
-#Simulation Parameters
-
 ```{r,fig.height=5,fig.width=8,eval=T}
-#Number of hummingbird species
+#Number of consumer species
 h_species=10
@@
 
-#Compute resource abundance
-
-```{r,eval=T}
-resources<-array(NA,dim=c(h_species,plant_species,cameras))
-
-#fill for each month
-for (x in 1:cameras){
-   resources[,,x]<-rpois(1,10)   
-  }
-resources<-array(data=scale(resources),dim=c(h_species,plant_species,cameras))
-```
-
 #Compute true interaction matrices
@@
 }
-
-#Observed counts in each day
-for(x in 1:h_species){
-  for (y in 1:plant_species){
-    for (z in 1:cameras){
-        for (d in 1:days){
-      #true detection rate of that observed count
-      obs[x,y,z,d]<-rbinom(1,N[x,y,z],p=detection[x])
-      }
-    }
-  }
-}
 ```
@@
 ggplot(mdat,aes(x=traitmatch,y=Interactions,col=as.factor(Bird))) + geom_point() + geom_smooth(aes(group=1),method="glm",method.args = list(family = "poisson")) + labs(col="Bird") + xlab("Absolute value of Bill Length - Corolla Length ")
-
-#Merge resources
-mr<-melt(resources)
-colnames(mr)<-c("Bird","Plant","Camera","Abundance")
-mdat<-merge(mdat,mr,by=c("Bird","Plant","Camera"))
 ```
@@
 ```
-
-# Hierarcichal Nmixture Model
-
-For hummingbird i visiting plant j recorded by camera k on day d:
-
-$$ Y_{i,j,k,d} \sim Binom(N_{i,j,k},\omega_i)$$
-$$N_{i,j,k} \sim Pois(\lambda_{i,j}) $$
-$$log(\lambda_{i,j})<-\alpha_i + \beta_{1,i} * |Bill_i - Corolla_j| * \beta_{2,i} * Abundance_{j,k} + \epsilon$$
-$$ \epsilon \sim Normal(0,\tau_\epsilon)$$
-
-**Priors**
-
-Please recall that jags parameterizes models using precision, not sd (precision = 1/sd^2)
-
-$$\omega_i \sim (\mu_{\omega},\tau_{\omega})$$  $$\mu_{\omega} \sim Normal(0,0.5)   
-$$\tau_{\omega} \sim Uniform(0,10)
-
-$$\alpha_i \sim Normal(\mu_{\alpha},\tau_{\alpha})$$
-$$\beta_{i,1} \sim Normal(\mu_{\beta_1},\tau_{\beta_1})$$
-
-**Hyperpriors**
-$$\mu_{\alpha} \sim Normal(0,0.0001)$$
-$$\mu_{\beta_1} \sim Normal(0,0.0001)$$
-$$\mu_{\beta_2} \sim Normal(0,0.0001)$$
-
-$$\tau_{\alpha} \sim Half-T(0.0001,0.0001)$$
-$$\sigma_{\alpha} = \sqrt[2]{\frac{1}{\tau_\alpha}}$$
-
-$$\tau_{\beta_1} \sim \sqrt[2]{\frac{1}{\sigma_{\beta_1}}}$$
-$$\sigma_{\beta_1} \sim Half-T(0,1)$$
-
-$$\tau_{\beta_2} \sim \sqrt[2]{\frac{1}{\sigma_{\beta_2}}}$$
-$$\sigma_{\beta_2} \sim Half-T(0,1)$$
-
-#Simulated data without detection
-
-```{r,eval=T}
-runs<-30000
-#Source model
-source("Bayesian/NoDetectNmixturePoissonRagged.R")
-
-#print model
-print.noquote(readLines("Bayesian//NoDetectNmixturePoissonRagged.R"))
-
-  #for parallel run
-  Yobs=obs.state$Yobs
-  Bird=obs.state$Bird
-  Plant=obs.state$Plant
-  Plants=max(obs.state$Plant)
-  Cameras=max(obs.state$Camera)
-  Camera=obs.state$Camera
-  Traitmatch=traitmatch
-  Birds=max(obs.state$Bird)
-  Nobs=length(obs.state$Yobs)
-  resources=resources
-
-  #Inits
-  Ninit<-array(dim=c(h_species,plant_species,Cameras),data=1)
-
-  InitStage <- function() {list(epsilon=Ninit)}
-
-  #Parameters to track
-  ParsStage <- c("alpha","beta1","alpha_mu","beta1_sigma","alpha_sigma","ynew","beta1_mu","fit","fitnew","tauE")
-
-  #MCMC options
-  ni <- runs  # number of draws from the posterior
-  nt <- 4   #thinning rate
-  nb <- max(0,runs-2000) # number to discard for burn-in
-  nc <- 2  # number of chains
-
-  Dat<-list("Yobs","Bird","Plant","Plants","Traitmatch","Birds","Nobs","Cameras","Camera","resources","Ninit")
-
-    sim_niave<-do.call(jags.parallel,list(data=Dat,InitStage,parameters.to.save=ParsStage,model.file="Bayesian/NoDetectNmixturePoissonRagged.jags",n.thin=nt, n.iter=ni,n.burnin=nb,n.chains=nc))
-
-```
-
-```{r,eval=F}
-#recompile if needed
-load.module("dic")
-runs<-5000
-recompile(sim_niave)
-sim_niave<-update(sim_niave,n.iter=runs,n.burnin=runs*.95,n.thin=10)
-```
-
-```{r}
-pars_niave<-extract_par(sim_niave,data=obs.state,ynew=T)
-pars_niave$Model<-c("Poisson GLMM")
-```
-
-##Assess Convergence
-
-```{r,cache=FALSE,fig.width=11,fig.height=5}
-ggplot(pars_niave[pars_niave$par %in% c("alpha","beta1"),],aes(x=Draw,y=estimate,col=as.factor(Chain))) + geom_line() + facet_grid(par~species,scale="free") + theme_bw() + labs(col="Chain") + ggtitle("Detection Probability")
-```
-
-```{r,fig.height=5,fig.width=11}
-ggplot(pars_niave[pars_niave$par %in% c("beta1_mu","beta1_sigma","tauE"),],aes(x=Draw,y=estimate,col=as.factor(Chain))) + geom_line() + theme_bw() + labs(col="Chain") + ggtitle("Trait-matching regression") + facet_wrap(~par,scales="free")
-```
-
-##Posteriors
-
-```{r,cache=FALSE,fig.width=7,fig.height=5}
-###Posterior Distributions
-p<-ggplot(pars_niave[pars_niave$par %in% c("alpha","beta1"),],aes(x=estimate)) + geom_histogram() + ggtitle("Estimate of parameters") + facet_grid(species~par,scales="free") + theme_bw() + ggtitle("Species Posteriors")
-
-#Add true values
-tr<-melt(data.frame(species=1:length(detection),alpha=alpha,beta1=beta1),id.var='species')
-colnames(tr)<-c("species","par","value")
-psim<-p + geom_vline(data=tr,aes(xintercept=value),col='red',linetype='dashed',size=1)
-psim
-ggsave("Figures/SimulationPosteriorsNoDetect.jpg",dpi=300,height=8,width=8)
-```
-
-```{r,cache=FALSE,fig.height=3,fig.width=10}
-p<-ggplot(pars_niave[pars_niave$par %in% c("beta1_mu","alpha_mu","alpha_sigma","beta1_sigma","tauE"),],aes(x=estimate)) + geom_histogram() + ggtitle("Hierarchical Posteriors") + facet_grid(~par,scale="free") + theme_bw()
-
-#Add true values
-tr<-melt(list(beta1_mu=beta1_mu,alpha_mu=alpha_mu,alpha_sigma=alpha_sigma,beta1_sigma=beta1_sigma,tauE=tauE))
-
-colnames(tr)<-c("value","par")
-
-psim2<-p + geom_vline(data=tr,aes(xintercept=value),linetype='dashed',size=1,col="red")
-```
-
-**True values are given in the dashed lines.**
-
-##Predicted Relationship
-
-```{r,fig.height=4,fig.width=4}
-castdf<-group_by(pars_niave,Chain) %>% select(par,estimate) %>% filter(par %in% c("beta1_mu","alpha_mu"))
-
-castdf<-dcast(pars_niave[pars_niave$par %in% c("beta1_mu","alpha_mu"),], Chain + Draw~par,value.var="estimate")
-
-#calculated predicted y
-predyniave_trait<-trajF(alpha=castdf$alpha_mu,beta1=castdf$beta1_mu,trait=as.numeric(traitmatch),resources=as.numeric(resources[,1,]),tauSigma = 1/sqrt(tauE))
-
-predyniave_both<-trajF(alpha=castdf$alpha_mu,beta1=castdf$beta1_mu,trait=as.numeric(traitmatch),resources=as.numeric(resources[,1,]),tauSigma=1/sqrt(tauE))
-```
-
-# Simulated data with detection
-
-```{r,eval=T}
-runs<-30000
-
-#Source model
-source("Bayesian/NmixturePoissonRagged.R")
-
-#print model
-print.noquote(readLines("Bayesian//NmixturePoissonRagged.R"))
-
-  #for parallel run
-  Yobs=obs.state$Yobs
-  Bird=obs.state$Bird
-  Plant=obs.state$Plant
-  Time=obs.state$Camera
-  Times=max(obs.state$Camera)
-  Traitmatch=traitmatch
-  Birds=max(obs.state$Bird)
-  Plants=max(obs.state$Plant)
-  Nobs=length(obs.state$Yobs)
-  resources=resources
-
-  #A blank Y matrix - all present
-  Ninit<-array(dim=c(h_species,plant_species,Times),data=max(obs.state$Yobs)*1)
-
-  #Inits
-  InitStage <- function() {list(N=Ninit,epsilon=Ninit/10)}
-
-  #Parameters to track
-  ParsStage <- c("detect","tauE","alpha","beta1","alpha_mu","alpha_sigma","beta1_sigma","ynew","beta1_mu","fit","fitnew")
-
-  #MCMC options
-  ni <- runs  # number of draws from the posterior
-  nt <- 8  #thinning rate
-  nb <- max(0,runs-2000) # number to discard for burn-in
-  nc <- 2  # number of chains
-
-  Dat<-list("Yobs","Bird","Plant","Plants","Traitmatch","resources","Birds","Nobs","Ninit","Time","Times")
-
-    system.time(sim_detect<-do.call(jags.parallel,list(Dat,InitStage,ParsStage,model.file="Bayesian/NmixturePoissonRagged.jags",n.thin=nt, n.iter=ni,n.burnin=nb,n.chains=nc)))
-```
-
-```{r,eval=F}
-#recompile if needed
-load.module("dic")
-runs<-10000
-recompile(sim_detect)
-system.time(sim_detect<-update(sim_detect,n.iter=runs,n.burnin=runs*.95,n.thin=10,parameters.to.save=ParsStage))
-```
-
-```{r}
-pars<-extract_par(sim_detect,data=obs.state,ynew=T)
-pars$Model<-"Nmixture"
-```
-
-##Assess Convergence
-
-```{r,cache=FALSE,fig.width=11,fig.height=5}
-ggplot(pars[pars$par %in% c("detect","alpha","beta1"),],aes(x=Draw,y=estimate,col=as.factor(Chain))) + geom_line() + facet_grid(par~species,scale="free") + theme_bw() + labs(col="Chain") + ggtitle("Detection Probability")
-```
-
-```{r,fig.height=5,fig.width=11}
-ggplot(pars[pars$par %in% c("beta1_mu","alpha_sigma","beta1_sigma","tauE"),],aes(x=Draw,y=estimate,col=as.factor(Chain))) + geom_line() + theme_bw() + labs(col="Chain") + ggtitle("Trait-matching regression") + facet_wrap(~par,scales="free")
-```
-
-##Posteriors
-
-```{r,cache=FALSE,fig.width=7,fig.height=5}
-###Posterior Distributions
-p<-ggplot(pars[pars$par %in% c("detect","alpha","beta1"),],aes(x=estimate)) + geom_histogram() + ggtitle("Estimate of parameters") + facet_grid(species~par,scales="free") + theme_bw() + ggtitle("Species Posteriors")
-
-#Add true values
-tr<-melt(data.frame(species=1:length(detection),detect=detection,alpha=alpha,beta1=beta1),id.var='species')
-colnames(tr)<-c("species","par","value")
-psim<-p + geom_vline(data=tr,aes(xintercept=value),col='red',linetype='dashed',size=1)
-#ggsave("Figures/SimulationPosteriors.jpg",dpi=300,height=8,width=8)
-```
-
-```{r,cache=FALSE,fig.height=3,fig.width=10}
-p<-ggplot(pars[pars$par %in% c("beta1_mu","alpha_mu","alpha_sigma","beta1_sigma","tauE"),],aes(x=estimate)) + geom_histogram() + ggtitle("Hierarchical Posteriors") + facet_wrap(~par,scale="free",nrow=2) + theme_bw()
-
-#Add true values
-tr<-melt(list(beta1_mu=beta1_mu,alpha_mu=alpha_mu,alpha_sigma=alpha_sigma,beta1_sigma=beta1_sigma,tauE=tauE))
-
-colnames(tr)<-c("value","par")
-
-psim2<-p + geom_vline(data=tr,aes(xintercept=value),linetype='dashed',size=1,col="black")
-#ggsave("Figures/SimulationH.jpg",dpi=300,height=4,width=10)
-```
-
-**True values are given in the dashed lines.**
-
-##Compare simulation posteriors with and without detection
-
-```{r,cache=FALSE,fig.width=8,fig.height=9}
-#Bind to other dataset
-parsall<-rbind.data.frame(pars[!pars$par %in% "ynew",],pars_niave[!pars_niave$par %in% "ynew",])
-parsall$Model<-as.factor(parsall$Model)
-
-###Posterior Distributions
-p<-ggplot(parsall[parsall$par %in% c("detect","alpha","beta1"),],aes(x=estimate,fill=Model)) + geom_histogram(position="identity") + ggtitle("Estimate of parameters") + facet_grid(species~par,scales="free") + theme_bw()
-
-#Add true values
-tr<-melt(data.frame(species=1:length(detection),detect=detection,alpha=alpha,beta1=beta1),id.var='species')
-colnames(tr)<-c("species","par","value")
-psim<-p + geom_vline(data=tr,aes(xintercept=value),col='black',linetype='dashed',size=1)
-psim
-#ggsave("Figures/SimulationPosteriorsBoth.jpg",dpi=300,height=8,width=8)
-```
-
-```{r,cache=FALSE,fig.height=3,fig.width=10}
-p<-ggplot(parsall[parsall$par %in% c("beta1_mu","alpha_mu","alpha_sigma","beta1_sigma","tauE_mu","tauE_tau"),],aes(x=estimate,fill=Model)) + geom_histogram(position="identity") + ggtitle("Hierarchical Posteriors") + facet_wrap(~par,scale="free",nrow=2) + theme_bw()
-
-#Add true values
-tr<-melt(list(beta1_mu=beta1_mu,alpha_mu=alpha_mu,alpha_sigma=alpha_sigma,beta1_sigma=beta1_sigma,tauE=tauE))
-
-colnames(tr)<-c("value","par")
-
-psim2<-p + geom_vline(data=tr,aes(xintercept=value),linetype='dashed',size=1,col="black")
-psim2
-#ggsave("Figures/SimulationHBoth.jpg",dpi=300,height=4,width=10)
-```
-
-##Strip plots
-Plot the posterior mean density
-```{r,fig.width=15,fig.height=13}
-spars<-parsall %>% filter(par %in% c("alpha","beta1","detect")) %>% group_by(Model,species,par) %>% summarize(mean=mean(estimate),lower=quantile(estimate,0.05),upper=quantile(estimate,0.95))
-
-tr<-melt(data.frame(species=1:length(detection),detect=detection,alpha=alpha,beta1=beta1,tauE=tauE),id.var='species')
-colnames(tr)<-c("species","par","value")
-
-ggplot(spars,aes(x=Model,ymin=lower,ymax=upper,y=mean,col=Model)) + geom_linerange(size=1.3) + facet_grid(par~species,scales="free") + geom_hline(data=tr,aes(yintercept=value),linetype='dashed',size=1,col="black") + geom_point(aes(y=mean),col='grey50',size=2) + theme_bw() + ylab("Estimate")
-
-#Hierarchical posteriors
-hpars<-parsall %>% filter(par %in% c("beta1_mu","alpha_mu","alpha_sigma","beta1_sigma","tauE")) %>% group_by(Model,species,par) %>% summarize(mean=mean(estimate),lower=quantile(estimate,0.05),upper=quantile(estimate,0.95))
-
-tr<-melt(data.frame(species=1:length(detection),beta1_mu=beta1_mu,alpha_mu=alpha_mu,beta1_sigma=beta1_sigma,alpha_sigma=alpha_sigma,tauE=tauE),id.var='species')
-colnames(tr)<-c("species","par","value")
-
-ggplot(hpars,aes(x=Model,ymin=lower,ymax=upper,y=mean,col=Model)) + geom_linerange(size=1.3) + facet_wrap(~par,scales="free",nrow=1) + geom_hline(data=tr,aes(yintercept=value),linetype='dashed',size=1,col="black") + geom_point(aes(y=mean),col='grey50',size=2) + theme_bw() + ylab("Estimate")
-#ggsave("Figures/StripPlots.svg",height=4,width=9)
-```
-
-##Correlation in posteriors for Nmixture Model
-
-```{r}
-castdf<- pars %>% filter(Model =="Nmixture") %>% group_by(Chain) %>% select(par,estimate,Draw) %>% filter(par %in% c("beta1_mu","alpha_mu")) %>% dcast(Chain+Draw~par,value.var="estimate")
-
-pairs(castdf[,3:4],main="Correlation in Group-Level Posteriors")
-
-castdf<- pars %>% filter(Model =="Nmixture") %>% group_by(Chain) %>% select(par,estimate,Draw,species) %>% filter(par %in% c("alpha","beta1","detect","tauE")) %>% dcast(species+Chain+Draw~par,value.var="estimate")
-
-pairs(castdf[,4:7],main="Correlation in Species-Level Posteriors")
-```
-
-##Predicted Relationship
-
-```{r,fig.height=4,fig.width=4}
-
-castdf<-dcast(pars[pars$par %in% c("beta1_mu","alpha_mu","tauE"),], Chain + Draw~par,value.var="estimate")
-```
-
-## Calculated predicted visitation rates
-
-### Traits
-
-```{r}
-predy_trait<-trajF(alpha=castdf$alpha_mu,beta1=castdf$beta1_mu,tauSigma =1/sqrt(castdf$tauE),trait=as.numeric(traitmatch),resources=as.numeric(resources))
-orig<-trajF(alpha=rnorm(1000,alpha_mu,alpha_sigma),beta1=rnorm(1000,beta1_mu,beta1_sigma),trait=as.numeric(traitmatch),resources=as.numeric(resources),tauSigma =1/sqrt(tauE))
-
-pm<-melt(list(Nmixture=predy_trait,'Poisson GLMM'=predyniave_trait),id.vars=colnames(predy_trait))
-
-tplot<-ggplot(data=pm[,],aes(x=trait)) + geom_ribbon(aes(ymin=lower,ymax=upper,fill=L1),alpha=0.5)  + geom_line(aes(y=mean,col=L1),size=.4,linetype="dashed") + theme_bw() + ylab("Daily Interactions") + xlab("Difference between Bill and Corolla Length") + geom_point(data=mdat,aes(x=traitmatch,y=Interactions),size=.5,alpha=.5) + labs(fill="Model",col="Model") + ggtitle("Simulation: Traits") + geom_line(data=orig,aes(y=mean),col='black',size=.8) + scale_fill_manual(values=c("grey70","Black"))+ scale_color_manual(values=c("grey70","Black"))
-tplot
-ggsave("Figures/SimPredictBoth.jpg",height=5,width=7)
-```
-
-Black line is the true relationship.
-The red line is the posterior mean with confidible intervals in shaded grey for the proposed bayesian model.
-The blue line is the same model, but assuming perfect detection rates.
-
-**Conclusion:** Accounting for detection and non-independence greatly increases the accuracy of the predicted state.
-The perfect detection model underestimates the strength of trait matching among hummingbirds and their foodplants.
-
-##Posterior Predictive Check
-
-Since I have simualted the data, it should fit as well as any random dataset drawn from the estimated parameters. An ideal fit would be posterior values sitting along the 1:1 line.
-
-```{r,fig.height=4,fig.width=8}
-fitstat<-droplevels(parsall[parsall$par %in% c("fit","fitnew"),])
-fitstat<-dcast(fitstat,Draw+Chain+Model~par,value.var="estimate")
-
-#add 1:1 line
-ymin<-round(min(c(fitstat$fit,fitstat$fitnew)))
-ymax<-round(max(c(fitstat$fit,fitstat$fitnew)))
-ab<-data.frame(x=ymin:ymax,y=ymin:ymax)
-p<-ggplot(fitstat,aes(x=fit,y=fitnew)) + geom_point(aes(col=Model)) + theme_bw() + coord_equal()
-psim4<-p  + labs(x="Discrepancy of observed data",y="Discrepancy of replicated data",col="Model") + geom_line(data=ab,aes(x=x,y=y)) + ggtitle("Simulated Data")
-psim4
-ggsave("Figures/SimulationDisc.jpeg",height=5,width=5)
-```
-
-```{r,echo=FALSE}
-#Write Figures as panels
-jpeg("Figures/SimulationPosteriorsBoth.jpg",height=12,width=12,units="in",res=300)
-grid.arrange(psim,psim2,ncol=1,heights=c(3/4,1/4))
-dev.off()
-
-```
-
-#By species
-
-#Compare using true known interactions
-
-## No Detection Nmixture Model
-
-```{r}
-true_state<-obs.state %>% group_by(Bird,Plant) %>% summarize(n=sum(Yobs)) %>% acast(.$Bird~.$Plant)
-
-N_niave<-pars_niave %>% filter(par== "ynew")
-
-#Discrepancy function
-#define discrep function
-chisq<-function(o,e){(o-e)^2/(e+0.5)}
-
-bydraw<-split(N_niave,list(N_niave$Chain,N_niave$Draw))
-
-#Don't need giant matrix sitting around
-rm(N_niave)
-gc()
-
-occ_nodetect_matrix<-lapply(bydraw,function(x){
-  r<-acast(x,species ~ plant,value.var = "estimate",fun.aggregate = sum)
-})
-
-rm(bydraw)
-gc()
-
-#calculate discrep on those deviates
-occ_nodetect<-lapply(occ_nodetect_matrix,function(r){
-    #for each position what is the chisq
-  rmerge<-matrix(nrow = nrow(true_state),ncol=ncol(true_state))
-  for (x in 1:nrow(r)){
-    for (y in 1:ncol(r)){
-     rmerge[x,y]<-chisq(e=r[x,y],o=true_state[x,y])
-      }
-    }
-  return(rmerge)
-})
-
-names(occ_nodetect)<-1:length(occ_nodetect)
-names(occ_nodetect_matrix)<-1:length(occ_nodetect_matrix)
-```
-
-##With Detection
-
-```{r}
-Ndetect<-pars %>% filter(par== "ynew")
-bydraw<-split(Ndetect,list(Ndetect$Chain,Ndetect$Draw))
-
-#Don't need giant matrix around
-rm(Ndetect)
-gc()
-
-occ_matrix<-lapply(bydraw,function(x){
-  r<-acast(x,species ~ plant,value.var = "estimate",fun.aggregate = sum)
-  })
-
-rm(bydraw)
-gc()
-
-#calculate discrep for those aggregated matrices
-occ<-lapply(occ_matrix,function(r){
-    #for each position what is the chisq
-  rmerge<-matrix(nrow = nrow(true_state),ncol=ncol(true_state))
-  for (x in 1:nrow(r)){
-    for (y in 1:ncol(r)){
-     rmerge[x,y]<-chisq(e=r[x,y],o=true_state[x,y])
-      }
-    }
-  return(rmerge)
-})
-
-names(occ)<-1:length(occ)
-names(occ_matrix)<-1:length(occ_matrix)
-```
-
-##Compare to observed data
-
-```{r,fig.width=11}
-#true number of observed interactions
-
-mmat<-melt(true_state)
-colnames(mmat)<-c("Bird","Plant","True_State")
-
-#append to predicted matrices
-
-#Nmixture with detection
-mocc<-melt(occ_matrix)
-colnames(mocc)<-c("Bird","Plant","Nmixture","Iteration")
-simdat<-merge(mocc,mmat,by=c("Bird","Plant"),all.x=T)
-
-#Nmixture with nodetection
-moccd<-melt(occ_nodetect_matrix)
-colnames(moccd)<-c("Bird","Plant","Poisson GLMM","Iteration")
-
-simdat<-merge(simdat,moccd,by=c("Bird","Plant","Iteration"))
-
-simdat<-melt(simdat,measure.vars = c("Nmixture","Poisson GLMM"))
-
-ggplot(simdat,aes(x=True_State,y=value,col=variable),alpha=1) + geom_point() + geom_abline() + labs(col="Model") + ylab("Predicted State") + xlab("True State") + theme_bw() + facet_wrap(~variable)
-
-ggsave("Figures/PredictedState.jpeg",height=3,width=8)
-
-ggplot(simdat[simdat$variable %in% c("Nmixture","Poisson GLMM"),],aes(x=True_State,y=value,col=variable)) + geom_point(alpha=.3) + geom_abline() + labs(col="Model") + ylab("Predicted State") + xlab("True State") + theme_bw()
-ggsave("Figures/PredictedState_noM.jpeg",height=3,width=8)
-
-#difference in the middle
-simd<-dcast(simdat,...~variable)
-simd$Diff<-simd$Nmixture-simd$`Poisson GLMM`
-ggplot(simd,aes(x=True_State,y=abs(Diff))) + geom_point() + ylab("|Nmixture - Poisson GLMM|") + theme_bw() + labs(x="True State")
-ggsave("Figures/Difference_Pred.jpeg",height=4,width=6)
-```
-
-## View predicted trait-matching relationship with the number of visits.
-```{r}
-simdat<-merge(simdat,traitmelt,c("Bird","Plant"))
-mmat<-merge(mmat,traitmelt,by=c("Bird","Plant"))
-```
-
-#Predicted total number of visits based on morphology
-
-```{r,fig.height=5,fig.width=5}
-simT<-simdat %>% group_by(variable,traitmatch) %>% summarize(Lower=quantile(value,0.05),Upper=quantile(value,0.95),y=mean(value))
-
-ggplot(simT,aes(x=traitmatch)) + geom_ribbon(aes(ymin=Lower,ymax=Upper,fill=variable),alpha=0.4) + geom_line(aes(y=y,col=variable),linetype='dashed') + theme_bw()  + geom_point(data=mmat,aes(y=True_State)) + labs(x="Difference in Bill and Corolla Length",y="Total Predicted Visits",fill="Model",col='Model')
-
-```
-
-View a couple example data points from across the type of interactions.
-
-```{r}
-h<-simdat[which.max(simdat$True_State),c("Bird","Plant")]
-d<-simdat[simdat$Bird %in% h$Bird & simdat$Plant %in% h$Plant,]
-
-ggplot(data=d,aes(x=value,fill=variable))+ geom_histogram(position="identity") + labs(fill="Model") + geom_vline(aes(xintercept=True_State)) + ggtitle("High Visitation Example")
-
-h<-simdat[which.min(simdat$True_State),c("Bird","Plant")]
-d<-simdat[simdat$Bird %in% h$Bird & simdat$Plant %in% h$Plant,]
-
-ggplot(data=d,aes(x=value,fill=variable))+ geom_histogram(position="identity") + labs(fill="Model") + geom_vline(aes(xintercept=True_State)) + ggtitle("Low Visitation Example")
-```
-
-##Summary of discrepancy of predicted matrices
-
-```{r}
-#Nmixture without detection
-occno_disc<-sapply(occ_nodetect,function(x) median(x))
-
-#Nmixture with detection
-occ_disc<-sapply(occ,function(x) median(x))
-
-#compared to bayesian
-ggplot() + xlab("Chi-squared Discrepancy") + geom_histogram(data=data.frame(occ_disc),aes(x=occ_disc),fill="red",alpha=.6) + theme_bw() +geom_vline(aes(xintercept=mean(occ_disc)),linetype="dashed",col="red") + geom_histogram(data=data.frame(occno_disc),aes(x=occno_disc),fill="orange",alpha=.6) + geom_vline(aes(xintercept=mean(occno_disc)),linetype="dashed",col="orange")
-```
-
-##Comparison of summary statistics
-
-```{r}
-d<-list(Nmixture=occ,Poisson_GLM=occ_nodetect)
-d<-melt(d)
-colnames(d)<-c("Bird","Plant","value","Iteration","Model")
-
-d %>% group_by(Model,Iteration) %>% summarize(mean=mean(value),sd=sd(value),sum=sum(value)) %>% group_by(Model) %>% summarize(mean_mean=round(mean(mean),2),mean_sd=round(sd(mean),2),mean_sum=round(mean(sum),2))
-```
-
-```{r}
-save.image("AbundanceSimulation.RData")
-```

### 2016-11-30 index.Rmd
add a makefile for R users that do not use Rstudio

@@
 
-[Dominique Gravel](), [Tim Poisot](), [Kevin Cazelles,]() [Ben Weinstein](benweinstein.weebly.com), [Ignacio Bartomeus](),Ignacio Morales Castilla, David Beauchesne, Mickaël Hedde, Mathilde Besson, Steve Vissault, Guillaume Blanchet
+[Dominique Gravel](), [Tim Poisot](), [Kevin Cazelles,](http://kevincazelles.fr/) [Ben Weinstein](benweinstein.weebly.com), [Ignacio Bartomeus](),Ignacio Morales Castilla, David Beauchesne, Mickaël Hedde, Mathilde Besson, Steve Vissault, Guillaume Blanchet
 

### 2016-11-30 index.Rmd
add spreadsheet and google docs links

--- 2016-11-14 index.Rmd
@@
 
-[Dominique Gravel](), [Tim Poisot](), [Kevin Cazelles,]() [Ben Weinstein](benweinstein.weebly.com), [Ignacio Bartomeus](),Ignacio Morales Castilla, David Beauchesne, Mickaël Hedde, Mathilde Besson, Steve Vissault, Guillaume Blanchet
+[Dominique Gravel](), [Tim Poisot](), [Kevin Cazelles,](http://kevincazelles.fr/) [Ben Weinstein](benweinstein.weebly.com), [Ignacio Bartomeus](),Ignacio Morales Castilla, David Beauchesne, Mickaël Hedde, Mathilde Besson, [Steve Vissault](http://steveviss.github.io/), Guillaume Blanchet
 
 # Access
+
+### Website
 
@@
 ```
+
+### Google doc:
+
+[Here](https://docs.google.com/document/d/1x1YHzF-vZXbrQUIF97TNqyIJ6bXyaEiFjMsi18nJDag)
+
+### Spreadsheet :
+
+[Here](https://docs.google.com/spreadsheets/d/1Nps6BsDrkWMbvLa_lsgc5MT8RBtTeFX6hWlUjsRlkO0/edit?ts=583f2190#gid=1879087403)

### 2016-12-01 SimulatedData.Rmd
trait-matching knitted

@@
 library(reshape2)
-library(pez)
+library(bipartite)
 library(ggplot2)
@@
 library(stringr)
-opts_chunk$set(message=FALSE,warning=FALSE,fig.width=5,fig.height=4,echo=TRUE,cache=F,fig.align='center',fig.path="figure/")
+library(R2jags)
+opts_chunk$set(message=FALSE,warning=FALSE,fig.width=5,fig.height=5,echo=TRUE,cache=F,fig.align='center',fig.path="figure/")
 ```
@@
 
-size_x = Number of level 1 (eg. Plants)
-size_y = Number of level 2 (eg. Pollinators)
+size_x = Number of level 1 
+size_y = Number of level 2
 traits_x = Trait frame for level 1
 traits_y = Trait frame for level 2
-phylogeny_x = phylogeny for level 1
-phylogeny_y = phylogeny for level 2
-Obs = a x by y matrix of visits
+Obs = x by y matrix of visits
 
 ```{r}
-#sim_assemblages<-function(x=,y=,traits=
-## Process Model
-```
-
-```{r,fig.height=5,fig.width=8,eval=T}
-#Number of consumer species
-h_species=10
-plant_species=10
-cameras<-5
-days<-3
-
-#Bill sizes
-Bill<-rpois(h_species,10)
-
-#Corolla sizes
-Corolla<-rpois(plant_species,15)
+simA<-function(size_x,size_y,traits_x,traits_y,beta1_sigma= 0.4,alpha_sigma= 0){
 
   #Subtract both and take absolute value, convert cm
-traitmatch<-abs(sapply(Corolla,function(x) x - Bill)/10)
+  traitmatch<-abs(sapply(traits_y,function(x) x - traits_x))
 
@@
 
-#abundance
-
   #species variance in slopes
-beta1_sigma<- 0.2
+  beta1_sigma<- 0.1
 
@@
   alpha_mu<- 3
-alpha_sigma<- 0.2
-
-#Poisson overdipersion
-tauE<-10
+  alpha_sigma<- 0
 
   #species level
-detection= inv.logit(rnorm(h_species,0,1.67))
-beta1<-rnorm(h_species,beta1_mu,beta1_sigma)
-alpha<-rnorm(h_species,alpha_mu,alpha_sigma)
-```
+  beta1<-rnorm(size_x,beta1_mu,beta1_sigma)
+  alpha<-rnorm(size_x,alpha_mu,alpha_sigma)
 
-#Compute true interaction matrices
-
-```{r,eval=T}
   #for each species loop through and create a replicate dataframe
-obs<-array(dim=c(h_species,plant_species,cameras,days))
-lambda<-array(dim=c(h_species,plant_species,cameras))
-N<-array(dim=c(h_species,plant_species,cameras))
+  obs<-array(dim=c(size_x,size_y))
+  N<-array(dim=c(size_x,size_y))
 
   #create intensities
-for(x in 1:h_species){
-  for (y in 1:plant_species){
-    for (z in 1:cameras){
-      lambda[x,y,z]<-exp(alpha[x] + beta1[x] * traitmatch[x,y] + rnorm(1,0,1/sqrt(tauE)))
-  }
+  for(x in 1:size_x){
+    for (y in 1:size_y){
+
+        #intensity
+        N[x,y]<-exp(alpha[x] + beta1[x] * traitmatch[x,y])
+
+        #draw one state
+        obs[x,y]<-rpois(1,N[x,y])
     }
@@
 
-#draw latent states
-for(x in 1:h_species){
-  for (y in 1:plant_species){
-    for (z in 1:cameras){
-      # true latent count
-      N[x,y,z]<-rpois(1,lambda[x,y,z])
-    }
-  }
+#draw intensity
+
+tx<-data.frame(I=1:length(traits_x),TraitI=traits_x)
+ty<-data.frame(J=1:length(traits_y),TraitJ=traits_y)
+
+#view trait matching
+dat<-melt(obs)
+colnames(dat)<-c("I","J","Interactions")
+
+dat<-merge(dat,tx)
+dat<-merge(dat,ty)
+
+#define data types, make species
+dat$I<-letters[dat$I]
+dat$J<-letters[dat$J]
+
+return(dat)
 }
@@
 
-##View correlation in simulated latent state
+```{r}
+size_x = 5
+size_y = 19
+
+#Bill sizes
+traits_x<-rpois(size_x,10)
+
+#Corolla sizes
+traits_y<-rpois(size_y,10)
+
+dat<-simA(size_x=size_x,size_y=size_y,traits_x=traits_x,traits_y=traits_y)
+
+#view heatmap
+heatmap<-function(dat){
+  ggplot(data=dat,aes(x=I,y=J,fill=Interactions)) + geom_tile() +    scale_fill_continuous(low="white",high="red")
+}
+
+heatmap(dat)
+
+traitplot<-function(dat){
+    ggplot(data=dat,aes(y=Interactions,x=abs(TraitI-TraitJ))) + geom_point() + geom_smooth(aes(group=1),method="glm",method.args=list(family="poisson"))
+}
+
+traitplot(dat)
+```
+
+# Poisson Regression
+
+```{r,eval=T}
+#format traitmatch as matrix
+dat$Traitmatch<-abs(dat$TraitI-dat$TraitJ)
+Traitmatch<-acast(data=dat,I~J,value.var="Traitmatch")
+
+runs<-100000
+
+#Source model
+source("Bayesian/Poisson.R")
+
+#print model
+print.noquote(readLines("Bayesian//Poisson.R"))
+
+  #for parallel run
+  Yobs=dat$Interactions
+  Bird=as.numeric(as.factor(dat$I))
+  Plant=as.numeric(as.factor(dat$J))
+  Birds=length(unique(dat$I))
+  Traitmatch=Traitmatch
+  Plants=length(unique(dat$J))
+  Nobs<-length(Yobs)
+
+  #Parameters to track
+  ParsStage <- c("beta1","ynew","fit","fitnew")
+
+  #MCMC options
+  ni <- runs  # number of draws from the posterior
+  nt <- 1  #thinning rate
+  nb <- max(0,runs-500) # number to discard for burn-in
+  nc <- 2  # number of chains
+
+  modelDat<-list("Yobs","Bird","Plant","Plants","Traitmatch","Birds","Nobs")
+
+    m1<-do.call(jags.parallel,list(data=modelDat,parameters.to.save=ParsStage,model.file="Bayesian/Poisson.jags",n.thin=nt, n.iter=ni,n.burnin=nb,n.chains=nc,DIC=F))
+```
 
 ```{r}
-mdat<-melt(N)
-colnames(mdat)<-c("Bird","Plant","Camera","Interactions")
+parsO<-melt(m1$BUGSoutput$sims.array)
+colnames(parsO)<-c("Draw","Chain","parameter","estimate")
 
-traitmelt<-melt(traitmatch)
-colnames(traitmelt)<-c("Bird","Plant","traitmatch")
+#label species and plants
+l<-levels(parsO$parameter)
 
-mdat<-merge(mdat,traitmelt,c("Bird","Plant"))
-ggplot(mdat,aes(x=traitmatch,y=Interactions,col=as.factor(Bird))) + geom_point() + geom_smooth(aes(group=1),method="glm",method.args = list(family = "poisson")) + labs(col="Bird") + xlab("Absolute value of Bill Length - Corolla Length ")
+#parameters to save
+totrack<-m1$parameters.to.save
+
+#assign species index to ragged frame.
+sp_pl<-data.frame(parameter=l,Index=as.numeric(str_match(l,pattern="\\[(\\d+)]")[,2]),par=str_extract(l,"\\w+"))
+
+#merge levels
+pars<-merge(parsO,sp_pl)
+
+pars<-pars[!pars$par %in% "deviance",]
+
 ```
 
-##View Detection Rates
+```{r,fig.width=15}
+#view chaings
+ggplot(pars[pars$par %in% c("beta1"),],aes(x=Draw,col=as.factor(Chain),y=estimate))  + geom_line() + facet_wrap(~Index,scales="free") 
+```
+
+##Posterior predictive check
+
+```{r,fig.height=4,fig.width=8}
+fitstat<-droplevels(pars[pars$par %in% c("fit","fitnew"),])
+fitstat<-dcast(fitstat,Draw+Chain~par,value.var="estimate")
+
+#add 1:1 line
+ymin<-round(min(c(fitstat$fit,fitstat$fitnew)))
+ymax<-round(max(c(fitstat$fit,fitstat$fitnew)))
+ab<-data.frame(x=ymin:ymax,y=ymin:ymax)
+p<-ggplot(fitstat,aes(x=fit,y=fitnew)) + geom_point() + theme_bw() + coord_equal()
+psim4<-p  + labs(x="Discrepancy of observed data",y="Discrepancy of replicated data") + geom_line(data=ab,aes(x=x,y=y)) + ggtitle("Simulated Data")
+psim4
+
+```
+
+#View simulated and predicted traitmatching
 
 ```{r}
-obs.state<-melt(obs)
-colnames(obs.state)<-c("Bird","Plant","Camera","Day","Yobs")
-obs.state<-merge(mdat,obs.state,by=c("Bird","Plant","Camera"))
-ggplot(obs.state,aes(x=Interactions,y=Yobs,col=Camera)) + geom_point() + theme_bw() + geom_abline() + coord_equal()
+
+#merge the dataframes.
+dat$Index<-1:nrow(dat)
+
+#posterior distribution for each index
+ypred<-pars %>% filter(par=="ynew") %>% group_by(Index) %>% summarize(mean=mean(estimate),lower=quantile(estimate,0.05),higher=quantile(estimate,0.95))
+
+datpred<-merge(dat,ypred)
+
+ggplot(datpred) + geom_point(aes(x=Traitmatch,y=Interactions),col="black") + geom_ribbon(aes(x=Traitmatch,ymin=higher,ymax=lower),alpha=0.6) + theme_bw()
+
+ggplot(datpred) + geom_point(aes(x=Traitmatch,y=Interactions),col="black") + geom_ribbon(aes(x=Traitmatch,ymin=higher,ymax=lower),alpha=0.6) + theme_bw() + facet_wrap(~I)
 ```
+
+# Compute emergent statistics
+
+```{r}
+
+makeN<-function(x){
+  predweb<-acast(data=x,I~J,value.var="estimate")
+
+  #calculate network statistic
+  nstat<-networklevel(predweb,index=c("weighted connectance","weighted NODF"))
+  data.frame(Metric=names(nstat),nstat)
+}
+
+#calculate network stats for each model
+nstats<-pars %>% filter(par=="ynew") %>% group_by(Draw,Chain) %>% inner_join(dat) %>% do(makeN(.))
+
+#calculate true network stat
+tweb<-acast(dat,I~J,value.var="Interactions")
+tstat<-networklevel(tweb,index=c("weighted connectance","weighted NODF"))
+tstat<-data.frame(Metric=names(tstat),tstat)
+
+ggplot(nstats,aes(x=nstat)) + geom_density(alpha=0.6,fill="black") + facet_wrap(~Metric,scales='free',nrow=2) + theme_bw()  + geom_vline(col="red",linetype="dashed",data=tstat,aes(xintercept=tstat)) + labs(x="value")
+```

### 2016-12-01 Bayesian/Poisson.R

@@ -1,0 +1,68 @@
+
+sink("Bayesian/Poisson.jags")
+
+cat("
+    model {
+
+    #Compute intensity for each pair of birds and plants
+    for (i in 1:Birds){
+    for (j in 1:Plants){
+    
+    #Process Model with log normal overdispersion
+    log(lambda[i,j])<-alpha[i] + beta1[i] * Traitmatch[i,j] 
+    }
+    }
+    
+    
+    #Prediction
+    for (x in 1:Nobs){
+    
+    # Observed State
+    Yobs[x] ~ dpois(lambda[Bird[x],Plant[x]]) 
+    
+    #Assess Model Fit
+    
+    #Fit discrepancy statistics
+    eval[x]<-lambda[Bird[x],Plant[x]]
+    E[x]<-pow((Yobs[x]-eval[x]),2)/(eval[x]+0.5)
+    
+    ynew[x]~dpois(lambda[Bird[x],Plant[x]])
+    E.new[x]<-pow((ynew[x]-eval[x]),2)/(eval[x]+0.5)
+    
+    }
+    
+    ###Priors
+    
+    #Process Model
+    
+    #Species level priors
+    for (i in 1:Birds){
+    
+    #Intercept
+    alpha[i] ~ dnorm(alpha_mu,alpha_tau)
+    
+    #Traits slope 
+    beta1[i] ~ dnorm(beta1_mu,beta1_tau)    
+    }
+    
+    #Group process priors
+    
+    #Intercept 
+    alpha_mu ~ dnorm(0,0.001)
+    alpha_tau ~ dunif(0,10)
+    alpha_sigma<-pow(1/alpha_tau,0.5) 
+    
+    #Trait
+    beta1_mu~dnorm(0,0.001)
+    beta1_tau ~ dunif(0,10)
+    beta1_sigma<-pow(1/beta1_tau,0.5)
+    
+    #derived posterior check
+    fit<-sum(E[]) #Discrepancy for the observed data
+    fitnew<-sum(E.new[])
+    
+    
+    }
+    ",fill=TRUE)
+
+sink()

### 2016-12-01 Data/data_explnation.Rmd
Add data (not completelly ready, but usefl)

---
title: "Data description"
author: "I. Bartomeus"
date: "12/1/2016"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

__Note:__ This data needs more curation and can be used to play with methods, but needs taxonomic unification, better traits, etc... before looking at results.

Data collected in 2015 across 16 sites. Each site was visited 7 times. Each time we recorded three data types:

- All interactions in 100 m transects
- Focal interactions to selected plants
- Oportunistic new interactions not detected by the last two methods (`out`)

there are three datasets so far:

__Interactions:__ This include `transect` and `visitation` data. Data can be used pooled (1445 interactions (5643 visits) general question abouts trait matching) or splited by site (`Site_ID`) and `Round`. Pollinator traits in this table are measured per individual.
```{r}
int <- read.csv("interactions.csv", h = T)
head(int)
nrow(int)
sum(int$Frequency, na.rm = TRUE)
```
Challenges:  
- long tail of singletones  
- not all species are taxonomically resolved. What to do with specimens we ony know Genus?  
- Low sample size when you split per site and round

__pollinator traits:__
```{r}
pol_tr <- read.csv("traits_pollinators.csv", h = T)
head(pol_tr)
```
Challenges:  
- We are not sure which trait are driving interactions (quite generalized system)  
- Need to inegrate traits across different orders

__plant traits:__ Not ready yet

__independent plant flower abundance__: Ca be combined by Site_ID and Round
```{r}
flw <- read.csv("flowers.csv", h = T)
head(flw)
```

__Phylogenies__ work in progress.  
- Plants: Phylomatic  
- Pollinators: Available at genus level only

### 2016-12-01 Analysis/niche.bin.R
addition of functions for binary niche model

@@ -1,0 +1,47 @@
+
+# This function use a pair of traits 
+# By convention, i is the target layer and j is the opposed layer
+
+# Data: a data frame with Ti as traits of the layer i and Tj traits of the layer j
+# each row correspond to a pair of species with known interaction
+niche.bin = function(L, data, thresh) {
+
+	# Fit model
+	library(quantreg)
+	model_o = lm(Tj ~ Ti, data)
+	model_lo = rq(Tj ~ Ti, tau = thresh, data)
+	model_up = rq(Tj ~ Ti, tau = 1 - thresh, data)
+
+	# Return results
+	return(list(model_o = model_o, model_lo = model_lo, model_up = model_up))
+	 
+}
+
+# newdata: a dataframe with a data frame with Ti as traits of the layer i and Tj traits of the layer j
+# each row correspond to a pair of trait for which we want to know if there is an 
+# interaction happenning
+predict.niche.bin = function(newdata, models, net_type = "unipartite") {
+	with(newdata, {
+		npairs = length(Ti)
+
+		# Lower boundary
+		lo = models$model_lo[1] + models$model_lo[2]*Ti
+
+		# Upper boundary
+		up = models$model_up[1] + models$model_up[2]*Ti
+
+		# Compute interactions among pairs of species
+		L = numeric(npairs)
+		k = 1
+		for(i in 1:npairs)
+			for(j in 1:npairs) {
+				if(Tj[j] > lo[i] & newTj[j] < up[i]) L[k] = 1
+				k = k+1
+			}
+	})
+	return(list(L = L))
+}
+
+
+
+

### 2016-12-01 Analysis/niche.bin.R
edits to niche.bin

@@ -20,7 +20,9 @@ niche.bin = function(L, data, thresh) {
 # newdata: a dataframe with a data frame with Ti as traits of the layer i and Tj traits of the layer j
 # each row correspond to a pair of trait for which we want to know if there is an 
 # interaction happenning
-predict.niche.bin = function(newdata, models, net_type = "unipartite") {
+predict.niche.bin = function(models, newdata) {
+	# Check if the newdata has the right format
+
 	with(newdata, {
 		npairs = length(Ti)
 
@@ -32,12 +34,8 @@ predict.niche.bin = function(newdata, models, net_type = "unipartite") {
 
 		# Compute interactions among pairs of species
 		L = numeric(npairs)
-		k = 1
-		for(i in 1:npairs)
-			for(j in 1:npairs) {
-				if(Tj[j] > lo[i] & newTj[j] < up[i]) L[k] = 1
-				k = k+1
-			}
+		L[Tj > lo & Tj < up] = 1
+
 	})
 	return(list(L = L))
 }

### 2016-12-01 Analysis/direct.centrality.R
Direct centrality

@@ -1,0 +1,28 @@
+# Direct matching centrality
+# Currently deals with logistic and poisson regression, and random forest
+
+direct.centrality = function(data, algorithm, ...) {
+
+	if(algorithm == "logistic") model = glm(L ~ .*., data, family = "binomial")		
+
+	else if(algorithm == "poisson") model = glm(L ~ .*., data, family = "poisson")
+
+	else if(algorithm == "RF") {
+		library(randomForest)
+		model = randomForest(L ~ .*., data, ...)
+	} 
+
+	return(model)
+}
+
+
+predict.direct.centrality = function(newdata, model, replicates) {
+
+	p = predict(model, newdata, type = "response")
+
+	L = list()
+	for(n in 1:replicates) L[[n]] = rbinom(nrow(newdata), size = 1, prob = p)
+
+	return(list(p = p, L = L))
+
+}

### 2016-12-01 Analysis/niche.bin.R

@@ -1,7 +1,5 @@
-
 # This function use a pair of traits 
 # By convention, i is the target layer and j is the opposed layer
-
 # Data: a data frame with Ti as traits of the layer i and Tj traits of the layer j
 # each row correspond to a pair of species with known interaction
 niche.bin = function(L, data, thresh) {
@@ -39,7 +37,3 @@ predict.niche.bin = function(models, newdata) {
 	})
 	return(list(L = L))
 }
-
-
-
-

### 2016-12-02 ms/intro_drafted.Rmd
First attempt to draft an introduction. Requires much much more work.

Making predictions of All Links In Ecological Networks

Introduction

Biotic interactions form the backbone of biodiversity (Bascompte & Jordano 2007). They pose selective pressures that are key to evolution (ref) and operate together with abiotic factors in constraining the distribution of biodiversity (Soberon et al. 2007). Documenting and analyzing ecological networks is essential to: (i) understanding the rules underlying pairwise species interactions (Bascompte et al. 2006), (ii) quantifying ecosystem functioning and services (Thomson et al 2012), or (iii) predicting changes in from of global change scenarios (ref). In spite of increasing efforts, it is unlikely that we will ever manage to empirically document the massive amount of interactions existing in nature (Morales-Castilla et al. 2015). Because of this, complementing field work with modelling is gaining acceptance among network ecologists (ref). In a context where novel ecosystems arise as a result of species invasions or due to geographic shifts tracking climate change, our ability to predict future interactions will limit how efficiently we anticipate and respond to threats on biodiversity (i.e. species local extinctions) and on humans (i.e. spread of infectious disease or plagues affecting to crops).

Species characteristics or traits are at the core of biological interactions (Bartomeus et al. 2016). This is because traits are the results of species adaptations to exploit abiotic or biotic resources. In fact, traits are rooted in all the principal theoretical mechanism proposed to explain when two species can interact, and hence are the perfect candidates to become the backbone of any framework that aims to predict species interactions. First, the idea of morphological trait matching directly relates to the idea that two species can only interact if there is certain concordance between the traits of both partners. For example, in fishes predator mouth gap  has to be bigger than prey body size in order to get a predatory interaction (Gravel et al. 2013). This idea is formalized in the niche model theory, which in its simpler form postulates that [species optimum is placed along a niche axes?…] (Williams...) and has been subsequently refined to include [more axes, prob] … (Gravel et al. 2013, Bartomeus et al. 2016). Second, species traits also determine the range of environmental conditions on which species can survive and co-occurrence in space and time is precisely the first requisite for two species to interact (Morales-Castilla et al 2015). Hence, patterns emerging through processes governed by island biogeography theory can be also viewed at the light of traits. Third, traits are the result of evolutionary history. Phylogenetic information can constrain trait evolution either directly (affecting upon trait matching) or indirectly (via niche conservationism, for example). Finally, it’s worth noting that traits are not only related to the niche differences among species, but also to the fitness effects gained by both interaction partners. This can potentially establish a link with to modern coexistence theory. In fact, neutral expectations postulating that species abundances are the main drivers of interactions, are not incompatible with the underlying role of species traits shaping species abundances via effects in fitness (Bartomeus et al. 2016).

We have showed that traits are a common currency that link mechanistic understanding of species interactions, but given that different traits are involved in different processes (e.g. traits involved in trait matching are not the same involved in species co-occurrence), one may think using traits for inferring biotic interactions would rapidly become intractable. Luckily, recent research has shown that traits can be organized in ecological syndromes and that less than five trait axis may suffice to describe ecological interactions (Eklöf).

Indeed, on the quest for describing species interactions, traits had played a pivotal role and several methods have recently emerged to test hypotheses of trait matching for pairwise interactions (Stang et al 2006, Delhing et al. 2014, Bartomeus et al 2016) or understand the phylogenetic structure of ecological networks (Rezende et al. 2007). However, generating biological meaningful predictions of species interactions and its emerging properties have proven challenging (Olito and Fox 2014). An ideal method is grounded in ecological theory,  which makes the description of ecological interactions interpretable. This method should be able to not only describe, but also predict ecological interactions based on a set of simple parameters. Importantly, this predictions should be accompanied by an associated uncertainty. In addition, given the difficulty of empirically assessing complete interaction networks, this method should be also robust to the use of incomplete data.

Here, we review the main methods proposed to evaluate plant-pollinator interactions with emphasis on its utility for predicting the position of a new species in the network. We use simulated and real data … Our objectives is to provide a guideline of the strengths and limitations of different methods in relation to the type of question pursued and data availability. There is an urgent need to better understand how communities interact and predict new interactions which will inevitably occur as a result of human induced rapid environmental change. In addition, our understanding of how species interact is limited to the few interactions are directly observable such as pollination of predation, but key species interactions not directly observable remains unlocked. Interactions inferences may be the only option for describing and predicting interactions in soil food webs or bacteria-phage networks.

### 2016-12-07 SimulatedData.Rmd
link to manuscript

@@
 library(knitr)
+library(boot)
+library(tidyr)
 library(dplyr)
@@
 opts_chunk$set(message=FALSE,warning=FALSE,fig.width=5,fig.height=5,echo=TRUE,cache=F,fig.align='center',fig.path="figure/")
+
+#source helper function
+source('Bayesian/Functions.R')
 ```
@@
 
-  #regression slopes
+  #regression slope
   #traits
@@
         #intensity
-        N[x,y]<-exp(alpha[x] + beta1[x] * traitmatch[x,y])
+        N[x,y]<-inv.logit(alpha[x] + beta1[x] * traitmatch[x,y])
 
         #draw one state
-        obs[x,y]<-rpois(1,N[x,y])
+        obs[x,y]<-rbinom(1,1,N[x,y])
     }
@@
 traitplot<-function(dat){
-    ggplot(data=dat,aes(y=Interactions,x=abs(TraitI-TraitJ))) + geom_point() + geom_smooth(aes(group=1),method="glm",method.args=list(family="poisson"))
+    ggplot(data=dat,aes(y=Interactions,x=abs(TraitI-TraitJ))) + geom_point() + geom_smooth(aes(group=1),method="glm",method.args=list(family="binomial")) + labs(x="Difference in traits",y="Probability of Interaction")
 }
 
-traitplot(dat)
+traitplot(dat) + theme_bw()
 ```
 
-# Poisson Regression
+# Intercept
 
 ```{r,eval=T}
+fitb<-function(dat,algorithm="Intercept"){
+
   #format traitmatch as matrix
@@
 
-runs<-100000
+  if(algorithm == "Intercept"){
+
+  runs<-10000
 
   #Source model
-source("Bayesian/Poisson.R")
+  source("Bayesian/Intercept.R")
 
   #print model
-print.noquote(readLines("Bayesian//Poisson.R"))
+  print.noquote(readLines("Bayesian//Intercept.R"))
+
+    #for parallel run
+    Yobs=dat$Interactions
+    Bird=as.numeric(as.factor(dat$I))
+    Plant=as.numeric(as.factor(dat$J))
+    Birds=length(unique(dat$I))
+    Traitmatch=Traitmatch
+    Plants=length(unique(dat$J))
+    Nobs<-length(Yobs)
+
+    #MCMC options
+    ni <- runs  # number of draws from the posterior
+    nt <- 1  #thinning rate
+    nb <- max(0,runs-500) # number to discard for burn-in
+    nc <- 2  # number of chains
+
+    modelDat<-list("Yobs","Bird","Plant","Plants","Birds","Nobs")
+
+    ParsStage <- c("alpha","alpha_mu","ynew","fit","fitnew")
+
+m1<-do.call(jags.parallel,list(data=modelDat,parameters.to.save=ParsStage,model.file="Bayesian/Intercept.jags",n.thin=nt, n.iter=ni,n.burnin=nb,n.chains=nc,DIC=F))
+} else
+
+    if(algorithm=="Binomial") {
+
+  runs<-10000
+
+  #Source model
+  source("Bayesian/Binomial.R")
+
+  #print model
+  print.noquote(readLines("Bayesian//Binomial.R"))
 
@@
     #Parameters to track
-  ParsStage <- c("beta1","ynew","fit","fitnew")
+    ParsStage <- c("alpha","beta","alpha_mu","beta_mu","ynew","fit","fitnew")
 
@@
 
-    m1<-do.call(jags.parallel,list(data=modelDat,parameters.to.save=ParsStage,model.file="Bayesian/Poisson.jags",n.thin=nt, n.iter=ni,n.burnin=nb,n.chains=nc,DIC=F))
+m1<-do.call(jags.parallel,list(data=modelDat,parameters.to.save=ParsStage,model.file="Bayesian/Binomial.jags",n.thin=nt, n.iter=ni,n.burnin=nb,n.chains=nc,DIC=F))  
+            }
+
+    m1$Algorithm<-algorithm
+    return(m1)
+}
+```
+
+# Interecpt Model
+```{r}
+intercept<-fitb(dat,"Intercept")
+```
+
+# Binomial Regression
+```{r}
+binomial<-fitb(dat,"Binomial")
+```
+
+#Extract parameters from the model
+
+```{r}
+getpars<-function(m1,n.iter=100,nt=1,nb=0,pars=c("beta1","alpha1")){  
+  parsm1<-extract_par(m1)
+  parsm1$Model<-m1$algorithm
+  return(parsm1)
+}
 ```
@@
 ```{r}
-parsO<-melt(m1$BUGSoutput$sims.array)
-colnames(parsO)<-c("Draw","Chain","parameter","estimate")
+#bind them together
+datpars<-list(intercept,binomial)
 
-#label species and plants
-l<-levels(parsO$parameter)
-
-#parameters to save
-totrack<-m1$parameters.to.save
-
-#assign species index to ragged frame.
-sp_pl<-data.frame(parameter=l,Index=as.numeric(str_match(l,pattern="\\[(\\d+)]")[,2]),par=str_extract(l,"\\w+"))
-
-#merge levels
-pars<-merge(parsO,sp_pl)
-
-pars<-pars[!pars$par %in% "deviance",]
-
+datpars<-lapply(datpars,getpars)
 ```
 
-```{r,fig.width=15}
-#view chaings
-ggplot(pars[pars$par %in% c("beta1"),],aes(x=Draw,col=as.factor(Chain),y=estimate))  + geom_line() + facet_wrap(~Index,scales="free") 
+#Predict new species
+
+```{r}
+size_x = 2
+size_y = 5
+
+#Bill sizes
+traits_x<-rpois(size_x,10)
+
+#Corolla sizes
+traits_y<-rpois(size_y,10)
+
+newtraits<-simA(size_x=size_x,size_y=size_y,traits_x=traits_x,traits_y=traits_y)
+
+#calculate traitmatch
+newtraits$Traitmatch<-abs(newtraits$TraitJ-newtraits$TraitI)
+
+predictM<-function(m1,newtraits){
+  parsm1<-extract_par(m1)
+
+  #matching function
+a<-parsm1 %>% filter(par %in% c("alpha_mu","beta_mu")) %>% select(Draw,Chain,par,estimate) %>% dcast(.,Draw + Chain ~ par,value.var="estimate") %>% group_by(Draw,Chain)%>%  do((pred.binomial(alpha=.$alpha_mu,beta=.$beta_mu,newdata=newtraits))) %>% group_by(I,J) %>% summarize(mean=mean(value),lower=quantile(value,0.95),higher=quantile(value,0.05))
+
+return(a)
+}
+
+predt<-predictM(binomial,newtraits)
 ```
 
-##Posterior predictive check
+# Model evaluation
+
+```{r}
+parsall<-rbind_all(list(parsm1,parsm2))
+```
+
+## Posterior predictive check
 
 ```{r,fig.height=4,fig.width=8}
-fitstat<-droplevels(pars[pars$par %in% c("fit","fitnew"),])
-fitstat<-dcast(fitstat,Draw+Chain~par,value.var="estimate")
+fitstat<-droplevels(parsall[parsall$par %in% c("fit","fitnew"),])
+fitstat<-dcast(fitstat,Draw+Model+Chain~par,value.var="estimate")
 
@@
 ab<-data.frame(x=ymin:ymax,y=ymin:ymax)
-p<-ggplot(fitstat,aes(x=fit,y=fitnew)) + geom_point() + theme_bw() + coord_equal()
-psim4<-p  + labs(x="Discrepancy of observed data",y="Discrepancy of replicated data") + geom_line(data=ab,aes(x=x,y=y)) + ggtitle("Simulated Data")
-psim4
+
+ggplot(data=fitstat,aes(x=fit,y=fitnew)) + geom_point(aes(col=Model)) + theme_bw() + coord_equal() + labs(x="Discrepancy of observed data",y="Discrepancy of replicated data") + geom_line(data=ab,aes(x=x,y=y)) + ggtitle("Simulated Data")
+```
+
+## Proportion of predicted links
+
+```{r}
+propc<-function(x,orig=dat){
+  predweb<-acast(data=x,I~J,value.var="Index")
+
+  #calculate network statistic
+  nstat<-networklevel(predweb,index=c("connectance","nestedness"))
+  data.frame(Metric=names(nstat),nstat)
+}
+
+prop_links<-parsall %>% filter(par=="ynew") %>% group_by(Model,Draw,Chain) %>% inner_join(dat) %>% do(propc(.))
 
@@
 
-#View simulated and predicted traitmatching
+###View trait-matching
 
 ```{r}
-
 #merge the dataframes.
@@
 
-# Compute emergent statistics
+### Compute emergent statistics
 
 ```{r}
-
 makeN<-function(x){
   predweb<-acast(data=x,I~J,value.var="estimate")
-
   #calculate network statistic
-  nstat<-networklevel(predweb,index=c("weighted connectance","weighted NODF"))
+  nstat<-networklevel(predweb,index=c("connectance","nestedness"))
   data.frame(Metric=names(nstat),nstat)
@@
 tweb<-acast(dat,I~J,value.var="Interactions")
-tstat<-networklevel(tweb,index=c("weighted connectance","weighted NODF"))
+tstat<-networklevel(tweb,index=c("connectance","nestedness"))
 tstat<-data.frame(Metric=names(tstat),tstat)
 
-ggplot(nstats,aes(x=nstat)) + geom_density(alpha=0.6,fill="black") + facet_wrap(~Metric,scales='free',nrow=2) + theme_bw()  + geom_vline(col="red",linetype="dashed",data=tstat,aes(xintercept=tstat)) + labs(x="value")
+ggplot(nstats[nstat$Metric %in% "connectance",],aes(x=nstat)) + geom_density(alpha=0.6,fill="black") + facet_wrap(~Metric,scales='free',nrow=1) + theme_bw()  + labs(x="value")
++ geom_vline(col="red",linetype="dashed",data=tstat,aes(xintercept=tstat)) 
 ```

### 2016-12-07 ms/intro_drafted.Rmd
link to manuscript

@@
 2006), (ii) quantifying ecosystem functioning and services (Thomson et al 2012), or (iii) predicting changes in from of global change scenarios (ref).
-In spite of increasing efforts, it is unlikely that we will ever manage to empirically document the massive amount of interactions existing in nature (Morales-Castilla et al.
+In spite of increasing efforts, it is unlikely that we will ever manage to empirically document the enormous amount of interactions existing in nature (Morales-Castilla et al.
 2015).
-Because of this, complementing field work with modelling is gaining acceptance among network ecologists (ref).
+Because of this, complementing field work with modelling of species interactions is gaining acceptance among network ecologists (ref).
 In a context where novel ecosystems arise as a result of species invasions or due to geographic shifts tracking climate change, our ability to predict future interactions will limit how efficiently we anticipate and respond to threats on biodiversity (i.e.

### 2016-12-07 Bayesian/Poisson.R

@@ -1,68 +1,0 @@
-
-sink("Bayesian/Poisson.jags")
-
-cat("
-    model {
-
-    #Compute intensity for each pair of birds and plants
-    for (i in 1:Birds){
-    for (j in 1:Plants){
-    
-    #Process Model with log normal overdispersion
-    log(lambda[i,j])<-alpha[i] + beta1[i] * Traitmatch[i,j] 
-    }
-    }
-    
-    
-    #Prediction
-    for (x in 1:Nobs){
-    
-    # Observed State
-    Yobs[x] ~ dpois(lambda[Bird[x],Plant[x]]) 
-    
-    #Assess Model Fit
-    
-    #Fit discrepancy statistics
-    eval[x]<-lambda[Bird[x],Plant[x]]
-    E[x]<-pow((Yobs[x]-eval[x]),2)/(eval[x]+0.5)
-    
-    ynew[x]~dpois(lambda[Bird[x],Plant[x]])
-    E.new[x]<-pow((ynew[x]-eval[x]),2)/(eval[x]+0.5)
-    
-    }
-    
-    ###Priors
-    
-    #Process Model
-    
-    #Species level priors
-    for (i in 1:Birds){
-    
-    #Intercept
-    alpha[i] ~ dnorm(alpha_mu,alpha_tau)
-    
-    #Traits slope 
-    beta1[i] ~ dnorm(beta1_mu,beta1_tau)    
-    }
-    
-    #Group process priors
-    
-    #Intercept 
-    alpha_mu ~ dnorm(0,0.001)
-    alpha_tau ~ dunif(0,10)
-    alpha_sigma<-pow(1/alpha_tau,0.5) 
-    
-    #Trait
-    beta1_mu~dnorm(0,0.001)
-    beta1_tau ~ dunif(0,10)
-    beta1_sigma<-pow(1/beta1_tau,0.5)
-    
-    #derived posterior check
-    fit<-sum(E[]) #Discrepancy for the observed data
-    fitnew<-sum(E.new[])
-    
-    
-    }
-    ",fill=TRUE)
-
-sink()

### 2016-12-19 Data/data_explnation.Rmd
Add clean data

@@
 
-__Note:__ This data needs more curation and can be used to play with methods, but needs taxonomic unification, better traits, etc... before looking at results.
-
 Data collected in 2015 across 16 sites. Each site was visited 7 times. Each time we recorded three data types:
 
-- All interactions in 100 m transects
-- Focal interactions to selected plants
+- All interactions in 100 m transects (`transects`)
+- Focal interactions to selected plants (`focal`)
 - Oportunistic new interactions not detected by the last two methods (`out`)
@@
 
-__Interactions:__ This include `transect` and `visitation` data.
-Data can be used pooled (1445 interactions (5643 visits) general question abouts trait matching) or splited by site (`Site_ID`) and `Round`.
-Pollinator traits in this table are measured per individual.
+__Interactions:__ This include `transect`, `visitation` and `out` data.
+Data can be used pooled (2067 interactions (5732 visits) general question abouts trait matching) or splited by site (`Site_ID`) and `Round`.
 ```{r}
@@
 ```
+__NOTE:__ Quite a lot of NA's still.
+
 Challenges:  
 - We are not sure which trait are driving interactions (quite generalized system)  
-- Need to inegrate traits across different orders
+- Need to inegrate traits across different orders (e.g. tongue lengths can be set to zero for coleoptera)
 
-__plant traits:__ Not ready yet
+__plant traits:__ 
+```{r}
+plant_tr <- read.csv("traits_plants.csv", h = T)
+head(plant_tr)
+```
+Those are complete.
 
@@
 
-__Phylogenies__ work in progress.  
-- Plants: Phylomatic  
-- Pollinators: Available at genus level only
+__Phylogenies__ Need to do.
+- Plants: Phylomatic? 
+- Pollinators: Bees available at genus level only. For other orders I have no idea.
+
+# combining data
+
+Question: should we use the pooled data for the package?
+```{r}
+head(dat)
+temp <- merge(int, pol_tr, by.x = "Pollinator_gen_sp", by.y = "Pollinator_gen_sp", all.x = TRUE)
+temp2 <- merge(temp, plant_tr, by.x = "Plant_gen_sp", by.y = "Plant_gen_sp", all.x = TRUE)
+pooled_data <- temp2
+head(pooled_data)
+```

### 2016-12-20 Data/data_explnation.Rmd
fix small typo

@@
 ```{r}
-head(dat)
 temp <- merge(int, pol_tr, by.x = "Pollinator_gen_sp", by.y = "Pollinator_gen_sp", all.x = TRUE)

### 2017-01-28 ms/ms_interactions.tex
Reorganized the manuscript in .tex format because the .md would not compile in its current form.

\input{ms_basis.tex}

\begin{document}

\section{Title:}\label{title}

\begin{itemize}
  \item Inferring species interactions from ecological data
  \item A quantitative comparison of methods for Inferring species interactions
  from ecological data
\item Comparing inferences for species interactions from ecological data
\item On the use of ecological data for comparing species interactions
\item Comparing methods for inferring species interactions
\item A statistical analysis of methods for inferring species interactions.
\item ?
\end{itemize}

\section{Authors*}\label{authors}

\begin{itemize}
  \item Ignasi Bartomeus
  \item Kevin Cazelles
  \item F. Guillaume Blanchet
  \item Mickaël Hedde
  \item Ignacio Morales Castilla
  \item Mathilde Besson
  \item David Beauchesne
  \item Dominique Gravel
  \item Timothée Poisot
  \item Ben Weinstein
  \item Steve Vissault
\end{itemize}

\emph{Current author order drawn from a multinomial distribution, which,
ironically, might be the only acceptable use case for this distribution in this
manuscript.}

\emph{Target Journals}: Methods in Ecology and Evolution, Ecological Monographs,
Annual Review of Ecology, Evolution, and Systematics.

\section{Introduction}\label{introduction}

Biotic interactions form the backbone of biodiversity
\cite{bascompte_plant-animal_2007}. Understanding the strength, direction, and
symmetry of species interactions provides insights into the maintenance of
assemblages and potential threats from anthropogenic change. Combined with
abiotic factors, biotic interactions constrain the distributions of species,
shape the evolution of phenotypes, and influence the stability of natural
systems \cite{schleuning_predicting_2015}. A common tool for understanding
interaction are ecological networks, in which each species (nodes) are connected
based on the strength and direction of interactions (links)(Jordano 1987).
Documenting and analyzing these networks is essential to understand the rules
underlying pairwise species interactions (Bascompte et al. 2006), quantifying
ecosystem functioning (Thomson et al 2012), and predicting the biodiversity
consequences due to anthropogenic change (ref).

Species traits are at the core of biological interactions (Bartomeus et al.
2016). In this paper, we defined traits as phenotypic adaptations to exploit abiotic or
biotic resources. We used traits as our principal mechanism to explain when
species can interact. The foremost requirement for species interactions is
species co-occurrence, which is itself driven by traits through environmental
filtering and local adaptation (Hillerislambers, Mayfield, but see Kraft). If
species co-occur, they may interact if their respective traits promote the
efficiency of resource extraction. In addition, traits can be represented by
phylogenetic information, in which evolution history constrains trait evolution
either directly through phylogenetic conservatism or indirectly through exposure
to similiar environmental conditions (Graham and Wiens 2005).

The traits promoting species co-occurrence and interaction may be hidden from
us, either through lack of sampling, or due to their complexity. In these cases,
the co-occurrence of species is used as a proxy for interaction. Given the
correlation in distribution or abundance of species across space or time,
co-occurrence based methods attempt to circumvent trait-based models by assuming
that species which consistently co-occur, and whose populations co-vary, may
have important biotic interactions. We therefore see co-occurrence based methods
as an extension of specific trait-based models when the mechanisms promoting
interactions are unknown or unobserved.

In spite of increasing efforts, it is unlikely that ecologists will empirically
document the enormous number of interactions existing in nature
(Morales-Castilla et al. 2015). Therefore, complementing field work with models
of species interactions is a promising avenue for understanding the complexity
of the natural world(ref). While direct inference will often be preferable, many
interactions are rare, infrequent, or impossible to detect by human observation.
Model-based inference for ecological networks provides three important aspects,
i) it establishes mechanistic relationships among species, ii) it estimates
uncertainty given empirical observation, and iii) it provides an avenue to
predict future interactions. We follow Poisot (2015) and Cazelles (2016) in
formulating our general conceptual model of species interactions (Fig. 1). There
is a pool of potential interactions, termed a metaweb, in which species could
interact based on co-occurrence. Any given observation we make is an attempt to
recover these relationships using empirical data. We use these data to fit a
model based on either species taxonomic identity or phenotypic characteristics.
This model can used to predict the interactions among observed species, given
the uncertainty due to spatial and temporal stochasticity of interactions, or a
new network, based on the presence of a novel species in the assemblage. From
these predictions, we can estimate structural network properties, such as
connectance and nestedness, while simultaneously propagating our uncertainty
throughout the analysis. In this paper, we review the methods proposed to build
models of species interactions, emphasizing each method's utility for predicting
the position of a novel species in the network.

From this conceptual framework, we can decide on a set of optimal criteria to
compare existing models. While there is unlikely to be any single method that
meets all our expectations, by defining a set of standards, we can more
transparently judge the relative merits of each method for our particular goal.
Our foremost goal is to generate a prediction of the links among co-occurring
species based on an estimated probability of interaction. This prediction should
come with a statement of uncertainty regarding our confidence in parameter
estimates based on empirical data. An ideal method will make reasonable
assumptions about the interdependence of observations, and be relatively
insensitive to low sampling effort. The mechanism determining the links among
species should be clear and grounded within ecological theory, such that we can
use our method to make a prediction about the placement of a new species in the
network based on a fitted model.

\subsection{Models of species
interactions}\label{models-of-species-interactions}

\subsubsection{Regression-based Models}\label{regression-based-models}

The simplest model of species interactions is

eq i.

\[Y_{i,j} = \rho_{i,j}\] Where the probability of interaction among species i
and species j is a fixed value (\(\rho_{i,j}\)). Using empirical data, we might
try to estimate this value by dividing the number of observations by the total
number of observations for any pair of species. While this is straightforward,
why might it be insufficient? Foremost, it contains no estimate of uncertainty.
The estimate is given without any consideration for the variance due to
sampling, spatial or temporal stochasticity. To add uncertainty, we draw our
observations from a statistical distribution.

eq ii.

\[ Y_{i,j} \sim Bernoulli(\rho_{i,j}) \]

We can now estimate the probability of interaction, as well as the uncertainty
based on sampling stochasticity. Yet this model is still unsatisfying. It is
non-mechanistic, it depends wholly on species identity, and uses no ecological
theory or inference. Therefore, it is difficult to ascribe meaning to a set of
interactions, or understand how they might change in the future. We cannot use
the inferences from data to make a prediction about a new species not in
dataset.

To make our model more useful, we might add a covariate, such that

eq. iii

\[ Y_{i,j} \sim Bernoulli(\rho_{i,j})\] \[ logit(\rho_{i,j}) = \alpha_{i,j} +
\beta_{i,j} * x_{i,j} \]

In which a link between species i and j is a Bernoulli trial with a probability
of success \(\rho_{i,j}\). The probability of success will vary based on some
ecological principle (\(x_{i,j}\)), for example, the similarity in functional
traits among species i and j:

eq. V

\[x_{i,j}= f(Trait_i,Trait_j)\]

A variety of trait-matching functions may be biologically plausible. For
example, the absolute value of the difference between traits (e.g. \(|trait_i -
trait_j|\)) or a binary difference (e.g. \(trait_i > trait_j\)), could be used
to predict which species may interact. Alternatively, the covariate may
represent the phylogenetic relatedness of species, or their relative response to
a given environmental variable.In some cases, we cannot provide a simple
trait-matching function (eg V), but must use a statistical tool to infer the
best possible combination of traits among co-occurring species that predicts
interactions. Machine learning tools provide an ideal compliment to traditional
linear models when we cannot anticipate the combinations of traits that can
influence species interaction rates. Unsupervised learning tools, such as random
forests, can give us the associations among species given a set of complex
traits. Regardless of the exact form of regression, this framework satisfies two
important goals: it allows for estimates of uncertainty in species interactions,
and it provides a mechanism shaping the probability of an interaction.

So far, we have only captured the uncertainty in the process model, the inherent
temporal or spatial stochasticity in our ecological process of interest (Hooten
and Hobbs). We may need to also account for the uncertainty in the observation
model, the ability to detect interactions given that they occur. The separation
of the observation and process model translates our ecological mechanism into
specific predictions about empirical data. For binary networks, one
straightforward way to account for the observation of interactions is to model
the detectability of network interactions.

eq iv.

\[ Y_{i,j,k} \sim Bernoulli(\phi_{i,j})\] \[ \phi_{i,j} = \omega_{i,j} *
z_{i,j}\] \[z_{i,j} \sim Bernoulii(\rho_{i,j})\] \[ logit(\rho_{i,j}) =
\alpha_{i,j} + \beta_{i,j} * x \]

Where the observation of a link between species i and species j at time k is a
Bernoulli trial with a probability \(\phi_{i,j}\), which is the outcome of the
detectability of an interaction (\(\omega_{i,j}\)) and the latent state
(\(z_{i,j}\)). This latent state is the true, but unobserved, existence of a
link, as predicted by our ecological mechanism of interest (see eq IV). Our
ability to detect this latent state depends on the detectability of an
interaction (\(\omega_{i,j}\)). The benefit of this approach is that
differentiates the probability of detection from the probability of occurrence.
Importantly, because we have explicitly defined some temporal window (k),
interactions with different levels of sampling effort can be directly compared.

\subsubsection{Logic-based methods}\label{logic-based-methods}

{[}Ben Weinstein: For someone to fill in.{]}

\subsubsection{Occurrence-based methods}\label{occurrence-based-methods}

{[}Ben Weinstein: For someone to fill in.{]}

\subsubsection{Non-parametric methods}\label{non-parametric-methods}

The above methods are model based, that is, they largely use parametric
distributions to estimate the liklihood of the data given estimated model
parameters. An alternative approach is to use multivariate statistics to create
association matrices. One such method is called fourth-corner analysis, which is
a type of three table ordination (sometimes called `RLQ' analysis) (Doledec
1996). In bipartite networks, fourth-corner analysis relates the matrix of
species interactions to a matrix of traits from one level (e.g.~plants) to
another level (e.g.~pollinators) via the matrix of species innteractions (Dray \&
Legendre, 2008). This produces a correlation matrix, which is then compared to a
null expectation through matrix randomization. Randomization is a familiar tool
in network ecology, in which permuatation rules are used to compare the observed
test statistic (e.g the correlation among species occurrences) to a null
distribution based on some matrix constraint. Common constraints include the
mantainance of row sums, column sums, or other all marginal values, such that
only the identity of partners are switched. After each permutation, the test
statistic is recalulcating, yielding a null distribution. By comparing whether
the observed test statistic falls within the upper or lower 5th quantile of the
null distribution, studies distinguish whether their results could have been
generated at random given their sampling effort.

\section{Methods}\label{methods}

Our goal is provide a conceptual framework for the broad array of potential
methods described above. Our three criteria for qualitative comparison are: 1)
Can the method account for the uncertainty in species observations?, 2) Does it
provide an ecological mechanism to infer the probability of species
interactions, and 3) Can it be used to predict an novel interaction? In addition
to qualitative comparison, we provide a quantitative analysis of methods for
inferring species interactions. From a quantitative perspective, we are
interested in methods which accurately reflect true relationships among data,
are relatively insensitive to sampling effort, and make reasonable assumptions
about the interdependence of species observations. We use a combination of
simulated and empirical data to establish a benchmark dataset to compare the
accuracy and sensitivity of currently proposed methods.

\section{Quantitative comparisons}\label{quantitative-comparisons}

-\textgreater{} results: - cross comparison \textbar{} real and simulated data -
-sensitivity \textbar{} - interpretation of probabilities - method description
-for comparison

\section{Results}\label{results}

\section{Discussion}\label{discussion}

\subsection{Perspectives}\label{perspectives}

\begin{itemize} \item guidelines \item challenges \item next steps \end{itemize}

\subsection{On the virtue of prediction of species
interactions.}\label{on-the-virtue-of-prediction-of-species-interactions.}

In a context where novel ecosystems arise as a result of species invasions or
due to geographic shifts tracking climate change, our ability to predict future
interactions will limit how efficiently we anticipate and respond to threats on
biodiversity (i.e.~species local extinctions) and on humans (i.e.~spread of
infectious disease or plagues affecting to crops).

\emph{{[}Ben Weinstein: Previous text from Nacho:{]}} Indeed, on the quest for
describing species interactions, traits had played a pivotal role and several
methods have recently emerged to test hypotheses of trait matching for pairwise
interactions (Stang et al 2006, Dehling et al. 2014, Bartomeus et al 2016) or
understand the phylogenetic structure of ecological networks (Rezende et al.
2007). However, generating biological meaningful predictions of species
interactions and its emerging properties have proven challenging (Olito and Fox
2014). An ideal method is grounded in ecological theory, which makes the
description of ecological interactions interpretable. This method should be able
to not only describe, but also predict ecological interactions based on a set of
simple parameters. Importantly, this predictions should be accompanied by an
associated uncertainty. In addition, given the difficulty of empirically
assessing complete interaction networks, this method should be also robust to
the use of incomplete data. There is an urgent need to better understand how
communities interact and predict new interactions which will inevitably occur as
a result of human induced rapid environmental change. In addition, our
understanding of how species interact is limited to the few interactions are
directly observable such as pollination of predation, but key species
interactions not directly observable remains unlocked. Interactions inferences
may be the only option for describing and predicting interactions in soil food
webs or bacteria-phage networks. Finally, traits also relate to the fitness
effects gained by both interaction partners. This can potentially establish a
link with to modern coexistence theory. In fact, neutral expectations
postulating that species abundances are the main drivers of interactions, are
not incompatible with the underlying role of species traits shaping species
abundances via effects in fitness (Bartomeus et al. 2016).

\section{Literature cited}\label{literature-cited}

\bibliography{referencesNew}

\section{Boxes}\label{potential-boxes}

\subsection{Soil ecology}\label{soil-ecology}

\subsection{Climate changes and alien
species}\label{climate-changes-and-alien-species}

\subsection{Community/evolution ecology}\label{communityevolution-ecology}

\input{ms_fig.tex}

\input{ms_tables.tex}

\end{document}

### 2017-01-28 ms/tables/comparison.tex
Reorganized the manuscript in .tex format because the .md would not compile in its current form.

\begin{sidewaystable}
\centering
\caption{Methods compared and how they apply}
\label{comparisonTab}
\begin{tabular}{llccccccccccc}
                                                & \multicolumn{5}{c}{Input Data}                                                                                                                                         & \multicolumn{1}{c}{}              & \multicolumn{1}{c}{}              & \multicolumn{1}{c}{}           & \multicolumn{1}{c}{}       & \multicolumn{1}{c}{}                   & \multicolumn{1}{c}{}                                                                                                                     \\
                                                & \multicolumn{1}{c}{abundance} & \multicolumn{1}{c}{Indirect  knowledge} & \multicolumn{1}{c}{Co-occurence} & \multicolumn{1}{c}{Trait} & \multicolumn{1}{c}{Phylogeny} & \multicolumn{1}{c}{Theory-driven} & \multicolumn{1}{c}{Type of links} & \multicolumn{1}{c}{Predictive} & \multicolumn{1}{c}{Output} & \multicolumn{1}{c}{Hypothesis testing} & \multicolumn{1}{c}{Example}                                                                                                              \\
Fourth-corner                                   & x                             &                                         &                                  & x                         &                               & yes                               & quantitative                      & mecanistic                     &                            &                                        & \begin{tabular}[c]{@{}l@{}}I\_eat (Beauschene et al. 2016)\\ Markov logic, Intuitive Logic Programming (Vacher et al. 2016)\end{tabular} \\
Maching learning                                & x                             & x                                       & (x)                              & (x)                       & (x)                           & no                                & quantitative                      & data-based                     &                            &                                        & Rohr et al. 2014, Rohr et al. 2016                                                                                                       \\
Direct mathcing-centrality                      &                               &                                         &                                  & x                         & x                             & yes                               & quantitative                      & mecanistic                     &                            &                                        &                                                                                                                                          \\
Indirect matching-centrality                    &                               &                                         &                                  &                           &                               &                                   &                                   &                                &                            &                                        &                                                                                                                                          \\
Niche model                                     &                               &                                         &                                  &                           &                               &                                   &                                   &                                &                            &                                        &                                                                                                                                          \\
Hierarchical model                              &                               &                                         &                                  &                           &                               &                                   &                                   &                                &                            &                                        &                                                                                                                                          \\
Null model                                      &                               &                                         &                                  &                           &                               &                                   &                                   &                                &                            &                                        &                                                                                                                                          \\
Joint Spatial Division and Multiplexing (JSDM)  &                               &                                         &                                  &                           &                               &                                   &                                   &                                &                            &                                        &                                                                                                                                          \\
Intercept                                       &                               &                                         &                                  &                           &                               &                                   &                                   &                                &                            &                                        &                                                                                                                                          \\
GLM/cooccur                                     &                               &                                         &                                  &                           &                               &                                   &                                   &                                &                            &                                        &                                                                                                                                          \\
Grouping species into nodes (functional groups) &                               & x                                       &                                  & x                         &                               & no                                & Binary (quantitative)             & no                             &                            &                                        &                                                                                                                                          \\
Expert-based assesment of links                 &                               & x                                       &                                  & x                         &                               & no                                & Binary (quantitative)             & no                             &                            &                                        &
\end{tabular}
\end{sidewaystable}

### 2017-02-03 ms/ms_interactions.tex
Change margin, add line numbers and double space manuscript
Worked on the Introduction

@@
 \begin{document}
+% Add line number
+\linenumbers
+\modulolinenumbers[1]
+
+% Double space
+\doublespacing
 
@@
 based on the strength and direction of interactions (links)(Jordano 1987).
-Documenting and analyzing these networks is essential to understand the rules
-underlying pairwise species interactions (Bascompte et al. 2006), quantifying
-ecosystem functioning (Thomson et al 2012), and predicting the biodiversity
-consequences due to anthropogenic change (ref).
+Documenting and analyzing these networks is essential to better understand the
+rules underlying pairwise species interactions (Bascompte et al. 2006),
+quantifying ecosystem functioning (Thomson et al 2012), and predicting the
+biodiversity consequences due to anthropogenic change (ref).
 
-Species traits are at the core of biological interactions (Bartomeus et al.
-2016). In this paper, we defined traits as phenotypic adaptations to exploit abiotic or
-biotic resources. We used traits as our principal mechanism to explain when
-species can interact. The foremost requirement for species interactions is
-species co-occurrence, which is itself driven by traits through environmental
-filtering and local adaptation (Hillerislambers, Mayfield, but see Kraft). If
-species co-occur, they may interact if their respective traits promote the
-efficiency of resource extraction. In addition, traits can be represented by
-phylogenetic information, in which evolution history constrains trait evolution
-either directly through phylogenetic conservatism or indirectly through exposure
-to similiar environmental conditions (Graham and Wiens 2005).
+The foremost requirement for species interactions is species co-occurrence,
+which is itself driven by species traits through environmental filtering and
+local adaptation (Hillerislambers, Mayfield, but see Kraft). Knowing where
+species are located gives us insights into the interactions they may have, but
+to more deeply understand the mechanisms that influence the relationships among
+species, traits are necessary. As such, species traits are at the core of
+biological interactions (Bartomeus et al. 2016). Here, we defined traits as
+phenotypic or genotypic adaptations, that may vary through time, to exploit
+abiotic or biotic resources. Note that this definition indirectly include
+phylogeny as a form of trait. From this definition, conceptually, if species
+co-occur, we assume that they may interact only if their respective traits
+promote the efficiency of resource extraction. When traits are represented by
+phylogenetic information, it is the evolution history that constrains trait
+evolution either directly through phylogenetic conservatism or indirectly
+through exposure to similiar environmental conditions (Graham and Wiens 2005).
 
-The traits promoting species co-occurrence and interaction may be hidden from
-us, either through lack of sampling, or due to their complexity. In these cases,
-the co-occurrence of species is used as a proxy for interaction. Given the
-correlation in distribution or abundance of species across space or time,
-co-occurrence based methods attempt to circumvent trait-based models by assuming
-that species which consistently co-occur, and whose populations co-vary, may
-have important biotic interactions. We therefore see co-occurrence based methods
-as an extension of specific trait-based models when the mechanisms promoting
-interactions are unknown or unobserved.
+In practice, the traits promoting species co-occurrence and interaction may be
+hidden from us, either because they were not sampled or due to their complexity.
+In such cases, co-occurrence of species is used as a proxy for interaction.
+Given the correlations in occurrence or abundance of species across space or
+through time, co-occurrence based methods attempt to circumvent trait-based
+models by assuming species that consistently co-occur, and whose populations
+co-vary, may have important biotic interactions. In this context, co-occurrence
+based methods are used to describe the potential relationships species may have
+while trait-based models consider the mechanisms promoting interactions.
 
@@
 of species interactions is a promising avenue for understanding the complexity
-of the natural world(ref). While direct inference will often be preferable, many
-interactions are rare, infrequent, or impossible to detect by human observation.
-Model-based inference for ecological networks provides three important aspects,
-i) it establishes mechanistic relationships among species, ii) it estimates
-uncertainty given empirical observation, and iii) it provides an avenue to
-predict future interactions. We follow Poisot (2015) and Cazelles (2016) in
-formulating our general conceptual model of species interactions (Fig. 1). There
-is a pool of potential interactions, termed a metaweb, in which species could
-interact based on co-occurrence. Any given observation we make is an attempt to
-recover these relationships using empirical data. We use these data to fit a
-model based on either species taxonomic identity or phenotypic characteristics.
-This model can used to predict the interactions among observed species, given
-the uncertainty due to spatial and temporal stochasticity of interactions, or a
-new network, based on the presence of a novel species in the assemblage. From
-these predictions, we can estimate structural network properties, such as
-connectance and nestedness, while simultaneously propagating our uncertainty
-throughout the analysis. In this paper, we review the methods proposed to build
-models of species interactions, emphasizing each method's utility for predicting
-the position of a novel species in the network.
+of the natural world(ref). While direct inference is often preferable, many
+interactions are rare, infrequent, or impossible to detect through human
+observation. Model-based inference for ecological networks provides three
+important aspects to characterize species interactions, i) it establishes
+mechanistic relationships among species, ii) it estimates uncertainty given
+empirical observation, and iii) it provides an avenue to predict future
+interactions. We follow Poisot (2015) and Cazelles (2016) in formulating our
+general conceptual model of species interactions (Fig. 1). There is a pool of
+true interactions, a metaweb, in which individuals of different species co-occur
+resulting in interactions among species. Any given observation we make on this
+metaweb is an attempt to recover the relationships among species using empirical
+data. Based on these data, model(s) are built to characterize and elucidate the
+relationships among species. Some models rely exclusively co-occurrence of
+individuals of different species at particular locations and may also include
+information on phylogenetic history and species traits. These models can than be
+used to predict the interactions among observed species, given an uncertainty
+due to the spatial and/or temporal stochasticity of the interactions. With these
+models it is also possible to define new network, based on the presence of a
+novel species in the assemblage. From these predictions, we can estimate
+structural network properties, such as connectance and nestedness, while
+simultaneously propagating our uncertainty throughout the analysis. In this
+paper, we reviewed the methods proposed to build models of species interactions,
+emphasizing each method's utility for predicting the position of a novel species
+in the network.
 
 From this conceptual framework, we can decide on a set of optimal criteria to
-compare existing models. While there is unlikely to be any single method that
-meets all our expectations, by defining a set of standards, we can more
-transparently judge the relative merits of each method for our particular goal.
-Our foremost goal is to generate a prediction of the links among co-occurring
-species based on an estimated probability of interaction. This prediction should
-come with a statement of uncertainty regarding our confidence in parameter
-estimates based on empirical data. An ideal method will make reasonable
-assumptions about the interdependence of observations, and be relatively
-insensitive to low sampling effort. The mechanism determining the links among
-species should be clear and grounded within ecological theory, such that we can
-use our method to make a prediction about the placement of a new species in the
-network based on a fitted model.
+compare existing approaches. While it is unlikely that any single method meets all
+our expectations, by defining a set of standards, we can more transparently
+judge the relative merits of each method to reach particular goals. The most
+important aspect of a method is for it to generate accurate prediction of the
+links among co-occurring species based on an estimated probability of
+interaction. This prediction should come with a statement of uncertainty
+regarding our confidence in parameter estimates based on empirical data. An
+ideal method would also be able to make reasonable assumptions about the interdependence of
+observations, and be insensitive to low sampling effort. The
+mechanism determining the links among species should be clear and grounded
+in ecological theory such that we can use our method to make a prediction
+about the placement of a new species in the network based on a fitted model.
 
@@
 
+A handful of soil is the home of countless number of organisms. Soils host about
+the quarter of Earth biodiversity (Decaëns et al 2006) and is recognised as the
+third biotic frontier (André et al 1994). But three crucial points arise when
+dealing with soil biodiversity and soil biotic interactions, (i) it is hard to
+observe organisms without destroying their habitat, (ii) most of soil organisms
+are tiny and cannot be bred or manipulated and (iii) soil organisms are prone to
+a “Linnean deficit”. Biotic interactions in soils are shaped by soil
+specificities that lead to an intense entanglement of trophic and non-trophic
+interactions. For instance, soils are physically structured by the activity of
+ecosystem engineers (earthworms, termites, ants, roots) that modifies, by their
+own niche construction, the accessibility to resources for other soil organisms
+(Jones et al 1994).
+
+Body size of soil organisms ranges from nanometric scale (phage, bacteria) to
+decimetric one (earthworms, moles, etc), leading to nested home ranges. Soils
+are highly organized environments (in the three dimensions and in time). This
+permits high level of niche segregation (to diminish competition) and numerous
+way to escape from predators. On the most famous example of soil trophic network
+comes from short grass meadows in Colorado (Hunt et al. 1987), in which nodes
+were a mix of taxa, feeding guilds and pools of organic matter. The observation
+of soil species interactions is hard and interaction can be inferred by various
+ways. In the case of trophic interaction, this can be achieved through, for
+example, gut content analyses (observation-based or DNA-based), (bio)chemical
+analyses (eg PLFA, stable isotopes content), predator-prey body size
+relationship, co-occurrence analysis, machine learning or a combination of them
+(Digelet al 2014). Given the lack of knowledge on microbes’ biology and ecology,
+co-occurrence is often interpreted as interaction (Barberan et al 2012).
+Conversely to the green-world hypothesis (predator limitation of herbivores can
+cascade to facilitate plant growth), soil food webs are assumed to be bottom-up
+controlled detritus-based communities, but yet the ecosystem role of predators
+in detrital food webs is less explored. Models of energy fluxes through soil
+trophic networks remains rare (de Vries et al 2013), and nutrient cycling is
+generally ignored while connecting above and belowground trophic networks (Zou
+et al 2016). Hence, soil food webs studies are mainly descriptive and have poor
+predictive ability.
+
+Given the importance of plant for humankind, it is crucial to understand the how
+multiple interactions control plant production for food, fibre and wood. Soils
+are the place of numerous symbiotic relationships between plants and bacteria
+(eg Rhizobia) or fungi (mycorrhiza). The soil microbial loop an example of
+hypothesis that links multiple partners of the rhizospheric microbial food web
+to explain protozoa (predators) effects on plant growth. As carbon-rich root
+exudates (high C to N ratio) are continuously released into the plant
+rhizosphere, those soil bacteria that consume the exudates need other sources of
+nitrogen to support their growth. It is suggested that the bacteria therefore
+increase their assimilation of indigenous soil organic nitrogen. When protozoa
+subsequently consume the bacteria, part of this nitrogen will be released into
+the soil inorganic nitrogen pool, which will increase the nitrogen pool
+available for plant uptake. It has also been demonstrated that signal molecules
+emitted by invertebrates can mediate plant growth and plant disease immunity
+(resistance or recovery). Interactions between soil organisms and roots also
+shape the plant community assembly and the aboveground trophic interactions, up
+to fourth levels (plant-herbivore-parasite-parasitoïd).
+
+Recently, DNA-based approach has been democratised to study co-occurrence of
+cryptic interactions like Ericaceous plant-fungus symbioses (Toju et al 2016),
+earthworm-gut bacteria symbioses (Moller et al 2015) or predator-prey
+interactions (Bohan xxx). The tremendous development of next-generation DNA
+sequencing will surely provide the most comprehensive and accurate measure of
+trophic links in ecosystems, regardless of the size of organisms and without a
+priori on the species consumed by a given organism, by matching sequences
+amplified from the gut content to wide gene databases (Pompanon et al. 2012).
+Finally, the integration of biogeography and network theory will yield important
+results and further our understanding and forecasting ability of species
+interactions, in a context of global changes.
+
 \subsection{Climate changes and alien

### 2017-02-09 Data/data_explnation.Rmd
update explnation.Rmd

@@
 Data can be used pooled (2067 interactions (5732 visits) general question abouts trait matching) or splited by site (`Site_ID`) and `Round`.
+
 ```{r}
@@
 ```
+
+# Reshaping the dataset as input of the `as.alienData()` function
+
+**Extract from the roxygen2 documentation:**
+
+```
+#' @param idObs A data.frame which is mandatory and will help to check consistency and prevent errors among unique identifiers of each alienData arguments.
+The first column (idSite) contains unique identifier of where the observation was made.
+The second column (idTime) is not mandatory and contains temporal information: an unique identifier at the time the sample has been taken (needed for timeseries analysis).
+The third column (idSpcies) is an unique identifier of the species sampled at time (idTime) and location (idSite).
+The fourth column is an unique identifier of individu of species (idSp) observed at time (idTime) and location (idSite).
+#' @param interactPair A data.frame which contains interaction at the finest level (individus or species).
+The first two columns are idFrom and idTo and determine the sens of the interaction.
+idFrom and idTo are unique identifier of species or individu documented in the idObs data.frame.
+Finaly, the thrid column is the strength of the interaction (Please see details).
+#' @param coOcc A square symmetric matrix of 0s and 1s that define co-occurence patterns among pairs of species.
+If this matrix is not provided, the co-occurence matrix is derived from the coAbund matrix else the interactSp matrix (see return section).
+#' @param coAbund A square symmetric matrix that includes any types of values, defining co-abundance patterns among pairs of species. TODO: Not implemented yet.
+#' @param siteEnv A matrix or a data.frame where each column is a descriptor of the sites. TODO: siteEnv should cover the possibility that environmental variables could be taken at several times.
+#' @param traitSp A matrix or a data.frame where each column is a trait characterizing all species. The first column is a unique identifier of the species documented in idObs data.frame.
+#' @param traitInd A matrix or a data.frame where each column is a trait characterizing an individual. The first column is a unique identifier of the individu documented in idObs data.frame.
+#' @param phylo A square symmetric matrix describing the phylogenetic relationships between pairs of all species (see details). TODO: Not implemented yet.
+#' @param scaleSiteEnv Logical. Whether the columns of X should be centred and divided by the standard deviation. Default is TRUE.
+#' @param scaleTrait Logical. Whether the rows of Tr should be centred and divided by the standard deviation. Default is TRUE.
+#' @param interceptSiteEnv Logical. Whether a column of 1s should be added to X. Default is TRUE.
+#' @param interceptTrait Logical. Whether a row of 1s should be added to Tr. Default is TRUE.
+#'
+#' @details
+#'
+#' The strength of the interactions defined in the third column of \code{interactPair} can be a 0 if no direct interaction has been observed (defined as true absence of interaction) or any numerical value.
+Undocumented interactions among species or individus will be assumed as NA by default.
+```
+
+```{r, echo=FALSE}
+# Create dir for formatted data
+dir.create('./Data/alienData', showWarnings = FALSE)
+```
+
+### Each section below transform your data into as.alienData args
+
+#### Get idObs
+
+```{r}
+# load data
+int <- read.csv("interactions.csv", h = T, stringsAsFactors=FALSE)
+flw <- read.csv("flowers.csv", h = T, stringsAsFactors=FALSE)
+plant_tr <- read.csv("traits_plants.csv", h = T, stringsAsFactors=FALSE)
+pol_tr <- read.csv("traits_pollinators.csv", h = T, stringsAsFactors=FALSE)
+
+# Create unique IDs for pollinator individus
+int$idPol <- seq(1,nrow(int),1)
+# Create unique IDs for plant individus
+int$idPlant <- seq(nrow(int)+1,nrow(int)+nrow(int),1)
+
+# subset pollinator and plant with their own ids
+pol <- unique(int[,c("Site_ID", "Round","Pollinator_gen_sp","idPol")])
+colnames(pol) <- c("idSite","idTime","idSp","idInd")
+plant <- unique(int[,c("Site_ID", "Round","Plant_gen_sp","idPlant")])
+colnames(plant) <- c("idSite","idTime","idSp","idInd")
+
+# Set idObs by merging both dfs
+idObs <- rbind(pol,plant)
+
+#Turn all columns as factor
+idObs <- as.data.frame(lapply(idObs,as.factor))
+```
+
+#### Get InteracPair
+
+```{r}
+interactPair <- int[,c("idPol","idPlant","Frequency")]
+```
+
+#### traitInd
+
+Trait by individus
+
+```{r}
+traitInd <- int[,c("idPol","Pollinator_sex")]
+traitInd$traitName <- "sex"
+traitInd <- traitInd[,c('idPol','traitName','Pollinator_sex')]
+names(traitInd)[3] <- 'value'
+```
+
+#### traitSp
+
+Trait by species
+Note Removed family as a trait
+
+```{r}
+library(reshape2)
+
+traitSpPlant <- melt(plant_tr[,-2],id=c("Plant_gen_sp"),value.name="value",variable.name="traitName")
+names(traitSpPlant)[1] <- "idSp"
+
+traitSpPol <- melt(pol_tr[,-4],id=c("Pollinator_gen_sp"),value.name="value",variable.name="traitName")
+names(traitSpPol)[1] <- "idSp"
+
+traitSp <- rbind(traitSpPlant,traitSpPol)
+traitSp$idSp <- as.factor(traitSp$idSp)
+
+#### errors
+# Some entries in traitSp (merge of pol_tr and plant_tr) are documenting species which are not present in `ìnt`, the interaction pairwise data.frame.
+removeSpecies <- levels(traitSp$idSp)[which(!levels(traitSp$idSp) %in% levels(idObs$idSp))]
+traitSp <- subset(traitSp, !idSp %in% removeSpecies )
+traitSp$idSp <- droplevels(traitSp$idSp)
+```
+
+#### Save all args of the as.alienData() into a RData file.
+
+```{r}
+save(idObs,interactPair,traitSp,traitInd,file="./alienData/argsAlienData.RData")
+```

### 2017-02-09 ms/ms_interactions.tex
Continued working on the method section of the manuscript

@@
 
-\section{Title:}\label{title}
+\section{Title:}
+\label{title}
 
@@
 
-\section{Introduction}\label{introduction}
+\section*{Abstract}
+\label{abstract}
+
+\section*{Introduction}
+\label{introduction}
 
@@
 document the enormous number of interactions existing in nature
-(Morales-Castilla et al. 2015). Therefore, complementing field work with models
+(Morales-Castilla et al. 2015). Therefore, complementing fieldwork with models
 of species interactions is a promising avenue for understanding the complexity
-of the natural world(ref). While direct inference is often preferable, many
+of the natural world (ref). While direct inference is often preferable, many
 interactions are rare, infrequent, or impossible to detect through human
 observation. Model-based inference for ecological networks provides three
-important aspects to characterize species interactions, i) it establishes
-mechanistic relationships among species, ii) it estimates uncertainty given
-empirical observation, and iii) it provides an avenue to predict future
+important aspects to characterize species interactions, 1) it establishes
+mechanistic relationships among species, 2) it estimates uncertainty given
+empirical observation, and 3) it provides an avenue to predict future
 interactions. We follow Poisot (2015) and Cazelles (2016) in formulating our
-general conceptual model of species interactions (Fig. 1). There is a pool of
-true interactions, a metaweb, in which individuals of different species co-occur
-resulting in interactions among species. Any given observation we make on this
-metaweb is an attempt to recover the relationships among species using empirical
-data. Based on these data, model(s) are built to characterize and elucidate the
-relationships among species. Some models rely exclusively co-occurrence of
-individuals of different species at particular locations and may also include
-information on phylogenetic history and species traits. These models can than be
-used to predict the interactions among observed species, given an uncertainty
-due to the spatial and/or temporal stochasticity of the interactions. With these
-models it is also possible to define new network, based on the presence of a
-novel species in the assemblage. From these predictions, we can estimate
-structural network properties, such as connectance and nestedness, while
-simultaneously propagating our uncertainty throughout the analysis. In this
-paper, we reviewed the methods proposed to build models of species interactions,
-emphasizing each method's utility for predicting the position of a novel species
-in the network.
+general conceptual model of species interactions (Figure \ref{fig:concept}).
+There is a pool of true interactions, a metaweb, in which individuals of
+different species co-occur resulting in interactions among species. Any given
+observation we make on this metaweb is an attempt to recover the relationships
+among species using empirical data. Based on these data, model(s) are built to
+characterize and elucidate the relationships among species. Some models rely
+exclusively co-occurrence of individuals of different species at particular
+locations and may also include information on phylogenetic history and species
+traits. These models can than be used to predict the interactions among observed
+species, given an uncertainty due to the spatial and/or temporal stochasticity
+of the interactions. With these models it is also possible to define new
+networks, based on the presence of new species in the assemblage. From these
+predictions, we can estimate structural network properties such as connectance
+and nestedness, while simultaneously propagating our uncertainty throughout the
+analysis. In this paper, we reviewed the methods proposed to build models of
+species interactions, emphasizing each method's utility for predicting the
+position of a novel species in the network.
 
 From this conceptual framework, we can decide on a set of optimal criteria to
-compare existing approaches. While it is unlikely that any single method meets all
-our expectations, by defining a set of standards, we can more transparently
+compare existing approaches. While it is unlikely that any single method meets
+all our expectations, by defining a set of standards, we can more transparently
 judge the relative merits of each method to reach particular goals. The most
@@
 regarding our confidence in parameter estimates based on empirical data. An
-ideal method would also be able to make reasonable assumptions about the interdependence of
-observations, and be insensitive to low sampling effort. The
-mechanism determining the links among species should be clear and grounded
-in ecological theory such that we can use our method to make a prediction
-about the placement of a new species in the network based on a fitted model.
+ideal method would also be able to make reasonable assumptions about the
+interdependence of observations, and be insensitive to low sampling effort. The
+mechanism determining the links among species should be clear and grounded in
+ecological theory such that we can use our method to make a prediction about the
+placement of a new species in the network based on a fitted model.
 
-\subsection{Models of species
-interactions}\label{models-of-species-interactions}
+\section*{Methods}\label{methods}
 
-\subsubsection{Regression-based Models}\label{regression-based-models}
+One goal of this paper is to frame the different techniques used to model
+ecological networks within a conceptual framework that would allow ecologists to
+efficiently choose which technique to use with regard to various properties of
+the methods and the type of ecological questions they can answer. A way to
+assess the different properties of the modelling techniques is to compare them
+based on specific criteria. We defined three criteria, formulated as questions,
+that can be used to qualitatively compare the methods used to model ecological
+networks: 1) Can the method account for the uncertainty in species
+observations?; 2) Does the method provide ecological mechanisms to infer the
+probability of species interactions?; and 3) Can the method be used to predict
+an novel interaction?
 
-The simplest model of species interactions is
+In addition to qualitative comparison, we also provide a quantitative comparison
+of methods for inferring species interactions. From a quantitative perspective,
+we are interested in methods that: 1) accurately reflect the true underlying
+relationships among species, 2) are insensitive to sampling effort, and 3) make
+reasonable assumptions about the interdependence of species observations. To
+quantitatively compare the different methods, we used simulated and reference
+empirical data to establish benchmarks that will later be used to compare the
+accuracy and sensitivity of currently proposed methods.
 
-eq i.
+\subsection*{Modelling ecological networks - A diversity of methods}
+\label{models-of-species-interactions}
 
-\[Y_{i,j} = \rho_{i,j}\] Where the probability of interaction among species i
-and species j is a fixed value (\(\rho_{i,j}\)). Using empirical data, we might
-try to estimate this value by dividing the number of observations by the total
-number of observations for any pair of species. While this is straightforward,
-why might it be insufficient? Foremost, it contains no estimate of uncertainty.
-The estimate is given without any consideration for the variance due to
-sampling, spatial or temporal stochasticity. To add uncertainty, we draw our
-observations from a statistical distribution.
+When modelling ecological networks, there are different types of methods that
+have been proposed in the litterature. In this section, we presented the
+different techniques and group them based on their properties.
 
-eq ii.
+\subsubsection*{Co-occurrences models}
+\label{regression-based-models}
+%%% Regression-based models on co-occurrences subsubsection will probably have to be moved somewhere else to follow the structure of the Introduction where traits and phylogeny related methods are discussed firsts and co-occurrence methods are discussed later.
 
-\[ Y_{i,j} \sim Bernoulli(\rho_{i,j}) \]
+The simplest model of that can be designed to characterize the co-occurrence
+patterns between species $i$ and $j$ in matrix $Y$ can be defined as follow
 
-We can now estimate the probability of interaction, as well as the uncertainty
-based on sampling stochasticity. Yet this model is still unsatisfying. It is
-non-mechanistic, it depends wholly on species identity, and uses no ecological
-theory or inference. Therefore, it is difficult to ascribe meaning to a set of
-interactions, or understand how they might change in the future. We cannot use
-the inferences from data to make a prediction about a new species not in
-dataset.
+\begin{equation}
+  \label{eq:coOccSimple}
+  Y_{i,j} = \rho_{i,j}
+\end{equation}
 
-To make our model more useful, we might add a covariate, such that
+where the probability of interaction among species $i$ and $j$ is a fixed value
+($\rho_{i,j}$). Using empirical data, we might try to estimate this value by
+dividing the number of observations by the total number of observations for any
+pair of species. While this is straightforward, it is insufficient because it
+contains no estimate of uncertainty. The estimate is given without any
+consideration for the variance due to sampling, environmental, spatial or
+temporal stochasticity. This model is overly simple. By assuming that
+observations are randomly drawn from a statistical distribution with parameter
+$\rho_{i,j}$, we obtain a simple model that account for sampling stochasticity
+in the data.
 
-eq. iii
+\begin{equation}
+  \label{eq:coOccBern}
+  Y_{i,j} \sim \text{Bernoulli}(\rho_{i,j})
+\end{equation}
 
-\[ Y_{i,j} \sim Bernoulli(\rho_{i,j})\] \[ logit(\rho_{i,j}) = \alpha_{i,j} +
-\beta_{i,j} * x_{i,j} \]
+In equation \ref{eq:coOccBern}, we used a Bernoulli distribution, which assumes
+that $Y_{i,j}$ is either 0 or 1. If $Y_{i,j}$ includes other types of data (e.g.
+counts or weights) other distributions may be used instead. From equation
+\ref{eq:coOccBern}, we can now estimate the probability of interaction while
+accounting sampling stochasticity. Although this model is more informative than
+the one presented in equation \ref{eq:coOccSimple}, it is still unsatisfying
+because it depends wholly on species identity and uses no ecological theory or
+inference; in short it is not mechanistic. Therefore, it is difficult to ascribe
+meaning to the species association that could be highlighted by this model, or
+understand how they might change in the future. In addition, from this model, we
+cannot make a prediction about potential new species.
 
-In which a link between species i and j is a Bernoulli trial with a probability
-of success \(\rho_{i,j}\). The probability of success will vary based on some
-ecological principle (\(x_{i,j}\)), for example, the similarity in functional
-traits among species i and j:
+To make our model more ecologically relevant, we might add covariates $X_{i,j}$
+that influence the relationship among a pair of species
 
-eq. V
+\begin{equation}
+  \label{eq:coOccCov}
+  Y_{i,j} \sim \text{Bernoulli}(\rho_{i,j})
+  \text{logit}(\rho_{i,j}) = \alpha_{i,j}+\beta_{i,j}X_{i,j}
+\end{equation}
 
-\[x_{i,j}= f(Trait_i,Trait_j)\]
+where $\beta_{i,j}$ is a regression coefficient and $\alpha_{i,j}$ is the model
+intercept. In equation \ref{eq:coOccCov}, a link between species $i$ and $j$ is
+a Bernoulli trial with a probability of success $\rho_{i,j}$. The probability of
+success will vary based on some ecological principle ($x_{i,j}$) such as the
+similarity in functional traits among species $i$ and $j$:
+
+%%% GB: I am not sure if this equation is right... Need to check carefully.
+\begin{equation}
+  \label{eq:coOccTrait}
+  x_{i,j}= f(\text{Trait}_i,\text{Trait}_j)
+\end{equation}
+
+%%% ICI %%%
 
@@
 
-\subsubsection{Logic-based methods}\label{logic-based-methods}
+\subsubsection*{Logic-based methods}\label{logic-based-methods}
 
@@
 
-\subsubsection{Occurrence-based methods}\label{occurrence-based-methods}
+\subsubsection*{Occurrence-based methods}\label{occurrence-based-methods}
 
@@
 
-\subsubsection{Non-parametric methods}\label{non-parametric-methods}
+\subsubsection*{Non-parametric methods}\label{non-parametric-methods}
 
 The above methods are model based, that is, they largely use parametric
-distributions to estimate the liklihood of the data given estimated model
+distributions to estimate the likelihood of the data given estimated model
 parameters. An alternative approach is to use multivariate statistics to create
@@
 
-\section{Methods}\label{methods}
-
-Our goal is provide a conceptual framework for the broad array of potential
-methods described above. Our three criteria for qualitative comparison are: 1)
-Can the method account for the uncertainty in species observations?, 2) Does it
-provide an ecological mechanism to infer the probability of species
-interactions, and 3) Can it be used to predict an novel interaction? In addition
-to qualitative comparison, we provide a quantitative analysis of methods for
-inferring species interactions. From a quantitative perspective, we are
-interested in methods which accurately reflect true relationships among data,
-are relatively insensitive to sampling effort, and make reasonable assumptions
-about the interdependence of species observations. We use a combination of
-simulated and empirical data to establish a benchmark dataset to compare the
-accuracy and sensitivity of currently proposed methods.
-
-\section{Quantitative comparisons}\label{quantitative-comparisons}
+\subsection*{Quantitative comparisons}\label{quantitative-comparisons}
 
@@
 
-\section{Results}\label{results}
+\section*{Results}\label{results}
 
-\section{Discussion}\label{discussion}
+\section*{Discussion}\label{discussion}
 
-\subsection{Perspectives}\label{perspectives}
+\subsection*{Perspectives}\label{perspectives}
 
@@
 
-\subsection{On the virtue of prediction of species
+\subsection*{On the virtue of prediction of species
 interactions.}\label{on-the-virtue-of-prediction-of-species-interactions.}
@@
 
-\section{Literature cited}\label{literature-cited}
+\section*{Literature cited}\label{literature-cited}
 
@@
 
-\section{Boxes}\label{potential-boxes}
+\section*{Boxes}\label{potential-boxes}
 
-\subsection{Soil ecology}\label{soil-ecology}
+\subsection*{Soil ecology}\label{soil-ecology}
 
@@
 
-\subsection{Climate changes and alien
-species}\label{climate-changes-and-alien-species}
+\subsection*{Climate changes and alien species}
+\label{climate-changes-and-alien-species}
 
-\subsection{Community/evolution ecology}\label{communityevolution-ecology}
+\subsection*{Community/evolution ecology}
+\label{communityevolution-ecology}
 

### 2017-02-10 Data/data_explnation.Rmd
Added cleaned data (more to come)

@@
 __Interactions:__ This include `transect`, `visitation` and `out` data.
-Data can be used pooled (2067 interactions (5732 visits) general question abouts trait matching) or splited by site (`Site_ID`) and `Round`.
+Data can be used pooled (1472 interactions ; 5706 visits) for general question abouts trait matching or splited by site (`Site_ID`) and `Round`.
 
@@
 ```
-__NOTE:__ Quite a lot of NA's still.
+__NOTE:__ Most IT (intertegular length) for bees are actually measured.
+For floral visitors without wings we measured torax width.
+When species is unknown (Genus sp.) we calculated the mean of the genus or family.
+Tongue length is measured only for a few individuals.
+Bees tongue length is estimated following Cariveau et al 2016.
+For families without an actual tongue the value is set to zero.
+For other species, tongue length is estimated to covary with allometry using same equation as for colletidaes (the shortest tongue bee family).
 
@@
 - We are not sure which trait are driving interactions (quite generalized system)
-- Need to inegrate traits across different orders (e.g. tongue lengths can be set to zero for coleoptera)
 
@@
 ```{r}
+head(int)
+unique(int[,3:4]) #why NA NA
+unique(int$Plant_gen_sp[which(!int$Plant_gen_sp %in% plant_tr$Plant_gen_sp)]) #FUCK!
+unique(int$Pollinator_gen_sp[which(!int$Pollinator_gen_sp %in% pol_tr$Pollinator_gen_sp)])
 temp <- merge(int, pol_tr, by.x = "Pollinator_gen_sp", by.y = "Pollinator_gen_sp", all.x = TRUE)
@@
 pooled_data <- temp2
-head(pooled_data)
+head(pooled_data, 20)
 ```
@@
 ```
-#' @param idObs A data.frame which is mandatory and will help to check consistency and prevent errors among unique identifiers of each alienData arguments.
-The first column (idSite) contains unique identifier of where the observation was made.
-The second column (idTime) is not mandatory and contains temporal information: an unique identifier at the time the sample has been taken (needed for timeseries analysis).
-The third column (idSpcies) is an unique identifier of the species sampled at time (idTime) and location (idSite).
-The fourth column is an unique identifier of individu of species (idSp) observed at time (idTime) and location (idSite).
-#' @param interactPair A data.frame which contains interaction at the finest level (individus or species).
-The first two columns are idFrom and idTo and determine the sens of the interaction.
-idFrom and idTo are unique identifier of species or individu documented in the idObs data.frame.
-Finaly, the thrid column is the strength of the interaction (Please see details).
-#' @param coOcc A square symmetric matrix of 0s and 1s that define co-occurence patterns among pairs of species.
-If this matrix is not provided, the co-occurence matrix is derived from the coAbund matrix else the interactSp matrix (see return section).
-#' @param coAbund A square symmetric matrix that includes any types of values, defining co-abundance patterns among pairs of species. TODO: Not implemented yet.
-#' @param siteEnv A matrix or a data.frame where each column is a descriptor of the sites. TODO: siteEnv should cover the possibility that environmental variables could be taken at several times.
-#' @param traitSp A matrix or a data.frame where each column is a trait characterizing all species. The first column is a unique identifier of the species documented in idObs data.frame.
-#' @param traitInd A matrix or a data.frame where each column is a trait characterizing an individual. The first column is a unique identifier of the individu documented in idObs data.frame.
-#' @param phylo A square symmetric matrix describing the phylogenetic relationships between pairs of all species (see details). TODO: Not implemented yet.
-#' @param scaleSiteEnv Logical. Whether the columns of X should be centred and divided by the standard deviation. Default is TRUE.
-#' @param scaleTrait Logical. Whether the rows of Tr should be centred and divided by the standard deviation. Default is TRUE.
-#' @param interceptSiteEnv Logical. Whether a column of 1s should be added to X. Default is TRUE.
-#' @param interceptTrait Logical. Whether a row of 1s should be added to Tr. Default is TRUE.
-#'
+#' @param idObs A data.frame which is mandatory and will help to check
+consistency and prevent errors among unique identifiers of each alienData
+arguments. The first column (idSite) contains unique identifier of where the
+observation was made. The second column (idTime) is not mandatory and contains
+temporal information: an unique identifier at the time the sample has been taken
+(needed for timeseries analysis). The third column (idSpcies) is an unique
+identifier of the species sampled at time (idTime) and location (idSite). The
+fourth column is an unique identifier of individu of species (idSp) observed at
+time (idTime) and location (idSite). 
+#' @param interactPair A data.frame which
+contains interaction at the finest level (individus or species). The first two
+columns are idFrom and idTo and determine the sens of the interaction. idFrom
+and idTo are unique identifier of species or individu documented in the idObs
+data.frame. Finaly, the thrid column is the strength of the interaction (Please
+see details). 
+#' @param coOcc A square symmetric matrix of 0s and 1s that define
+co-occurence patterns among pairs of species. If this matrix is not provided,
+the co-occurence matrix is derived from the coAbund matrix else the interactSp
+matrix (see return section). 
+#' @param coAbund A square symmetric matrix that
+includes any types of values, defining co-abundance patterns among pairs of
+species. TODO: Not implemented yet. 
+#' @param siteEnv A matrix or a data.frame
+where each column is a descriptor of the sites. TODO: siteEnv should cover the
+possibility that environmental variables could be taken at several times. 
+#' @param traitSp A matrix or a data.frame where each column is a trait
+characterizing all species. The first column is a unique identifier of the
+species documented in idObs data.frame. 
+#' @param traitInd A matrix or a data.frame where each column is a trait characterizing an individual. The first
+column is a unique identifier of the individu documented in idObs data.frame. 
+#' @param phylo A square symmetric matrix describing the phylogenetic relationships
+between pairs of all species (see details). TODO: Not implemented yet. 
+#' @param scaleSiteEnv Logical. Whether the columns of X should be centred and divided by
+the standard deviation. Default is TRUE. 
+#' @param scaleTrait Logical. Whether the rows of Tr should be centred and divided by the standard deviation. Default
+is TRUE. 
+#' @param interceptSiteEnv Logical. Whether a column of 1s should be
+added to X. Default is TRUE. 
+#' @param interceptTrait Logical. Whether a row of
+1s should be added to Tr. Default is TRUE. 
 #' @details 
-#'
-#' The strength of the interactions defined in the third column of \code{interactPair} can be a 0 if no direct interaction has been observed (defined as true absence of interaction) or any numerical value.
-Undocumented interactions among species or individus will be assumed as NA by default.
+#'The strength of the interactions defined in the third column of \code{interactPair} can be a 0
+if no direct interaction has been observed (defined as true absence of
+interaction) or any numerical value. Undocumented interactions among species or
+individus will be assumed as NA by default.
 ```

### 2017-02-14 Data/data_explnation.Rmd
Data completely clean and complete.

@@
 __Interactions:__ This include `transect`, `visitation` and `out` data.
-Data can be used pooled (1472 interactions ; 5706 visits) for general question abouts trait matching or splited by site (`Site_ID`) and `Round`.
+Data can be used pooled (1469 interactions ; 5706 visits) for general question abouts trait matching or splited by site (`Site_ID`) and `Round`.
 
@@
 - not all species are taxonomically resolved. What to do with specimens we ony know Genus?
+As for now:
+  . sp means we only know the genus
+  . morpho1,2,3 Unidentified species, but we know is a unique species diferent from anything else.
+  . Family NA- means we only know the family. 
 - Low sample size when you split per site and round
@@
 ```
-__NOTE:__ Most IT (intertegular length) for bees are actually measured.
+__NOTE:__ Most IT (intertegular length) for bees are actual measures.
 For floral visitors without wings we measured torax width.
@@
 Tongue length is measured only for a few individuals.
-Bees tongue length is estimated following Cariveau et al 2016.
+Bees tongue length for most species is estimated following Cariveau et al 2016.
 For families without an actual tongue the value is set to zero.
 For other species, tongue length is estimated to covary with allometry using same equation as for colletidaes (the shortest tongue bee family).
-
-Challenges:  
-- We are not sure which trait are driving interactions (quite generalized system)
 
@@
 
+Challenges:  
+- We are not sure which trait are driving interactions (quite generalized system)
+
 __independent plant flower abundance__: Ca be combined by Site_ID and Round
@@
 __Phylogenies__ Need to do.
+
+```{r}
+#Nacho this is the final list of species
+pol_tr$Pollinator_gen_sp #277 entries. 
+#sp means we don't know the species. Ignore for phylogeny and use a random species in this genus
+#morpho1,2,3 Unidentified, but we know is a unique species. Idem. Ignore for phylogenies and use a species in genus as replacement?
+#Family NA- means we only know the family. It will be intersting to make sure those families are represented, e.g. by adding a ranom species?
+plant_tr$Plant_gen_sp
+```
+
 - Plants: Phylomatic?
@@
 
-Question: should we use the pooled data for the package?
 ```{r}
 head(int)
-unique(int[,3:4]) #why NA NA
-unique(int$Plant_gen_sp[which(!int$Plant_gen_sp %in% plant_tr$Plant_gen_sp)]) #FUCK!
+unique(int[,3:4])
+unique(int$Plant_gen_sp[which(!int$Plant_gen_sp %in% plant_tr$Plant_gen_sp)]) #ok
 unique(int$Pollinator_gen_sp[which(!int$Pollinator_gen_sp %in% pol_tr$Pollinator_gen_sp)])
@@
 # Create dir for formatted data
-dir.create('./Data/alienData', showWarnings = FALSE)
+#dir.create('./Data/alienData', showWarnings = FALSE)
 ```
@@
 # Create unique IDs for plant individus
-int$idPlant <- seq(nrow(int)+1,nrow(int)+nrow(int),1)
+int$idPlant <- seq(nrow(int)+1,nrow(int)+nrow(int),1) #technically, some individuus will be repeated, although there is not way to know. Good for now
 
 # subset pollinator and plant with their own ids
-pol <- unique(int[,c("Site_ID", "Round","Pollinator_gen_sp","idPol")])
+pol <- unique(int[,c("Site_ID", "Round","Pollinator_gen_sp","idPol")]) #The unique is unececsary here, right?
 colnames(pol) <- c("idSite","idTime","idSp","idInd")
-plant <- unique(int[,c("Site_ID", "Round","Plant_gen_sp","idPlant")])
+plant <- unique(int[,c("Site_ID", "Round","Plant_gen_sp","idPlant")]) #idem
 colnames(plant) <- c("idSite","idTime","idSp","idInd")
@@
 idObs <- rbind(pol,plant)
+#head(idObs)
 
@@
 names(traitInd)[3] <- 'value'
+#head(traitInd)
 ```
@@
 library(reshape2)
-
 traitSpPlant <- melt(plant_tr[,-2],id=c("Plant_gen_sp"),value.name="value",variable.name="traitName")
@@
 
-traitSpPol <- melt(pol_tr[,-4],id=c("Pollinator_gen_sp"),value.name="value",variable.name="traitName")
+#head(pol_tr)
+traitSpPol <- melt(pol_tr[,c(1,2,5)],id=c("Pollinator_gen_sp"),value.name="value",variable.name="traitName") #I also remove order
 names(traitSpPol)[1] <- "idSp"
@@
 # Some entries in traitSp (merge of pol_tr and plant_tr) are documenting species which are not present in `ìnt`, the interaction pairwise data.frame.
+#IB: This is ok! I collect traits for some species we do not have interactions.
 removeSpecies <- levels(traitSp$idSp)[which(!levels(traitSp$idSp) %in% levels(idObs$idSp))]

### 2017-02-14 Data/data_explnation.Rmd
Play a bit with trait-matching functions

@@
 ```
+
+```{r}
+#IB plays with trait matching
+head(pooled_data, 20)
+
+#simple match
+scatter.smooth(pooled_data$nectar_tube_depth_mm ~ pooled_data$tongue_length)
+temp <- pooled_data$tongue_length - pooled_data$nectar_tube_depth_mm 
+plot(temp) #A few outlyers, including negative values.
+
+pooled_data[which(temp < 0),c(1,2,12,17)]
+#Clearly there is something with Anchusa azurea: Robing? Exerted pollen?
+#Andryala integrifolia: idem
+#Armeria velutina: Very open tube? idem with convolvulus and equium, rosmarinus...
+#one interaction of Astragalus lusitanicus	Oxythyrea funesta, check?
+pooled_data[which(temp > 10),c(1,2,12,17)]
+#some buterflies
+#lots of long tongued bees visiting open flowers (cistus, asphodelus) This is real and suggests a barrier model.
+
+#complex match
+temp <- ifelse(pooled_data$IT > pooled_data$nectar_tube_diameter_mm,
+               pooled_data$tongue_length - pooled_data$nectar_tube_depth_mm,
+               0)
+plot(temp) #A few outlyers, including negative values.
+
+pooled_data[which(temp < 0),c(1,2,12,17)]
+```

### 2017-02-15 Data/phylogenies.Rmd
Assembling phylogenies for empirical data

---
title: "Obtaining and assembling phylogenies"
author: "Ignacio Morales-Castilla"
date: "February 14, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Reading data
```{r reading data in}
## remove objects
rm(list=ls())
options(stringsAsFactors = FALSE)

setwd("~/GitHub/Montreal_Workshop/Data")
int <- read.csv("interactions.csv", h = T)
#head(int)

plants<-sub(" ","_",unique(int$Plant_gen_sp))
insects<-sub(" ","_",unique(int$Pollinator_gen_sp))

```

## Loading plant phylogeny and matching taxonomy

Out of the initial list of 56 plant species, 33 are in the phylogeny. For 21 of the remaining 23 species, we can assign the phylogenetic placement of a sister species (within the same genus). Two species can't be placed in this preliminary phylogeny (Mirtus_communis & Halimium_halimifolium). We could insert them as polytomies.

```{r phylo plants, echo=FALSE}
library(phytools)
plant.phylo<-read.tree("Plants_phylo.tre")

# plants not in phylogeny
not.in.phylo<-plants[which(!plants%in%plant.phylo$tip.label)]

# check if genus of these plants is in phylogeny
genus.plants<-sapply(strsplit(plants, "_"), "[", 1)
genus.phylo<-sapply(strsplit(plant.phylo$tip.label, "_"), "[", 1)
genus.not.in.phylo<-genus.plants[which(!genus.plants%in%genus.phylo)]

# retrieve sister species for those species not in phylo
rep.each.genus<-table(genus.plants)
new.names<-array(NA,length(not.in.phylo))
for (i in 1:length(not.in.phylo)){
  #print(i)
  sps.i<-not.in.phylo[i]
  genus.i<-strsplit(sps.i, "_")[[1]][1]
  length(which(genus.phylo!=genus.i))

  if(!genus.i%in%genus.not.in.phylo){
  if(length(which(genus.phylo==genus.i))>1){
  phylo.genus.i<-drop.tip(plant.phylo,which(genus.phylo!=genus.i))
  phylo.genus.i<-drop.tip(phylo.genus.i,which(phylo.genus.i$tip.label%in%plants))

  #if(rep.each.genus[genus.i]<length(phylo.genus.i$tip.label)){
    new.names[i]<-sample(phylo.genus.i$tip.label,1)
  } else {
    new.names[i]<-plant.phylo$tip.label[which(genus.phylo==genus.i)]
  }
  } 
}

# check that there are not repeated species
#sum(which(new.names%in%plants))

# replace names
plants.new<-plants
plants.new[which(!plants%in%plant.phylo$tip.label)]<-new.names
plants.new[is.na(plants.new)]<-"Mirtus_communis"

```

## The plant phylogeny
Phylogenetic tree with 54 plant species 
```{r phylo plants, echo=FALSE}
# prune and obtain final plant phylogeny:
plant.phylo.int<-drop.tip(plant.phylo,which(!plant.phylo$tip.label%in%plants.new))

# compare and get vector with original names in same order as names in phylogeny
final.sps.list<-unique(plants.new[which(plants.new%in%plant.phylo$tip.label)])
names.compar<-data.frame(names.phylo=final.sps.list,order.phylo=as.numeric(match(final.sps.list,plant.phylo.int$tip.label)),names.original=plants[-c(34,53)])

# reordering to match phylogeny
names.compar<-names.compar[with(names.compar,order(order.phylo)),]

# re-assign names to phylogeny
plant.phylo.int$tip.label<-names.compar$names.original
plot(plant.phylo.int,cex=0.6)

```

## Repeating for Pollinator phylogeny
This one is going to be trickier. Only 16 sps of our list of arthropods are included in TTOL. The following genera are not in the phylogeny. Should we go to familial level?

```{r phylo insects, echo=FALSE}
all.phylo<-read.tree("TTOL_arthropods.nwk")

# pollinators not in phylogeny
not.in.phylo<-insects[which(!insects%in%all.phylo$tip.label)]

# check if genus of these plants is in phylogeny
genus.insects<-sapply(strsplit(insects, "_"), "[", 1)
genus.phylo<-sapply(strsplit(all.phylo$tip.label, "_"), "[", 1)
genus.not.in.phylo<-unique(genus.insects[which(!genus.insects%in%genus.phylo)])
genus.not.in.phylo

```

### 2017-02-17 Data/phylogenies.Rmd
just play a bit with this.

@@
 setwd("~/GitHub/Montreal_Workshop/Data")
+#setwd("~/Documents/github/local/Montreal_Workshop/Data")
 int <- read.csv("interactions.csv", h = T)
@@
 We could insert them as polytomies.
+
+IB: Halimium_conmutatum, is represented, can we use it as sister species?
 
@@
 # re-assign names to phylogeny
-plant.phylo.int$tip.label<-names.compar$names.original
+plant.phylo.int$tip.label<-as.character(names.compar$names.original)
 plot(plant.phylo.int,cex=0.6)
@@
 genus.insects<-sapply(strsplit(insects, "_"), "[", 1)
+species.insects<-sapply(strsplit(insects, "_"), "[", 2)
+
 genus.phylo<-sapply(strsplit(all.phylo$tip.label, "_"), "[", 1)
@@
 
+#check how many species per genus there are
+temp <- colSums(table(data.frame(sp = species.insects, gen = genus.insects)))
+summary(temp)
+
 ```

### 2017-03-02 ms/ms_interactions.tex
Reorganized the co-occurrence models section

@@
 When modelling ecological networks, there are different types of methods that
-have been proposed in the litterature. In this section, we presented the
+have been proposed in the literature. In this section, we presented the
 different techniques and group them based on their properties.
@@
 \subsubsection*{Co-occurrences models}
-\label{regression-based-models}
-%%% Regression-based models on co-occurrences subsubsection will probably have to be moved somewhere else to follow the structure of the Introduction where traits and phylogeny related methods are discussed firsts and co-occurrence methods are discussed later.
+\label{Co-occurrences models}
 
-The simplest model of that can be designed to characterize the co-occurrence
-patterns between species $i$ and $j$ in matrix $Y$ can be defined as follow
+% Regression-based models on co-occurrences subsubsection will probably have
+% to be moved somewhere else to follow the structure of the Introduction where
+% traits and phylogeny related methods are discussed firsts and co-occurrence
+% methods are discussed later.
 
-\begin{equation}
-  \label{eq:coOccSimple}
-  Y_{i,j} = \rho_{i,j}
-\end{equation}
-
-where the probability of interaction among species $i$ and $j$ is a fixed value
-($\rho_{i,j}$). Using empirical data, we might try to estimate this value by
-dividing the number of observations by the total number of observations for any
-pair of species. While this is straightforward, it is insufficient because it
-contains no estimate of uncertainty. The estimate is given without any
-consideration for the variance due to sampling, environmental, spatial or
-temporal stochasticity. This model is overly simple. By assuming that
-observations are randomly drawn from a statistical distribution with parameter
-$\rho_{i,j}$, we obtain a simple model that account for sampling stochasticity
-in the data.
+Co-occurence models are designed to characterize species co-occurrence directly.
+In these models, the co-occurrences $Z_{ij}$ we try to characterize describes
+the presence (1) or absence (0) of a trophic link between species $i$ and $j$.
+In the context of co-occurrence models,  a way to formalize the relationship
+between co-occurrences and traits is as follow
 
@@
   \label{eq:coOccBern}
-  Y_{i,j} \sim \text{Bernoulli}(\rho_{i,j})
+  Z_{ij} \sim \text{Bernoulli}(\rho_{ij})
 \end{equation}
 
-In equation \ref{eq:coOccBern}, we used a Bernoulli distribution, which assumes
-that $Y_{i,j}$ is either 0 or 1. If $Y_{i,j}$ includes other types of data (e.g.
-counts or weights) other distributions may be used instead. From equation
-\ref{eq:coOccBern}, we can now estimate the probability of interaction while
-accounting sampling stochasticity. Although this model is more informative than
-the one presented in equation \ref{eq:coOccSimple}, it is still unsatisfying
-because it depends wholly on species identity and uses no ecological theory or
-inference; in short it is not mechanistic. Therefore, it is difficult to ascribe
-meaning to the species association that could be highlighted by this model, or
-understand how they might change in the future. In addition, from this model, we
-cannot make a prediction about potential new species.
+where $\rho_{ij}$ is the probability of interaction among species $i$ and $j$.
+In equation \ref{eq:coOccBern}, we used a Bernoulli distribution because we
+assumed that $Z_{ij}$ is binary. However, if $Z_{ij}$ is characterized by other
+types of data (e.g. counts or weights) other distributions may be used instead.
+From equation \ref{eq:coOccBern}, we can estimate the probability of interaction
+while accounting for sampling stochasticity. However, because this model depends
+solely on species identity it may lead to misinterpretation of the mechanisms
+that characterize the co-occurrences.
 
-To make our model more ecologically relevant, we might add covariates $X_{i,j}$
-that influence the relationship among a pair of species
+A way to solve this problem is to add a layer of complexity to the model in
+equation \ref{eq:coOccBern} by including traits. To do this, the probability of
+interaction between species $i$ and $j$ defined by $\rho_{ij}$ can depend on the
+similarity in functional traits for species $i$ and $j$
 
 \begin{equation}
-  \label{eq:coOccCov}
-  Y_{i,j} \sim \text{Bernoulli}(\rho_{i,j})
-  \text{logit}(\rho_{i,j}) = \alpha_{i,j}+\beta_{i,j}X_{i,j}
+  \label{eq:coOccTraits}
+  \text{logit}(\rho_{ij}) = \beta_{ij}X_{ij}
 \end{equation}
 
-where $\beta_{i,j}$ is a regression coefficient and $\alpha_{i,j}$ is the model
-intercept. In equation \ref{eq:coOccCov}, a link between species $i$ and $j$ is
-a Bernoulli trial with a probability of success $\rho_{i,j}$. The probability of
-success will vary based on some ecological principle ($x_{i,j}$) such as the
-similarity in functional traits among species $i$ and $j$:
-
-%%% GB: I am not sure if this equation is right... Need to check carefully.
-\begin{equation}
-  \label{eq:coOccTrait}
-  x_{i,j}= f(\text{Trait}_i,\text{Trait}_j)
-\end{equation}
-
-%%% ICI %%%
+where $\beta_{ij}$ is a regression coefficient and $X_{ij}$ is a covariates that
+influence the relationship among a pair of species. For example, $X_{ij}$ could
+be a trait-matching function between species $i$ and $j$. Note that a model
+intercept is accounted for in equation \ref{eq:coOccTraits} by assuming that $X$
+includes a column of 1s. In equation \ref{eq:coOccTraits}, we used a logit link
+function to ensure that $\rho$ ranges between 0 and 1, however, it is also
+possible to use other link functions such as a probit or a complementary log-log
+link function.
 
 A variety of trait-matching functions may be biologically plausible. For
-example, the absolute value of the difference between traits (e.g. \(|trait_i -
-trait_j|\)) or a binary difference (e.g. \(trait_i > trait_j\)), could be used
-to predict which species may interact. Alternatively, the covariate may
-represent the phylogenetic relatedness of species, or their relative response to
-a given environmental variable.In some cases, we cannot provide a simple
-trait-matching function (eg V), but must use a statistical tool to infer the
-best possible combination of traits among co-occurring species that predicts
-interactions. Machine learning tools provide an ideal compliment to traditional
-linear models when we cannot anticipate the combinations of traits that can
-influence species interaction rates. Unsupervised learning tools, such as random
-forests, can give us the associations among species given a set of complex
-traits. Regardless of the exact form of regression, this framework satisfies two
-important goals: it allows for estimates of uncertainty in species interactions,
-and it provides a mechanism shaping the probability of an interaction.
+example, the absolute value of the difference between traits or a binary
+difference, can be used to predict which species may interact. Alternatively,
+the covariate may also represent phylogenetic relatedness, or their relative
+response to a given environmental variable.
 
-So far, we have only captured the uncertainty in the process model, the inherent
-temporal or spatial stochasticity in our ecological process of interest (Hooten
-and Hobbs). We may need to also account for the uncertainty in the observation
-model, the ability to detect interactions given that they occur. The separation
-of the observation and process model translates our ecological mechanism into
-specific predictions about empirical data. For binary networks, one
-straightforward way to account for the observation of interactions is to model
-the detectability of network interactions.
+In the litterature, \#\#\#REF(Ben?)\#\#\# also proposed to account for
+observation uncertainty; to account for the ability to detect interactions given
+that they really occurred. In this formulation, the idea is to try to seperate
+the error resulting from the observation and the processes defining the
+co-occurrences. As such, we can define a model that account for observation uncertainty and ecological processes through the following equations
 
-eq iv.
+\begin{align}
+  Z_{ijk} \sim \text{Bernoulli}(\phi_{ij})\\
+  \phi_{ij} = \omega_{ij} \times \Omega_{ij}\\
+  \Omega_{ij} \sim \text{Bernoulli}(\rho_{i,j})\\
+  logit(\rho_{ij}) = \beta_{ij} X_{ij}
+\end{align}
 
-\[ Y_{i,j,k} \sim Bernoulli(\phi_{i,j})\] \[ \phi_{i,j} = \omega_{i,j} *
-z_{i,j}\] \[z_{i,j} \sim Bernoulii(\rho_{i,j})\] \[ logit(\rho_{i,j}) =
-\alpha_{i,j} + \beta_{i,j} * x \]
+where the observation at time $k$ of a co-occurrence between species $i$ and $j$
+is a Bernoulli trial with a probability $\phi_{ij}$, which is the outcome of the
+detectability of an interaction ($\omega_{ij}$) and the latent state
+($\Omega_{ij}$). The latent state $\Omega_{ij}$ is the true, but unobserved,
+existence of a co-occurrence, as predicted by the ecological mechanism of
+interest. The ability to assess this latent state depends on the detectability
+of an interaction ($\omega_{i,j}$).
 
-Where the observation of a link between species i and species j at time k is a
-Bernoulli trial with a probability \(\phi_{i,j}\), which is the outcome of the
-detectability of an interaction (\(\omega_{i,j}\)) and the latent state
-(\(z_{i,j}\)). This latent state is the true, but unobserved, existence of a
-link, as predicted by our ecological mechanism of interest (see eq IV). Our
-ability to detect this latent state depends on the detectability of an
-interaction (\(\omega_{i,j}\)). The benefit of this approach is that
-differentiates the probability of detection from the probability of occurrence.
-Importantly, because we have explicitly defined some temporal window (k),
-interactions with different levels of sampling effort can be directly compared.
+\subsubsection*{fourth-corner approaches}
+\label{fourth-corner approaches}
+
+A way to assess
+
+An alternative approach is to use multivariate statistics to create association
+matrices. One such method is called fourth-corner analysis, which is a type of
+three table ordination (sometimes called `RLQ' analysis) (Doledec 1996). In
+bipartite networks, fourth-corner analysis relates the matrix of species
+interactions to a matrix of traits from one level (e.g.~plants) to another level
+(e.g.~pollinators) via the matrix of species innteractions (Dray \& Legendre,
+2008). This produces a correlation matrix, which is then compared to a null
+expectation through matrix randomization. Randomization is a familiar tool in
+network ecology, in which permuatation rules are used to compare the observed
+test statistic (e.g the correlation among species occurrences) to a null
+distribution based on some matrix constraint. Common constraints include the
+mantainance of row sums, column sums, or other all marginal values, such that
+only the identity of partners are switched. After each permutation, the test
+statistic is recalulcating, yielding a null distribution. By comparing whether
+the observed test statistic falls within the upper or lower 5th quantile of the
+null distribution, studies distinguish whether their results could have been
+generated at random given their sampling effort.
+
+\end{document}
+
+%%% I think these bits of text would fit better in the discussion
+%% The text in 1) were all from the co-occurrence model section.
+% 1.1)
+In some cases, we cannot provide a simple trait-matching function but must use a
+statistical tool to infer the best possible combination of traits among
+co-occurring species that predicts interactions. Machine learning tools provide
+an ideal compliment to traditional linear models when we cannot anticipate the
+combinations of traits that can influence species interaction rates.
+Unsupervised learning tools, such as random forests, can give us the
+associations among species given a set of complex traits. Regardless of the
+exact form of regression, this framework satisfies two important goals: it
+allows for estimates of uncertainty in species interactions, and it provides a
+mechanism shaping the probability of an interaction.
+
+% 1.2)
+The benefit of this approach is that differentiates the probability of detection
+from the probability of occurrence. Importantly, because we have explicitly
+defined some temporal window (k), interactions with different levels of sampling
+effort can be directly compared.
 
@@
 {[}Ben Weinstein: For someone to fill in.{]}
-
-\subsubsection*{Non-parametric methods}\label{non-parametric-methods}
-
-The above methods are model based, that is, they largely use parametric
-distributions to estimate the likelihood of the data given estimated model
-parameters. An alternative approach is to use multivariate statistics to create
-association matrices. One such method is called fourth-corner analysis, which is
-a type of three table ordination (sometimes called `RLQ' analysis) (Doledec
-1996). In bipartite networks, fourth-corner analysis relates the matrix of
-species interactions to a matrix of traits from one level (e.g.~plants) to
-another level (e.g.~pollinators) via the matrix of species innteractions (Dray \&
-Legendre, 2008). This produces a correlation matrix, which is then compared to a
-null expectation through matrix randomization. Randomization is a familiar tool
-in network ecology, in which permuatation rules are used to compare the observed
-test statistic (e.g the correlation among species occurrences) to a null
-distribution based on some matrix constraint. Common constraints include the
-mantainance of row sums, column sums, or other all marginal values, such that
-only the identity of partners are switched. After each permutation, the test
-statistic is recalulcating, yielding a null distribution. By comparing whether
-the observed test statistic falls within the upper or lower 5th quantile of the
-null distribution, studies distinguish whether their results could have been
-generated at random given their sampling effort.
 

### 2017-03-09 ms/ms_interactions.tex
Added some text in the fourth corner section

@@
 
-\subsubsection*{fourth-corner approaches}
-\label{fourth-corner approaches}
+\subsubsection*{fourth-corner modelling}
+\label{fourth-corner modelling}
 
-A way to assess
+Fourth corner models originated from the work of \cite{doledec_matching_1996},
+who proposed a way to perform an ordination called the `RLQ analysis', and
+\cite{legendre_relating_1997}, who proposed a statistical  procedure to tests
+the relationships among three data tables called the `fourth corner problem'.
+Originally, the three data tables described a set of species found at different
+locations, a set of environmental variables found at the same locations and a
+set of species traits related associated to each species (\#\#\# Figure
+conceptual 4th corner \#\#\#). Following these two seminal papers, a series of
+developments have been proposed extending and generalizing and linking the `RLQ
+analysis' and the `fourth corner problem' \cite{dray_testing_2008,
+dray_combining_2014}.
 
-An alternative approach is to use multivariate statistics to create association
-matrices. One such method is called fourth-corner analysis, which is a type of
-three table ordination (sometimes called `RLQ' analysis) (Doledec 1996). In
-bipartite networks, fourth-corner analysis relates the matrix of species
-interactions to a matrix of traits from one level (e.g.~plants) to another level
-(e.g.~pollinators) via the matrix of species innteractions (Dray \& Legendre,
-2008). This produces a correlation matrix, which is then compared to a null
-expectation through matrix randomization. Randomization is a familiar tool in
-network ecology, in which permuatation rules are used to compare the observed
-test statistic (e.g the correlation among species occurrences) to a null
-distribution based on some matrix constraint. Common constraints include the
-mantainance of row sums, column sums, or other all marginal values, such that
-only the identity of partners are switched. After each permutation, the test
-statistic is recalulcating, yielding a null distribution. By comparing whether
-the observed test statistic falls within the upper or lower 5th quantile of the
-null distribution, studies distinguish whether their results could have been
-generated at random given their sampling effort.
+In the context of ecological interactions, a different set of three matrices are
+used. A matrix of species interactions is used as a basis which is related with
+a matrix of traits for both side of the interacting species (\#\#\# Figure
+conceptual 4th corner \#\#\#). These different developments are interesting to
+evaluate and test the relationship among species, however they are not designed
+to make predictions. To solve this issue, \cite{nygaard_new_2004},
+\cite{pollock_role_2012}, \cite{jamil_selecting_2013},
+\cite{brown_fourth-corner_2014} have independently proposed developments to make
+predictions when dealing with three related matrices. In essence, these new
+developments rely on hierarchical models \cite{gelman_data_2007}.
+
+%%% ICI %%%
+%%% ICI %%%
+%%% ICI %%%
+%%% ICI %%%
+\section*{Literature cited}\label{literature-cited}
+
+\bibliography{referencesNew}
+
+\input{ms_fig.tex}
+
+\input{ms_tables.tex}
 
@@
 
-\input{ms_fig.tex}
-
-\input{ms_tables.tex}
-
 \end{document}

### 2017-07-24 Data/alien_reshaping.Rmd
alienData v2 - new structure after some corrections inherents to methods

--- 2017-02-14 Data/data_explnation.Rmd
@@
----
-title: "Data description"
-author: "I. Bartomeus"
-date: "12/1/2016"
-output: github_document
----
-
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-```
-
-Data collected in 2015 across 16 sites. Each site was visited 7 times. Each time we recorded three data types:
-
-- All interactions in 100 m transects (`transects`)
-- Focal interactions to selected plants (`focal`)
-- Oportunistic new interactions not detected by the last two methods (`out`)
-
-there are three datasets so far:
-
-__Interactions:__ This include `transect`, `visitation` and `out` data.
-Data can be used pooled (1469 interactions ; 5706 visits) for general question abouts trait matching or splited by site (`Site_ID`) and `Round`.
-
-```{r}
-int <- read.csv("interactions.csv", h = T)
-head(int)
-nrow(int)
-sum(int$Frequency, na.rm = TRUE)
-```
-Challenges:  
-- long tail of singletones  
-- not all species are taxonomically resolved. What to do with specimens we ony know Genus?
-As for now:
-  . sp means we only know the genus
-  . morpho1,2,3 Unidentified species, but we know is a unique species diferent from anything else.
-  . Family NA- means we only know the family. 
-- Low sample size when you split per site and round
-
-__pollinator traits:__
-```{r}
-pol_tr <- read.csv("traits_pollinators.csv", h = T)
-head(pol_tr)
-```
-__NOTE:__ Most IT (intertegular length) for bees are actual measures.
-For floral visitors without wings we measured torax width.
-When species is unknown (Genus sp.) we calculated the mean of the genus or family.
-Tongue length is measured only for a few individuals.
-Bees tongue length for most species is estimated following Cariveau et al 2016.
-For families without an actual tongue the value is set to zero.
-For other species, tongue length is estimated to covary with allometry using same equation as for colletidaes (the shortest tongue bee family).
-
-__plant traits:__
-```{r}
-plant_tr <- read.csv("traits_plants.csv", h = T)
-head(plant_tr)
-```
-Those are complete.
-
-Challenges:  
-- We are not sure which trait are driving interactions (quite generalized system)
-
-__independent plant flower abundance__: Ca be combined by Site_ID and Round
-```{r}
-flw <- read.csv("flowers.csv", h = T)
-head(flw)
-```
-
-__Phylogenies__ Need to do.
-
-```{r}
-#Nacho this is the final list of species
-pol_tr$Pollinator_gen_sp #277 entries. 
-#sp means we don't know the species. Ignore for phylogeny and use a random species in this genus
-#morpho1,2,3 Unidentified, but we know is a unique species. Idem. Ignore for phylogenies and use a species in genus as replacement?
-#Family NA- means we only know the family. It will be intersting to make sure those families are represented, e.g. by adding a ranom species?
-plant_tr$Plant_gen_sp
-```
-
-- Plants: Phylomatic?
-- Pollinators: Bees available at genus level only. For other orders I have no idea.
-
-# combining data
-
-```{r}
-head(int)
-unique(int[,3:4])
-unique(int$Plant_gen_sp[which(!int$Plant_gen_sp %in% plant_tr$Plant_gen_sp)]) #ok
-unique(int$Pollinator_gen_sp[which(!int$Pollinator_gen_sp %in% pol_tr$Pollinator_gen_sp)])
-temp <- merge(int, pol_tr, by.x = "Pollinator_gen_sp", by.y = "Pollinator_gen_sp", all.x = TRUE)
-temp2 <- merge(temp, plant_tr, by.x = "Plant_gen_sp", by.y = "Plant_gen_sp", all.x = TRUE)
-pooled_data <- temp2
-head(pooled_data, 20)
-```
-
 # Reshaping the dataset as input of the `as.alienData()` function
-
-**Extract from the roxygen2 documentation:**
-
-```
-#' @param idObs A data.frame which is mandatory and will help to check
-consistency and prevent errors among unique identifiers of each alienData
-arguments. The first column (idSite) contains unique identifier of where the
-observation was made. The second column (idTime) is not mandatory and contains
-temporal information: an unique identifier at the time the sample has been taken
-(needed for timeseries analysis). The third column (idSpcies) is an unique
-identifier of the species sampled at time (idTime) and location (idSite). The
-fourth column is an unique identifier of individu of species (idSp) observed at
-time (idTime) and location (idSite). 
-#' @param interactPair A data.frame which
-contains interaction at the finest level (individus or species). The first two
-columns are idFrom and idTo and determine the sens of the interaction. idFrom
-and idTo are unique identifier of species or individu documented in the idObs
-data.frame. Finaly, the thrid column is the strength of the interaction (Please
-see details). 
-#' @param coOcc A square symmetric matrix of 0s and 1s that define
-co-occurence patterns among pairs of species. If this matrix is not provided,
-the co-occurence matrix is derived from the coAbund matrix else the interactSp
-matrix (see return section). 
-#' @param coAbund A square symmetric matrix that
-includes any types of values, defining co-abundance patterns among pairs of
-species. TODO: Not implemented yet. 
-#' @param siteEnv A matrix or a data.frame
-where each column is a descriptor of the sites. TODO: siteEnv should cover the
-possibility that environmental variables could be taken at several times. 
-#' @param traitSp A matrix or a data.frame where each column is a trait
-characterizing all species. The first column is a unique identifier of the
-species documented in idObs data.frame. 
-#' @param traitInd A matrix or a data.frame where each column is a trait characterizing an individual. The first
-column is a unique identifier of the individu documented in idObs data.frame. 
-#' @param phylo A square symmetric matrix describing the phylogenetic relationships
-between pairs of all species (see details). TODO: Not implemented yet. 
-#' @param scaleSiteEnv Logical. Whether the columns of X should be centred and divided by
-the standard deviation. Default is TRUE. 
-#' @param scaleTrait Logical. Whether the rows of Tr should be centred and divided by the standard deviation. Default
-is TRUE. 
-#' @param interceptSiteEnv Logical. Whether a column of 1s should be
-added to X. Default is TRUE. 
-#' @param interceptTrait Logical. Whether a row of
-1s should be added to Tr. Default is TRUE. 
-#' @details 
-#'The strength of the interactions defined in the third column of \code{interactPair} can be a 0
-if no direct interaction has been observed (defined as true absence of
-interaction) or any numerical value. Undocumented interactions among species or
-individus will be assumed as NA by default.
-```
-
-```{r, echo=FALSE}
-# Create dir for formatted data
-#dir.create('./Data/alienData', showWarnings = FALSE)
-```
 

### 2017-07-24 Data/data_explo.Rmd
alienData v2 - new structure after some corrections inherents to methods

--- 2017-02-14 Data/data_explnation.Rmd
@@
----
-title: "Data description"
-author: "I. Bartomeus"
-date: "12/1/2016"
-output: github_document
----
-
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-```
-
-Data collected in 2015 across 16 sites. Each site was visited 7 times. Each time we recorded three data types:
-
-- All interactions in 100 m transects (`transects`)
-- Focal interactions to selected plants (`focal`)
-- Oportunistic new interactions not detected by the last two methods (`out`)
-
-there are three datasets so far:
-
-__Interactions:__ This include `transect`, `visitation` and `out` data.
-Data can be used pooled (1469 interactions ; 5706 visits) for general question abouts trait matching or splited by site (`Site_ID`) and `Round`.
-
-```{r}
-int <- read.csv("interactions.csv", h = T)
-head(int)
-nrow(int)
-sum(int$Frequency, na.rm = TRUE)
-```
-Challenges:  
-- long tail of singletones  
-- not all species are taxonomically resolved. What to do with specimens we ony know Genus?
-As for now:
-  . sp means we only know the genus
-  . morpho1,2,3 Unidentified species, but we know is a unique species diferent from anything else.
-  . Family NA- means we only know the family. 
-- Low sample size when you split per site and round
-
-__pollinator traits:__
-```{r}
-pol_tr <- read.csv("traits_pollinators.csv", h = T)
-head(pol_tr)
-```
-__NOTE:__ Most IT (intertegular length) for bees are actual measures.
-For floral visitors without wings we measured torax width.
-When species is unknown (Genus sp.) we calculated the mean of the genus or family.
-Tongue length is measured only for a few individuals.
-Bees tongue length for most species is estimated following Cariveau et al 2016.
-For families without an actual tongue the value is set to zero.
-For other species, tongue length is estimated to covary with allometry using same equation as for colletidaes (the shortest tongue bee family).
-
-__plant traits:__
-```{r}
-plant_tr <- read.csv("traits_plants.csv", h = T)
-head(plant_tr)
-```
-Those are complete.
-
-Challenges:  
-- We are not sure which trait are driving interactions (quite generalized system)
-
-__independent plant flower abundance__: Ca be combined by Site_ID and Round
-```{r}
-flw <- read.csv("flowers.csv", h = T)
-head(flw)
-```
-
-__Phylogenies__ Need to do.
-
-```{r}
-#Nacho this is the final list of species
-pol_tr$Pollinator_gen_sp #277 entries. 
-#sp means we don't know the species. Ignore for phylogeny and use a random species in this genus
-#morpho1,2,3 Unidentified, but we know is a unique species. Idem. Ignore for phylogenies and use a species in genus as replacement?
-#Family NA- means we only know the family. It will be intersting to make sure those families are represented, e.g. by adding a ranom species?
-plant_tr$Plant_gen_sp
-```
-
-- Plants: Phylomatic?
-- Pollinators: Bees available at genus level only. For other orders I have no idea.
-
-# combining data
-
-```{r}
-head(int)
-unique(int[,3:4])
-unique(int$Plant_gen_sp[which(!int$Plant_gen_sp %in% plant_tr$Plant_gen_sp)]) #ok
-unique(int$Pollinator_gen_sp[which(!int$Pollinator_gen_sp %in% pol_tr$Pollinator_gen_sp)])
-temp <- merge(int, pol_tr, by.x = "Pollinator_gen_sp", by.y = "Pollinator_gen_sp", all.x = TRUE)
-temp2 <- merge(temp, plant_tr, by.x = "Plant_gen_sp", by.y = "Plant_gen_sp", all.x = TRUE)
-pooled_data <- temp2
-head(pooled_data, 20)
-```
-
-# Reshaping the dataset as input of the `as.alienData()` function
-
-**Extract from the roxygen2 documentation:**
-
-```
-#' @param idObs A data.frame which is mandatory and will help to check
-consistency and prevent errors among unique identifiers of each alienData
-arguments. The first column (idSite) contains unique identifier of where the
-observation was made. The second column (idTime) is not mandatory and contains
-temporal information: an unique identifier at the time the sample has been taken
-(needed for timeseries analysis). The third column (idSpcies) is an unique
-identifier of the species sampled at time (idTime) and location (idSite). The
-fourth column is an unique identifier of individu of species (idSp) observed at
-time (idTime) and location (idSite). 
-#' @param interactPair A data.frame which
-contains interaction at the finest level (individus or species). The first two
-columns are idFrom and idTo and determine the sens of the interaction. idFrom
-and idTo are unique identifier of species or individu documented in the idObs
-data.frame. Finaly, the thrid column is the strength of the interaction (Please
-see details). 
-#' @param coOcc A square symmetric matrix of 0s and 1s that define
-co-occurence patterns among pairs of species. If this matrix is not provided,
-the co-occurence matrix is derived from the coAbund matrix else the interactSp
-matrix (see return section). 
-#' @param coAbund A square symmetric matrix that
-includes any types of values, defining co-abundance patterns among pairs of
-species. TODO: Not implemented yet. 
-#' @param siteEnv A matrix or a data.frame
-where each column is a descriptor of the sites. TODO: siteEnv should cover the
-possibility that environmental variables could be taken at several times. 
-#' @param traitSp A matrix or a data.frame where each column is a trait
-characterizing all species. The first column is a unique identifier of the
-species documented in idObs data.frame. 
-#' @param traitInd A matrix or a data.frame where each column is a trait characterizing an individual. The first
-column is a unique identifier of the individu documented in idObs data.frame. 
-#' @param phylo A square symmetric matrix describing the phylogenetic relationships
-between pairs of all species (see details). TODO: Not implemented yet. 
-#' @param scaleSiteEnv Logical. Whether the columns of X should be centred and divided by
-the standard deviation. Default is TRUE. 
-#' @param scaleTrait Logical. Whether the rows of Tr should be centred and divided by the standard deviation. Default
-is TRUE. 
-#' @param interceptSiteEnv Logical. Whether a column of 1s should be
-added to X. Default is TRUE. 
-#' @param interceptTrait Logical. Whether a row of
-1s should be added to Tr. Default is TRUE. 
-#' @details 
-#'The strength of the interactions defined in the third column of \code{interactPair} can be a 0
-if no direct interaction has been observed (defined as true absence of
-interaction) or any numerical value. Undocumented interactions among species or
-individus will be assumed as NA by default.
-```
-
-```{r, echo=FALSE}
-# Create dir for formatted data
-#dir.create('./Data/alienData', showWarnings = FALSE)
-```
-
-### Each section below transform your data into as.alienData args
-
-#### Get idObs
-
-```{r}
-# load data
-int <- read.csv("interactions.csv", h = T, stringsAsFactors=FALSE)
-flw <- read.csv("flowers.csv", h = T, stringsAsFactors=FALSE)
-plant_tr <- read.csv("traits_plants.csv", h = T, stringsAsFactors=FALSE)
-pol_tr <- read.csv("traits_pollinators.csv", h = T, stringsAsFactors=FALSE)
-
-# Create unique IDs for pollinator individus
-int$idPol <- seq(1,nrow(int),1) 
-# Create unique IDs for plant individus
-int$idPlant <- seq(nrow(int)+1,nrow(int)+nrow(int),1) #technically, some individuus will be repeated, although there is not way to know. Good for now
-
-# subset pollinator and plant with their own ids
-pol <- unique(int[,c("Site_ID", "Round","Pollinator_gen_sp","idPol")]) #The unique is unececsary here, right?
-colnames(pol) <- c("idSite","idTime","idSp","idInd")
-plant <- unique(int[,c("Site_ID", "Round","Plant_gen_sp","idPlant")]) #idem
-colnames(plant) <- c("idSite","idTime","idSp","idInd")
-
-# Set idObs by merging both dfs
-idObs <- rbind(pol,plant)
-#head(idObs)
-
-#Turn all columns as factor
-idObs <- as.data.frame(lapply(idObs,as.factor))
-```
-
-#### Get InteracPair
-
-```{r}
-interactPair <- int[,c("idPol","idPlant","Frequency")]
-```
-
-#### traitInd
-
-Trait by individus
-
-```{r}
-traitInd <- int[,c("idPol","Pollinator_sex")] 
-traitInd$traitName <- "sex"
-traitInd <- traitInd[,c('idPol','traitName','Pollinator_sex')]
-names(traitInd)[3] <- 'value'
-#head(traitInd)
-```
-
-#### traitSp
-
-Trait by species
-Note Removed family as a trait
-
-```{r}
-library(reshape2)
-traitSpPlant <- melt(plant_tr[,-2],id=c("Plant_gen_sp"),value.name="value",variable.name="traitName")
-names(traitSpPlant)[1] <- "idSp"
-
-#head(pol_tr)
-traitSpPol <- melt(pol_tr[,c(1,2,5)],id=c("Pollinator_gen_sp"),value.name="value",variable.name="traitName") #I also remove order
-names(traitSpPol)[1] <- "idSp"
-
-traitSp <- rbind(traitSpPlant,traitSpPol)
-traitSp$idSp <- as.factor(traitSp$idSp)
-
-#### errors
-# Some entries in traitSp (merge of pol_tr and plant_tr) are documenting species which are not present in `ìnt`, the interaction pairwise data.frame.
-#IB: This is ok! I collect traits for some species we do not have interactions.
-removeSpecies <- levels(traitSp$idSp)[which(!levels(traitSp$idSp) %in% levels(idObs$idSp))]
-traitSp <- subset(traitSp, !idSp %in% removeSpecies )
-traitSp$idSp <- droplevels(traitSp$idSp)
-```
-
-#### Save all args of the as.alienData() into a RData file.
-
-```{r}
-save(idObs,interactPair,traitSp,traitInd,file="./alienData/argsAlienData.RData")
-```
-
 ```{r}

### 2017-07-26 Data/alien_reshaping.Rmd
Reshaping to alienData v2 / tracking taxo with NCBI and ITIS

@@
 # Create unique IDs for plant individus
-int$idPlant <- seq(nrow(int)+1,nrow(int)+nrow(int),1) #technically, some individuus will be repeated, although there is not way to know. Good for now
+int$idPlant <- seq(nrow(int)+1,nrow(int)+nrow(int),1) # IB: technically, some individus will be repeated, although there is not way to know. Good for now
 
 # subset pollinator and plant with their own ids
-pol <- unique(int[,c("Site_ID", "Round","Pollinator_gen_sp","idPol")]) #The unique is unececsary here, right?
+# if no_captured set his own ids
+# int$idPol <- int$Pollinator_id
+# id <- 1
+# for(i in 1:nrow(int)){
+#   if(int[i,'idPol'] == 'no_captured'){
+#     int[i,'idPol'] <- id
+#     id <- id+1
+#   }
+# }
+
+pol <- int[,c("Site_ID", "Round","Pollinator_gen_sp","idPol")] #The unique is unececsary here, right? SV: yes
 colnames(pol) <- c("idSite","idTime","idSp","idInd")
@@
 idObs <- rbind(pol,plant)
-#head(idObs)
+head(idObs)
 
-#Turn all columns as factor
-idObs <- as.data.frame(lapply(idObs,as.factor))
-```
-
-#### Get InteracPair
-
-```{r}
-interactPair <- int[,c("idPol","idPlant","Frequency")]
 ```
@@
 #head(traitInd)
+```
+
+#### Get InteracPair
+
+```{r}
+interactPair <- int[,c("idPol","idPlant","Frequency")]
 ```
@@
 #IB: This is ok! I collect traits for some species we do not have interactions.
-removeSpecies <- levels(traitSp$idSp)[which(!levels(traitSp$idSp) %in% levels(idObs$idSp))]
-traitSp <- subset(traitSp, !idSp %in% removeSpecies )
-traitSp$idSp <- droplevels(traitSp$idSp)
+# removeSpecies <- levels(traitSp$idSp)[which(!levels(traitSp$idSp) %in% levels(idObs$idSp))]
+# traitSp <- subset(traitSp, !idSp %in% removeSpecies )
+# traitSp$idSp <- droplevels(traitSp$idSp)
 ```
 
-#### Save all args of the as.alienData() into a RData file.
+#### Prepare each aguments for alienData v2
 
 ```{r}
-save(idObs,interactPair,traitSp,traitInd,file="./alienData/argsAlienData.RData")
+
+#### STEP A: dfNodes
+
+dfNodes <- idObs[,c("idInd","idSp")]
+
+## 1. ADD TRAITS FOR SP
+
+dfNodes <- merge(dfNodes,traitSp,by="idSp",all=TRUE)
+dfNodes <- dcast(idSp + idInd ~ traitName, data=dfNodes)
+
+## 2. ADD TRAITS FOR IND
+
+traitInd <- traitInd[,c('idPol','value')]
+names(traitInd)[2] <- "sex"
+dfNodes <- merge(dfNodes,traitInd,by.x="idInd",by.y="idPol",all=TRUE)
+
+## 3. ADD TAXO
+library(stringr)
+dfNodes$species <- word(str_trim(dfNodes$idSp),-1)
+dfNodes$genus <-  word(str_trim(dfNodes$idSp),-2)
+
+## retrieve higher taxonomy based on ITIS
+library(taxize)
+
+cleanup_taxo <- function(x){
+  if(all(!is.na(x))){
+    if(!is.null(x) & dim(x)[1] != 1){
+      r <- data.frame(t(x$name))
+      colnames(r) <- x$rank
+      return(r)
+    } else {
+      return(NA)
+    }
+  } else {
+    return(NA)
+  }
+  }
+
+# resolved_itis <- classification(unique(str_trim(dfNodes$idSp)),db="itis")
+# saveRDS(resolved_itis,file='./resolved_itis.rds')
+resolved_itis <- readRDS('resolved_itis.rds')
+
+res_itis <- lapply(resolved_itis,cleanup_taxo)
+
+res_itis <- res_itis[!is.na(res_itis)]
+res_itis <- plyr::rbind.fill(res_itis)
+res_itis <- res_itis[,names(res_itis)[which(colSums(apply(res_itis,2,is.na))==0)]]
+
+# removing genus
+res_itis <- res_itis[,-which(names(res_itis)=='genus')]
+dfNodes <- merge(dfNodes,res_itis,by.x="idSp",by.y="species",all.x=TRUE)
+
+## retrieve higher MISSING taxonomy based on GENUS with NCBI
+
+taxo_missing <- unique(subset(dfNodes,is.na(family))[,'genus'])
+resolved_ncbi <- readRDS('./resolved_ncbi.rds')
+# resolved_ncbi <- classification(str_trim(taxo_missing),db="ncbi")
+# saveRDS(resolved_ncbi,file='./resolved_ncbi.rds')
+
+resolved_ncbi <- resolved_ncbi[!is.na(resolved_ncbi)]
+res_ncbi <- lapply(resolved_ncbi,cleanup_taxo)
+res_ncbi <- plyr::rbind.fill(res_ncbi)
+res_ncbi <- res_ncbi[,c('genus',names(res_ncbi)[which(colSums(apply(res_ncbi,2,is.na))==0)])]
+
+# remove taxo for species with no genus
+res_ncbi <- subset(res_ncbi,!is.na(genus))
+res_ncbi <- res_ncbi[,-2]
+
+# cast classification items as factors
+cols <- names(dfNodes)[11:19]
+for(c in cols) dfNodes[,c] <- as.character(dfNodes[,c])
+
+# fill missing columns on match
+for(i in 1:nrow(res_ncbi)){
+  dfNodes[which(res_ncbi[i,'genus'] == dfNodes$genus),'kingdom'] <- as.character(res_ncbi[i,'kingdom'])
+  dfNodes[which(res_ncbi[i,'genus'] == dfNodes$genus),'order'] <- as.character(res_ncbi[i,'order'])
+  dfNodes[which(res_ncbi[i,'genus'] == dfNodes$genus),'family'] <- as.character(res_ncbi[i,'family'])
+}
+
+## TODO: STILL TAXO LEVELS MISSINGS
+
+## Final reshaping
+dfNodes <- dfNodes[,c(2:ncol(dfNodes))]
+names(dfNodes)[1] <- "idNodes"
+
+#### STEP B: dfEdges
+
+dfEdges <- int[,c('Site_ID','idPol','idPlant','Frequency')]
+names(dfEdges) <- c('idSite','idFrom','idTo','value')
+
+#### STEP C: dfSites
+
+dfSites <- unique(dfEdges[,'idSite'])
+
+#### Step D: dfOcc
+
+dfOcc <- unique(dfEdges[,c('idSite','idTo')])
+names(dfOcc)[2] <- "idSp"
+dfOccFrom <- unique(dfEdges[,c('idSite','idFrom')])
+names(dfOccFrom)[2] <- "idSp"
+dfOcc <- rbind(dfOcc,dfOccFrom)
+
 ```
+
+### Transform to alienData object
+
+```{r}
+plantsPol <-  alienData(dfNodes=dfNodes,dfEdges=dfEdges,verbose=TRUE,taxo=c('kingdom','class','order','family','genus','species'),trait=c('flower_morphology','flower_form','nectar_tube_diameter_mm','nectar_tube_depth_mm','IT','tongue_length','sex'))
+
+save(plantsPol,file='../../alien/data/plantsPol.rda')
+
+```

### 2017-07-26 Data/data_explnation.Rmd
Reshaping to alienData v2 / tracking taxo with NCBI and ITIS

--- 2017-02-14 Data/data_explnation.Rmd
@@
 ```
-
-# Reshaping the dataset as input of the `as.alienData()` function
-
-**Extract from the roxygen2 documentation:**
-
-```
-#' @param idObs A data.frame which is mandatory and will help to check
-consistency and prevent errors among unique identifiers of each alienData
-arguments. The first column (idSite) contains unique identifier of where the
-observation was made. The second column (idTime) is not mandatory and contains
-temporal information: an unique identifier at the time the sample has been taken
-(needed for timeseries analysis). The third column (idSpcies) is an unique
-identifier of the species sampled at time (idTime) and location (idSite). The
-fourth column is an unique identifier of individu of species (idSp) observed at
-time (idTime) and location (idSite). 
-#' @param interactPair A data.frame which
-contains interaction at the finest level (individus or species). The first two
-columns are idFrom and idTo and determine the sens of the interaction. idFrom
-and idTo are unique identifier of species or individu documented in the idObs
-data.frame. Finaly, the thrid column is the strength of the interaction (Please
-see details). 
-#' @param coOcc A square symmetric matrix of 0s and 1s that define
-co-occurence patterns among pairs of species. If this matrix is not provided,
-the co-occurence matrix is derived from the coAbund matrix else the interactSp
-matrix (see return section). 
-#' @param coAbund A square symmetric matrix that
-includes any types of values, defining co-abundance patterns among pairs of
-species. TODO: Not implemented yet. 
-#' @param siteEnv A matrix or a data.frame
-where each column is a descriptor of the sites. TODO: siteEnv should cover the
-possibility that environmental variables could be taken at several times. 
-#' @param traitSp A matrix or a data.frame where each column is a trait
-characterizing all species. The first column is a unique identifier of the
-species documented in idObs data.frame. 
-#' @param traitInd A matrix or a data.frame where each column is a trait characterizing an individual. The first
-column is a unique identifier of the individu documented in idObs data.frame. 
-#' @param phylo A square symmetric matrix describing the phylogenetic relationships
-between pairs of all species (see details). TODO: Not implemented yet. 
-#' @param scaleSiteEnv Logical. Whether the columns of X should be centred and divided by
-the standard deviation. Default is TRUE. 
-#' @param scaleTrait Logical. Whether the rows of Tr should be centred and divided by the standard deviation. Default
-is TRUE. 
-#' @param interceptSiteEnv Logical. Whether a column of 1s should be
-added to X. Default is TRUE. 
-#' @param interceptTrait Logical. Whether a row of
-1s should be added to Tr. Default is TRUE. 
-#' @details 
-#'The strength of the interactions defined in the third column of \code{interactPair} can be a 0
-if no direct interaction has been observed (defined as true absence of
-interaction) or any numerical value. Undocumented interactions among species or
-individus will be assumed as NA by default.
-```
-
-```{r, echo=FALSE}
-# Create dir for formatted data
-#dir.create('./Data/alienData', showWarnings = FALSE)
-```
-
-### Each section below transform your data into as.alienData args
-
-#### Get idObs
-
-```{r}
-# load data
-int <- read.csv("interactions.csv", h = T, stringsAsFactors=FALSE)
-flw <- read.csv("flowers.csv", h = T, stringsAsFactors=FALSE)
-plant_tr <- read.csv("traits_plants.csv", h = T, stringsAsFactors=FALSE)
-pol_tr <- read.csv("traits_pollinators.csv", h = T, stringsAsFactors=FALSE)
-
-# Create unique IDs for pollinator individus
-int$idPol <- seq(1,nrow(int),1) 
-# Create unique IDs for plant individus
-int$idPlant <- seq(nrow(int)+1,nrow(int)+nrow(int),1) #technically, some individuus will be repeated, although there is not way to know. Good for now
-
-# subset pollinator and plant with their own ids
-pol <- unique(int[,c("Site_ID", "Round","Pollinator_gen_sp","idPol")]) #The unique is unececsary here, right?
-colnames(pol) <- c("idSite","idTime","idSp","idInd")
-plant <- unique(int[,c("Site_ID", "Round","Plant_gen_sp","idPlant")]) #idem
-colnames(plant) <- c("idSite","idTime","idSp","idInd")
-
-# Set idObs by merging both dfs
-idObs <- rbind(pol,plant)
-#head(idObs)
-
-#Turn all columns as factor
-idObs <- as.data.frame(lapply(idObs,as.factor))
-```
-
-#### Get InteracPair
-
-```{r}
-interactPair <- int[,c("idPol","idPlant","Frequency")]
-```
-
-#### traitInd
-
-Trait by individus
-
-```{r}
-traitInd <- int[,c("idPol","Pollinator_sex")] 
-traitInd$traitName <- "sex"
-traitInd <- traitInd[,c('idPol','traitName','Pollinator_sex')]
-names(traitInd)[3] <- 'value'
-#head(traitInd)
-```
-
-#### traitSp
-
-Trait by species
-Note Removed family as a trait
-
-```{r}
-library(reshape2)
-traitSpPlant <- melt(plant_tr[,-2],id=c("Plant_gen_sp"),value.name="value",variable.name="traitName")
-names(traitSpPlant)[1] <- "idSp"
-
-#head(pol_tr)
-traitSpPol <- melt(pol_tr[,c(1,2,5)],id=c("Pollinator_gen_sp"),value.name="value",variable.name="traitName") #I also remove order
-names(traitSpPol)[1] <- "idSp"
-
-traitSp <- rbind(traitSpPlant,traitSpPol)
-traitSp$idSp <- as.factor(traitSp$idSp)
-
-#### errors
-# Some entries in traitSp (merge of pol_tr and plant_tr) are documenting species which are not present in `ìnt`, the interaction pairwise data.frame.
-#IB: This is ok! I collect traits for some species we do not have interactions.
-removeSpecies <- levels(traitSp$idSp)[which(!levels(traitSp$idSp) %in% levels(idObs$idSp))]
-traitSp <- subset(traitSp, !idSp %in% removeSpecies )
-traitSp$idSp <- droplevels(traitSp$idSp)
-```
-
-#### Save all args of the as.alienData() into a RData file.
-
-```{r}
-save(idObs,interactPair,traitSp,traitInd,file="./alienData/argsAlienData.RData")
-```
